---
title: "model_description"
output: html_document
date: "2025-08-23"
---


README: walking through the basic model dynamics (updated with parameters from the literature)

```{r setup, include=FALSE}

# command + option + i for new code chunk
# command + shift + enter to run entire code chunk
# command + enter to run single line of code

library(tidyverse) # command + shift + m for %>%

```


load the functions

```{r}

source("coral_demographic_funs.R")
source("rse_funs.R")
```


import the parameter values


```{r}
# fun_test <- function(x){
#   
#   if(x > 1){
#     y <- NA
#     print("error")
#   } else{
#     y <- x + 1
#   }
#   
#   return(y)
# }
# 
# fun_test(0)
# fun_test(2)

```


```{r}

# extra_x <- 5
# xset <- c(7, 0, 3, 6, 2)
# #xset <- 6
# removed_i <- rep(NA, extra_x)
# for(i in 1:extra_x){
#   indx <- which(xset == max(xset))[1]
#   xset[indx] <- xset[indx] - 1
#   removed_i[i] <- indx
#
# }
#
# xset
# removed_i

# max_x <- 10
# xset <- c(1, 11)
#
# prop_fits_x <- max_x/sum(xset)
#
# xset*prop_fits_x
#
# sum(xset*prop_fits_x)

max_x <- 10
xset <- c(1, 11) # sum = 12


```



```{r}
# NOTE: Data files are stored in the separate Detmer-2025-coral-parameters repository.
# Update DATA_PATH to point to your local copy of that repository.
DATA_PATH <- "../Detmer-2025-coral-parameters"

# fragmentation data
apal_frag_summ <- read.csv(file.path(DATA_PATH, "standardized_data/apal_fragmentation_summ.csv"))
apal_frag <- read.csv(file.path(DATA_PATH, "standardized_data/apal_fragmentation.csv"))


# survival data
field_surv <- readRDS(file.path(DATA_PATH, "parameter_lists/field_surv_pars.rds"))
nurs_surv <- readRDS(file.path(DATA_PATH, "parameter_lists/nurs_surv_pars.rds"))
lab_surv <- readRDS(file.path(DATA_PATH, "parameter_lists/lab_surv_pars.rds"))


# growth data
field_growth <- readRDS(file.path(DATA_PATH, "parameter_lists/field_growth_pars.rds"))
nurs_growth <- readRDS(file.path(DATA_PATH, "parameter_lists/nurs_growth_pars.rds"))
lab_growth <- readRDS(file.path(DATA_PATH, "parameter_lists/lab_growth_pars.rds"))

# short-term lab survival data
apal_lab_short_surv <- read.csv(file.path(DATA_PATH, "standardized_data/apal_surv_lab_short.csv"))


#View(field_surv$SC_surv_summ_df)

#View(field_surv$SC_surv_df)

# col_choice <- "mean"
# 
# field_growth$summ_list[[1]][-c(1:2) , col_choice] # 1 = first col, 2 = second, etc. 
# 
# field_growth$summ_list[[5]][ , col_choice] 
# 
# field_growth$mat_list[[1]][1:10, ]
# 
# nrow(field_growth$mat_list[[1]])


#View(apal_frag_summ)

```


```{r}
# test parameter function
# 
# # survival
# par_test <- par_list_fun(par_type = "survival", sample_dt = F, summ_df = field_surv$SC_surv_summ_df, summ_metric = "mean", full_df = NA, n_sample = NA)
# 
# par_test <- par_list_fun(par_type = "survival", sample_dt = T, summ_df = NA, summ_metric = NA, full_df = field_surv$SC_surv_df, n_sample = 5)
# 
# par_test$surv_pars
# 
# # growth/shrinkage
# par_test <- par_list_fun(par_type = "growth", sample_dt = F, summ_df = field_growth$summ_list, summ_metric = "mean", full_df = NA, n_sample = NA)
# 
# par_test$growth_pars
# par_test$shrink_pars
# 
# par_test <- par_list_fun(par_type = "growth", sample_dt = T, summ_df = NA, summ_metric = NA, full_df = field_growth$mat_list, n_sample = 5)
# 
# par_test$growth_pars[[1]]
# par_test$shrink_pars[[1]]
# 
# # fragmentation
# par_test <- par_list_fun(par_type = "fragmentation", sample_dt = F, summ_df = apal_frag_summ, summ_metric = "mean", full_df = NA, n_sample = NA)
# 
# par_test <- par_list_fun(par_type = "fragmentation", sample_dt = T, summ_df = NA, summ_metric = NA, full_df = apal_frag, n_sample = 10)
# 
# par_test$frag_pars[[1]]
# 
# 
# #View(field_surv$SC_surv_summ_df)
# 
# 
# #field_growth$summ_list[[1]]
# 
# par_test <- par_list_fun(par_type = "survival", sample_dt = F, summ_df = nurs_surv$SC_surv_summ_df, summ_metric = "mean", full_df = NA, n_sample = NA)
# 
# par_test <- par_list_fun(par_type = "growth", sample_dt = F, summ_df = nurs_growth$summ_list, summ_metric = "mean", full_df = NA, n_sample = NA)
# 
# nurs_growth$summ_list
# 
# 
# par_test # remember the nursery data don't include the larger size classes so will need to fill in with the field data. Do this outside the function (modify the input data frames)

```


```{r}
# test summarized parameter function

# field_surv1 <- field_surv
# field_growth1 <- field_growth
# nurs_surv1 <- nurs_surv
# nurs_growth1 <- nurs_growth
# 
# # nursery data only available for size classes 1-2, so use the field data for the larger size classes
# nurs_surv1$SC_surv_summ_df <- nurs_surv1$SC_surv_summ_df %>% filter(size_class < 3) %>% rbind(field_surv1$SC_surv_summ_df[field_surv1$SC_surv_summ_df$size_class > 2,])
# 
# nurs_growth1$summ_list[[3]] <- field_growth1$summ_list[[3]]
# nurs_growth1$summ_list[[4]] <- field_growth1$summ_list[[4]]
# nurs_growth1$summ_list[[5]] <- field_growth1$summ_list[[5]]
# 
# summ_pars_test1 <- default_pars_fun(n_reef = 1, n_orchard = 1, n_lab = 2, summ_metric_list = list(field_surv = "mean", field_growth = "mean", field_shrink = "mean", field_frag = "mean", nurs_surv = "mean", nurs_growth = "mean", nurs_shrink = "mean"), field_surv1, field_growth1, nurs_surv1, nurs_growth1, apal_frag_summ)
#  
# summ_pars_test1$surv_pars.r
# 
# 
# summ_pars_test1$frag_pars.o

```


```{r}
# test random parameter function
# field_surv1 <- field_surv
# field_growth1 <- field_growth
# nurs_surv1 <- nurs_surv
# nurs_growth1 <- nurs_growth
# 
# # nursery data only available for size classes 1-2, so use the field data for larger size classes
# nurs_surv1$SC_surv_df <- nurs_surv1$SC_surv_df %>% filter(size_class < 3) %>% rbind(field_surv1$SC_surv_df[field_surv1$SC_surv_df$size_class > 2,])
# 
# nurs_growth1$mat_list[[3]] <- field_growth1$mat_list[[3]]
# nurs_growth1$mat_list[[4]] <- field_growth1$mat_list[[4]]
# nurs_growth1$mat_list[[5]] <- field_growth1$mat_list[[5]]
# 
# rand_test1 <- rand_pars_fun(n_reef = 2, n_orchard = 1, n_lab = 2, n_sample = 2, field_surv = field_surv1, field_growth = field_growth1, nurs_surv = nurs_surv1, nurs_growth = nurs_growth1, apal_frag)
# 
# rand_test1$surv_pars_L.r[[1]]
# 
# #rand_test1$frag_pars_L.o[[1]]

```



```{r}

years <- 20
n <- 5

# define size classes
SC1 <- 0
SC2 <- 10
SC3 <- 100
SC4 <- 900
SC5 <- 4000

A_mids <- c((SC1 + SC2)/2, (SC2 + SC3)/2, (SC3 + SC4)/2, (SC4 + SC5)/2, 9325) # midpoint areas of each size class (cm^2); last value is the 50% quantile of observations greater than SC5 in the standardized data from the literature

A_reef <- mean(c(4492.011, 10076.67, 8943.55)) # area of the reef

orchard_treatments <- c("orchard1") # orchards (could have multiple orchards and/or orchards with different post-outplanting treatments)
reef_treatments <- c("reef1") # reef treatments/subpopulations (could have "urchin outplanting" or "algal removal" or other postoutplanting treatments)
lab_treatments <- c("0_T1", "0_T2") # lab (cement = T1, ceramic = T2). "TX" is tile type, "X_" indicates whether recruits are outplanted immediately (0_) or the next year (0_1)
#lab_treatments <- c("0_T1") # lab (cement = T1, ceramic = T2). "TX" is tile type, "X_" indicates whether recruits are outplanted immediately (0_) or the next year (0_1)


# demographic parameters for each orchard and reef treatment/subpopulation
field_surv1 <- field_surv
field_growth1 <- field_growth
nurs_surv1 <- nurs_surv
nurs_growth1 <- nurs_growth

# nursery data only available for size classes 1-2, so use the field data for the larger size classes
nurs_surv1$SC_surv_summ_df <- nurs_surv1$SC_surv_summ_df %>% filter(size_class < 3) %>% rbind(field_surv1$SC_surv_summ_df[field_surv1$SC_surv_summ_df$size_class > 2,])

nurs_growth1$summ_list[[3]] <- field_growth1$summ_list[[3]]
nurs_growth1$summ_list[[4]] <- field_growth1$summ_list[[4]]
nurs_growth1$summ_list[[5]] <- field_growth1$summ_list[[5]]

# calculate mean parameter values (survival, growth, shrinkage, fragmentation) for each 
all_pars <- default_pars_fun(n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), summ_metric_list = list(field_surv = "mean", field_growth = "mean", field_shrink = "mean", field_frag = "mean", nurs_surv = "mean", nurs_growth = "mean", nurs_shrink = "mean"), field_surv1, field_growth1, nurs_surv1, nurs_growth1, apal_frag_summ)

# reef parameters: survival
surv_pars.r <- all_pars$surv_pars.r

# reef parameters: growth
growth_pars.r <- all_pars$growth_pars.r

# reef parameters: shrinkage
shrink_pars.r <- all_pars$shrink_pars.r

# reef parameters: fragmentation
frag_pars.r <- all_pars$frag_pars.r

# density dependent survival (smallest size class only for now)
dens_pars.r <- list()
dens_pars.r[[1]] <- list() # first reef treatment/subpop
dens_pars.r[[1]][[1]] <- 0.02 # first source to first reef (external recruits)
dens_pars.r[[1]][[2]] <- dens_pars.r[[1]][[1]] # second source to first reef (first lab treatment)
dens_pars.r[[1]][[3]] <- dens_pars.r[[1]][[1]] # third source to first reef (second lab treatment)

# reef parameters: fecundity (for tracking reef reproductive output)
fec_pars.r <- list()
fec_pars.r[[1]] <- list()
fec_pars.r[[1]][[1]] <- c(0, 0, 5000, 50000, 100000)
fec_pars.r[[1]][[2]] <- fec_pars.r[[1]][[1]]
fec_pars.r[[1]][[3]] <- fec_pars.r[[1]][[1]]

# orchard parameters: survival
surv_pars.o <- all_pars$surv_pars.o

# orchard parameters: growth
growth_pars.o <- all_pars$growth_pars.o

# orchard parameters: shrinkage
shrink_pars.o <- all_pars$shrink_pars.o

# orchard parameters: fragmentation
frag_pars.o <- all_pars$frag_pars.o

# density dependent survival (smallest size class only for now)
dens_pars.o <- list()
dens_pars.o[[1]] <- list() # first orchard
dens_pars.o[[1]][[1]] <- 0.02
dens_pars.o[[1]][[2]] <- dens_pars.o[[1]][[1]]

# orchard parameters: fecundity
fec_pars.o <- list()
fec_pars.o[[1]] <- list() # first treatment
fec_pars.o[[1]][[1]] <- c(0, 0, 5000, 50000, 100000) # first source
fec_pars.o[[1]][[2]] <- fec_pars.o[[1]][[1]] # second source

lambda <- 0 # external recruitment to reef

#lambda_R <- 10000 # mean number of larvae produced by reference reefs each year
lambda_R <- 1255111 # from Table 1 of Fundemar's report (total embryos collected in 2025)

# stochasticity parameters
sigma_s <- 0
sigma_f <- 0
ext_rand <- c(FALSE, FALSE)
seeds <- c(1000, 5000, 10000, 40000)

# initial conditions in each reef subpopulation
N0.r <- list()
N0.r[[1]] <- list() # second reef subpop
N0.r[[1]][[1]] <- rep(0, n) # first source (external recruits)
N0.r[[1]][[2]] <- rep(0, n) # second source
N0.r[[1]][[3]] <- rep(0, n) # third source

# initial conditions in each orchard subpopulation
N0.o <- list()
N0.o[[1]] <- list() # first orchard treatment
N0.o[[1]][[1]] <- rep(0, n) # first source
N0.o[[1]][[2]] <- rep(0, n) # second source

N0.l <- list()
N0.l[[1]] <- 0
N0.l[[2]] <- 0


# disturbance parameters
dist_yrs <- c(years + 10) # years when disturbance occurs, can set to > years for none

# effects of disturbance on each reef subpop
dist_effects.r <- list()
dist_effects.r[[1]] <- list() # list where each element is the effects of disturbance on corals from each source in the 1st reef subpop (either "survival" for survival, "Tmat" for growth/survival, "Fmat" for fragmentation, or "fecundity" for reproduction)
dist_effects.r[[1]][[1]] <- list() # effects of disturbances on corals from first source in first reef subpop
dist_effects.r[[1]][[1]][[1]] <- c("survival") # effects of 1st disturbance on corals from first source in first reef subpop ([[1]][[1]][[2]] would be second disturbance, etc.)
dist_effects.r[[1]][[2]] <- list() # second source
dist_effects.r[[1]][[2]][[1]] <- c("survival")
dist_effects.r[[1]][[3]] <- list() # third source
dist_effects.r[[1]][[3]][[1]] <- c("survival")


# disturbance parameters for each reef subpop (what happens to the population parameters at each disturbance eve)
dist_pars.r <- list()

# first reef population
dist_pars.r[[1]] <- list() 
# first source to second reef subpop
dist_pars.r[[1]][[1]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[1]], dist_surv0 = list(surv_pars.r[[1]][[1]]*0.1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL) # list of disturbance parameters for the first source to the first reef population (defaults for each parameter type are NULL unless the disturbance affects them, as specified in dist_effects)
# second source to second reef subpop
dist_pars.r[[1]][[2]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[2]], dist_surv0 = list(surv_pars.r[[1]][[2]]*0.1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)
# third source to second reef subpop
dist_pars.r[[1]][[3]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[3]], dist_surv0 = list(surv_pars.r[[1]][[3]]*0.1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)

# disturbance effects for each orchard subpop
dist_effects.o <- list()

dist_effects.o[[1]] <- list() 
dist_effects.o[[1]][[1]] <- list() # effects of disturbances on corals from first source in first orchard treatment
dist_effects.o[[1]][[1]][[1]] <- c("survival") # effects of 1st disturbance on corals from first source in first reef subpop ([[1]][[1]][[2]] would be second disturbance, etc.)
dist_effects.o[[1]][[2]] <- list() # second source
dist_effects.o[[1]][[2]][[1]] <- c("survival")


# disturbance parameters for each orchard subpop
dist_pars.o <- list()

# first orchard treatment
dist_pars.o[[1]] <- list() 
# first source in first orchard treatment
dist_pars.o[[1]][[1]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.o[[1]][[1]], dist_surv0 = list(surv_pars.o[[1]][[1]]*0.1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL) 
# second source in first orchard treatment
dist_pars.o[[1]][[2]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.o[[1]][[2]], dist_surv0 = list(surv_pars.o[[1]][[2]]*0.1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)


# lab parameters
s0 <- c(0.95, 0.95) # fraction of settled larvae in each lab treatment that survive to immediate outplanting
s1 <- c(0.7, 0.7) # fraction of settled larvae in each lab treatment that survive to 1 yr outplanting
m0 <- c(0.02, 0.02) # density dependent mortality rate of larvae in each lab treatment that get outplanted immediately
m1 <- c(0.02, 0.02) # density dependent mortality rate of larvae in each lab treatment that get outplanted after 1 year
sett_props <- list(T1 = 0.15, T2 = 0.15) # proportion of larvae that settle on each tile type
size_props <- matrix(NA, nrow = length(lab_treatments), ncol = n) # matrix for fractions of retained recruits in each size class at the end of their year in the lab
size_props[1, ] <- c(1, 0, 0, 0, 0) # first lab treatment (0_T1)
size_props[2, ] <- c(1, 0, 0, 0, 0) # second lab treatment (0_T2)
lab_pars <- list(s0 = s0, s1 = s1, m0 = m0, m1 = m1, sett_props = sett_props, size_props = size_props)

# restoration strategy parameters
tile_props <- list(T1 = 0.25, T2 = 0.75) # proportion of tiles that are each type
orchard_yield <- 0 # percent of new orchard babies successfully collected
reef_yield <- 1 # percent of new reef babies successfully collected and fertilized
spawn_target <- 2000000 # target number of embryos to collect each year (note: if this is set to zero, there won't be any settlers)
reef_prop <- c(1, 1) # proportion of lab recruits from each lab treatment outplanted to reef (1-proportion outplanted to orchard)
reef_out_props <- matrix(NA, nrow = length(lab_treatments), ncol = length(reef_treatments)) # proportion recruits going from each lab treatment to each reef treatment (row = origin lab treatment, column = destination reef treatment)
reef_out_props[1,] <- c(1) # from first lab treatment to each reef treatment (here there's just on intervention reef so they all go there)
reef_out_props[2,] <- c(1) # from second lab treatment to each reef treatment 

orchard_out_props <- matrix(NA, nrow = length(lab_treatments), ncol = length(orchard_treatments)) # proportion recruits going from each lab treatment to each orchard treatment
orchard_out_props[1,] <- c(1) # from first lab treatment to each orchard 
orchard_out_props[2,] <- c(1) # from second lab treatment to each orchard 

# sizes allocated to each treatment
reef_areas <- c(A_reef)*10000 # m^2 given to each post-outplanting reef treatment/subpopulation, convert to cm^2 by multiplying by 10,000
lab_max <- 3100 # total number of tiles that the lab can accommodate 
lab_retain_max <- 0 # total number of settlers that the lab can keep for a year (must be less than or equal to lab_max; if > 0 then must have at least one "1_" lab treatment, if equal to lab max then must have all treatments be "1_")
orchard_size <- c(30*80) # number of tiles each orchard has space for (500 stars x 30 tiles per star)

# coral transplanting
transplant <- rep(0, years) # years in which corals get transplanted from the orchard
#transplant[10] <- 1 # in 10th year, move corals from orchard to reef

null_mat <- matrix(0, nrow = years, ncol = n)
#rest_pars$trans_mats[[1]][[1]][which(transplant!=0)[1],] <- c(0, 0, 0, 2,0)# max number of colonies of each size class to move to reef in first transplant event
trans_mats <- list()
trans_mats[[1]] <- list() # first orchard 
trans_mats[[1]][[1]] <- null_mat # number of corals in each size class that originated from the first source of colonies for this orchard to transplant at each timepoint
trans_mats[[1]][[2]] <- null_mat # number of corals in each size class that originated from the second source of colonies for this orchard to transplant at each timepoint

# trans_mats[[2]] <- list() # second orchard
# trans_mats[[2]][[1]] <- null_mat # first source
# trans_mats[[2]][[2]] <- null_mat # second source

# where on the reef to put the transplants
#rest_pars$trans_reef[[1]][[1]][which(transplant!=0)[1],] <- c(1, 1)# reef and source subpop to transport the corals to
trans_reef <- list()
trans_reef[[1]] <- list() # first orchard
trans_reef[[1]][[1]] <- matrix(c(1, 2), nrow = years, ncol = 2, byrow = T) # first source corals in this orchard get transplanted to the second source of corals in the first reef
trans_reef[[1]][[2]] <- matrix(c(1, 3), nrow = years, ncol = 2, byrow = T) # second source corals in this orchard get transplanted to the third source of corals in the second reef

rest_pars <- list(tile_props = tile_props, orchard_yield = orchard_yield, reef_yield = reef_yield, spawn_target = spawn_target, reef_prop = reef_prop, reef_out_props = reef_out_props, orchard_out_props = orchard_out_props, reef_areas = reef_areas, lab_max = lab_max, lab_retain_max = lab_retain_max, orchard_size = orchard_size, transplant = transplant, trans_mats = trans_mats, trans_reef = trans_reef)


sim1 <- rse_mod1(years, n, A_mids, surv_pars.r, dens_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, surv_pars.o, dens_pars.o, growth_pars.o, shrink_pars.o, frag_pars.o, fec_pars.o, lambda, lambda_R, sigma_s, sigma_f, ext_rand, seeds, dist_yrs, dist_pars.r, dist_effects.r, dist_pars.o, dist_effects.o, orchard_treatments, reef_treatments, lab_treatments, lab_pars, rest_pars, N0.r, N0.o, N0.l)

```



check numbers (note: getting the numbers to match perfectly for the reported embryos collected and recruits outplanted is a matter of picking the in-lab density dependent and independent mortality rates)

```{r}
sim1$reef_out[[1]][[2]] + sim1$reef_out[[1]][[3]]

(sim1$reef_out[[1]][[2]] + sim1$reef_out[[1]][[3]])/lab_max


#sim1$orchard_out
```
53087.55

```{r}
# plot results

#reef1_tot <- apply(sim1$reef_pops_pre[[1]][[1]], MARGIN = 2, sum) + apply(sim1$reef_pops_pre[[1]][[2]], MARGIN = 2, sum) + apply(sim1$reef_pops_pre[[1]][[3]], MARGIN = 2, sum)
#orch1_tot <- apply(sim1$orchard_pops_pre[[1]][[1]], MARGIN = 2, sum) + apply(sim1$orchard_pops_pre[[1]][[2]], MARGIN = 2, sum)

reef1_tot <- model_summ(model_sim = sim1, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

orch1_tot <- model_summ(model_sim = sim1, location = "orchard", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))


plot(x = c(1:years), y = reef1_tot, type = "l", las = 1, ylim = c(0, max(max(reef1_tot), max(orch1_tot))), xlab = "year", ylab = "Pop'n size", lwd = 2)
lines(x = c(1:years), y = orch1_tot, type = "l", col = "dodgerblue", lwd = 2, lty = 2)
legend("topleft", legend = c("reef", "orchard"), col = c("black", "dodgerblue"), lwd = 2, lty = c(1, 2))


```

test disturbance

```{r}
dist_yrs <- c(10) # years when disturbance occurs, can set to > years or NA for none


# disturbance parameters for each reef subpop (what happens to the population parameters at each disturbance eve)
dist_pars.r <- list()

# first reef population
dist_pars.r[[1]] <- list() 
# first source to second reef subpop
dist_pars.r[[1]][[1]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[1]], dist_surv0 = list(surv_pars.r[[1]][[1]]*0.1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)# list of disturbance parameters for the first source to the first reef population (defaults for each parameter type are NULL unless the disturbance affects them, as specified in dist_effects)
# dist_surv0 = list where ith element is the survival probabilities for the ith disturbance
# second source to second reef subpop
dist_pars.r[[1]][[2]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[2]], dist_surv0 = list(surv_pars.r[[1]][[2]]*0.1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)
# third source to second reef subpop
dist_pars.r[[1]][[3]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[3]], dist_surv0 = list(surv_pars.r[[1]][[3]]*1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)


# disturbance parameters for each orchard 
dist_pars.o <- list()

# first orchard treatment
dist_pars.o[[1]] <- list() 
# first source in first orchard treatment
dist_pars.o[[1]][[1]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.o[[1]][[1]], dist_surv0 = list(surv_pars.o[[1]][[1]]*1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL) 
# second source in first orchard treatment
dist_pars.o[[1]][[2]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.o[[1]][[2]], dist_surv0 = list(surv_pars.o[[1]][[2]]*0.1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)


sim1 <- rse_mod1(years, n, A_mids, surv_pars.r, dens_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, surv_pars.o, dens_pars.o, growth_pars.o, shrink_pars.o, frag_pars.o, fec_pars.o, lambda, lambda_R, sigma_s, sigma_f, ext_rand, seeds, dist_yrs, dist_pars.r, dist_effects.r, dist_pars.o, dist_effects.o, orchard_treatments, reef_treatments, lab_treatments, lab_pars, rest_pars, N0.r, N0.o, N0.l)

```


```{r}

# reef1_tot <- apply(sim1$reef_pops_pre[[1]][[1]], MARGIN = 2, sum) + apply(sim1$reef_pops_pre[[1]][[2]], MARGIN = 2, sum) + apply(sim1$reef_pops_pre[[1]][[3]], MARGIN = 2, sum)
# orch1_tot <- apply(sim1$orchard_pops_pre[[1]][[1]], MARGIN = 2, sum) + apply(sim1$orchard_pops_pre[[1]][[2]], MARGIN = 2, sum)

reef1_tot <- model_summ(model_sim = sim1, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

orch1_tot <- model_summ(model_sim = sim1, location = "orchard", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

plot(x = c(1:years), y = reef1_tot, type = "l", las = 1, ylim = c(0, max(max(reef1_tot), max(orch1_tot))), xlab = "year", ylab = "Pop'n size", lwd = 2)
lines(x = c(1:years), y = orch1_tot, type = "l", col = "dodgerblue", lwd = 2, lty = 2)
legend("topleft", legend = c("reef", "orchard"), col = c("black", "dodgerblue"), lwd = 2, lty = c(1, 2))

plot(x = c(1:years), y = apply(sim1$reef_pops_pre[[1]][[1]], MARGIN = 2, sum), type = "l", las = 1, ylim = c(0, max(max(reef1_tot), max(orch1_tot))), xlab = "year", ylab = "Pop'n size", lwd = 2)
lines(x = c(1:years), y = apply(sim1$reef_pops_pre[[1]][[2]], MARGIN = 2, sum), lty = 2)
lines(x = c(1:years), y = apply(sim1$reef_pops_pre[[1]][[3]], MARGIN = 2, sum), lty = 3)

plot(x = c(1:years), y = apply(sim1$orchard_pops_pre[[1]][[1]], MARGIN = 2, sum), type = "l", col = "dodgerblue", las = 1, ylim = c(0, max(max(reef1_tot), max(orch1_tot))), xlab = "year", ylab = "Pop'n size", lwd = 2)
lines(x = c(1:years), y = apply(sim1$orchard_pops_pre[[1]][[2]], MARGIN = 2, sum), col = "dodgerblue", lty = 2)


```



test transplanting


```{r}
dist_yrs <- c(years + 10)

# coral transplanting
transplant <- rep(0, years) # years in which corals get transplanted from the orchard
transplant[10] <- 1 # in 10th year, move corals from orchard to reef

null_mat <- matrix(0, nrow = years, ncol = n)
null_mat[which(transplant!=0)[1],] <- c(0, 1, 0, 0,0)
trans_mats <- list()
trans_mats[[1]] <- list() # first orchard 
trans_mats[[1]][[1]] <- null_mat # number of corals in each size class that originated from the first source of colonies for this orchard to transplant at each timepoint
trans_mats[[1]][[2]] <- null_mat # number of corals in each size class that originated from the second source of colonies for this orchard to transplant at each timepoint

# trans_mats[[2]] <- list() # second orchard
# trans_mats[[2]][[1]] <- null_mat # first source
# trans_mats[[2]][[2]] <- null_mat # second source

# where on the reef to put the transplants
#rest_pars$trans_reef[[1]][[1]][which(transplant!=0)[1],] <- c(1, 1)# reef and source subpop to transport the corals to
trans_reef <- list()
trans_reef[[1]] <- list() # first orchard
trans_reef[[1]][[1]] <- matrix(c(1, 2), nrow = years, ncol = 2, byrow = T) # first source corals in this orchard get transplanted to the second source of corals in the first reef
trans_reef[[1]][[2]] <- matrix(c(1, 3), nrow = years, ncol = 2, byrow = T) # second source corals in this orchard get transplanted to the third source of corals in the second reef

#rest_pars <- list(orchard_yield = orchard_yield, reef_yield = reef_yield, reef_prop = reef_prop, reef_out_props = reef_out_props, orchard_out_props = orchard_out_props, reef_areas = reef_areas, lab_max = lab_max, lab_retain_max = lab_retain_max, orchard_size = orchard_size, transplant = transplant, trans_mats = trans_mats, trans_reef = trans_reef)


rest_pars <- list(tile_props = tile_props, orchard_yield = orchard_yield, reef_yield = reef_yield, reef_prop = reef_prop, reef_out_props = reef_out_props, orchard_out_props = orchard_out_props, reef_areas = reef_areas, lab_max = lab_max, lab_retain_max = lab_retain_max, orchard_size = orchard_size, transplant = transplant, trans_mats = trans_mats, trans_reef = trans_reef)

sim1 <- rse_mod1(years, n, A_mids, surv_pars.r, dens_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, surv_pars.o, dens_pars.o, growth_pars.o, shrink_pars.o, frag_pars.o, fec_pars.o, lambda, lambda_R, sigma_s, sigma_f, ext_rand, seeds, dist_yrs, dist_pars.r, dist_effects.r, dist_pars.o, dist_effects.o, orchard_treatments, reef_treatments, lab_treatments, lab_pars, rest_pars, N0.r, N0.o, N0.l)


```


```{r}
reef1_tot <- apply(sim1$reef_pops_pre[[1]][[1]], MARGIN = 2, sum) + apply(sim1$reef_pops_pre[[1]][[2]], MARGIN = 2, sum) + apply(sim1$reef_pops_pre[[1]][[3]], MARGIN = 2, sum)
orch1_tot <- apply(sim1$orchard_pops_pre[[1]][[1]], MARGIN = 2, sum) + apply(sim1$orchard_pops_pre[[1]][[2]], MARGIN = 2, sum)

plot(x = c(1:years), y = reef1_tot, type = "l", las = 1, ylim = c(0, max(max(reef1_tot), max(orch1_tot))), xlab = "year", ylab = "Pop'n size", lwd = 2)
lines(x = c(1:years), y = orch1_tot, type = "l", col = "dodgerblue", lwd = 2, lty = 1)
legend("topleft", legend = c("reef", "orchard"), col = c("black", "dodgerblue"), lwd = 2, lty = c(1, 1))

plot(x = c(1:years), y = apply(sim1$reef_pops_pre[[1]][[1]], MARGIN = 2, sum), type = "l", las = 1, ylim = c(0, max(max(reef1_tot), max(orch1_tot))), xlab = "year", ylab = "Pop'n size", lwd = 2)
lines(x = c(1:years), y = apply(sim1$reef_pops_pre[[1]][[2]], MARGIN = 2, sum), lty = 2)
lines(x = c(1:years), y = apply(sim1$reef_pops_pre[[1]][[3]], MARGIN = 2, sum), lty = 3)

plot(x = c(1:years), y = apply(sim1$orchard_pops_pre[[1]][[1]], MARGIN = 2, sum), type = "l", col = "dodgerblue", las = 1, ylim = c(0, max(max(reef1_tot), max(orch1_tot))), xlab = "year", ylab = "Pop'n size", lwd = 2)
lines(x = c(1:years), y = apply(sim1$orchard_pops_pre[[1]][[2]], MARGIN = 2, sum), col = "dodgerblue", lty = 2)
```



test no restoration


```{r}

years <- 20
n <- 5

# define size classes
SC1 <- 0
SC2 <- 10
SC3 <- 100
SC4 <- 900
SC5 <- 4000

A_mids <- c((SC1 + SC2)/2, (SC2 + SC3)/2, (SC3 + SC4)/2, (SC4 + SC5)/2, 9325) # midpoint areas of each size class (cm^2); last value is the 50% quantile of observations greater than SC5 in the standardized data from the literature

A_reef <- mean(c(4492.011, 10076.67, 8943.55)) # area of the reef

orchard_treatments <- c("orchard1") # orchards (could have multiple orchards and/or orchards with different post-outplanting treatments)
reef_treatments <- c("reef1") # reef treatments/subpopulations (could have "urchin outplanting" or "algal removal" or other postoutplanting treatments)
lab_treatments <- c("0_T1", "0_T2") # lab (cement = T1, ceramic = T2). "TX" is tile type, "X_" indicates whether recruits are outplanted immediately (0_) or the next year (0_1)


# demographic parameters for each orchard and reef treatment/subpopulation
field_surv1 <- field_surv
field_growth1 <- field_growth
nurs_surv1 <- nurs_surv
nurs_growth1 <- nurs_growth

# nursery data only available for size classes 1-2, so use the field data for the larger size classes
nurs_surv1$SC_surv_summ_df <- nurs_surv1$SC_surv_summ_df %>% filter(size_class < 3) %>% rbind(field_surv1$SC_surv_summ_df[field_surv1$SC_surv_summ_df$size_class > 2,])

nurs_growth1$summ_list[[3]] <- field_growth1$summ_list[[3]]
nurs_growth1$summ_list[[4]] <- field_growth1$summ_list[[4]]
nurs_growth1$summ_list[[5]] <- field_growth1$summ_list[[5]]

all_pars <- default_pars_fun(n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), summ_metric_list = list(field_surv = "mean", field_growth = "mean", field_shrink = "mean", field_frag = "mean", nurs_surv = "mean", nurs_growth = "mean", nurs_shrink = "mean"), field_surv1, field_growth1, nurs_surv1, nurs_growth1, apal_frag_summ)

# reef parameters: survival
surv_pars.r <- all_pars$surv_pars.r

# reef parameters: growth
growth_pars.r <- all_pars$growth_pars.r

# reef parameters: shrinkage
shrink_pars.r <- all_pars$shrink_pars.r

# reef parameters: fragmentation
frag_pars.r <- all_pars$frag_pars.r

# density dependent survival (smallest size class only for now)
dens_pars.r <- list()
dens_pars.r[[1]] <- list() # first reef treatment/subpop
dens_pars.r[[1]][[1]] <- 0
dens_pars.r[[1]][[2]] <- dens_pars.r[[1]][[1]]
dens_pars.r[[1]][[3]] <- dens_pars.r[[1]][[1]]

# reef parameters: fecundity (for tracking reef reproductive output)
fec_pars.r <- list()
fec_pars.r[[1]] <- list()
fec_pars.r[[1]][[1]] <- c(0, 0, 5000, 50000, 100000)
fec_pars.r[[1]][[2]] <- fec_pars.r[[1]][[1]]
fec_pars.r[[1]][[3]] <- fec_pars.r[[1]][[1]]

# orchard parameters: survival
surv_pars.o <- all_pars$surv_pars.o

# orchard parameters: growth
growth_pars.o <- all_pars$growth_pars.o

# orchard parameters: shrinkage
shrink_pars.o <- all_pars$shrink_pars.o

# orchard parameters: fragmentation
frag_pars.o <- all_pars$frag_pars.o

# density dependent survival (smallest size class only for now)
dens_pars.o <- list()
dens_pars.o[[1]] <- list() # first orchard
dens_pars.o[[1]][[1]] <- 0
dens_pars.o[[1]][[2]] <- dens_pars.o[[1]][[1]]

# orchard parameters: fecundity
fec_pars.o <- list()
fec_pars.o[[1]] <- list() # first treatment
fec_pars.o[[1]][[1]] <- c(0, 0, 5000, 50000, 100000) # first source
fec_pars.o[[1]][[2]] <- fec_pars.o[[1]][[1]] # second source

lambda <- 0 # external recruitment to reef

lambda_R <- 10000 # mean number of larvae produced by reference reefs each year

# stochasticity parameters
sigma_s <- 0
sigma_f <- 0
ext_rand <- c(FALSE, FALSE)
seeds <- c(1000, 5000, 10000, 40000)

# initial conditions in each reef subpopulation
N0.r <- list()
N0.r[[1]] <- list() # second reef subpop
N0.r[[1]][[1]] <- c(15, 25, 15, 10, 6)#rep(0, n) # first source (external recruits)
N0.r[[1]][[2]] <- rep(0, n) # second source
N0.r[[1]][[3]] <- rep(0, n) # third source

# initial conditions in each orchard subpopulation
N0.o <- list()
N0.o[[1]] <- list() # first orchard treatment
N0.o[[1]][[1]] <- rep(0, n) # first source
N0.o[[1]][[2]] <- rep(0, n) # second source

N0.l <- list()
N0.l[[1]] <- 0
N0.l[[2]] <- 0


# disturbance parameters
dist_yrs <- NA#c(years + 10) # years when disturbance occurs, can set to > years for none

# effects of disturbance on each reef subpop
dist_effects.r <- list()
dist_effects.r[[1]] <- list() # list where each element is the effects of disturbance on corals from each source in the 1st reef subpop (either "survival" for survival, "Tmat" for growth/survival, "Fmat" for fragmentation, or "fecundity" for reproduction)
dist_effects.r[[1]][[1]] <- list() # effects of disturbances on corals from first source in first reef subpop
dist_effects.r[[1]][[1]][[1]] <- c("survival") # effects of 1st disturbance on corals from first source in first reef subpop ([[1]][[1]][[2]] would be second disturbance, etc.)
dist_effects.r[[1]][[2]] <- list() # second source
dist_effects.r[[1]][[2]][[1]] <- c("survival")
dist_effects.r[[1]][[3]] <- list() # third source
dist_effects.r[[1]][[3]][[1]] <- c("survival")


# disturbance parameters for each reef subpop (what happens to the population parameters at each disturbance eve)
dist_pars.r <- list()

# first reef population
dist_pars.r[[1]] <- list() 
# first source to second reef subpop
dist_pars.r[[1]][[1]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[1]], dist_surv0 = list(surv_pars.r[[1]][[1]]*0.1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL) # list of disturbance parameters for the first source to the first reef population (defaults for each parameter type are NULL unless the disturbance affects them, as specified in dist_effects)
# second source to second reef subpop
dist_pars.r[[1]][[2]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[2]], dist_surv0 = list(surv_pars.r[[1]][[2]]*0.1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)
# third source to second reef subpop
dist_pars.r[[1]][[3]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[3]], dist_surv0 = list(surv_pars.r[[1]][[3]]*0.1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)

# disturbance effects for each orchard subpop
dist_effects.o <- list()

dist_effects.o[[1]] <- list() 
dist_effects.o[[1]][[1]] <- list() # effects of disturbances on corals from first source in first orchard treatment
dist_effects.o[[1]][[1]][[1]] <- c("survival") # effects of 1st disturbance on corals from first source in first reef subpop ([[1]][[1]][[2]] would be second disturbance, etc.)
dist_effects.o[[1]][[2]] <- list() # second source
dist_effects.o[[1]][[2]][[1]] <- c("survival")


# disturbance parameters for each orchard subpop
dist_pars.o <- list()

# first orchard treatment
dist_pars.o[[1]] <- list() 
# first source in first orchard treatment
dist_pars.o[[1]][[1]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.o[[1]][[1]], dist_surv0 = list(surv_pars.o[[1]][[1]]*0.1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL) 
# second source in first orchard treatment
dist_pars.o[[1]][[2]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.o[[1]][[2]], dist_surv0 = list(surv_pars.o[[1]][[2]]*0.1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)


# lab parameters
s0 <- c(0.9, 0.9)*0 # fraction of settled larvae in each lab treatment that survive to immediate outplanting
s1 <- c(0.7, 0.7)*0 # fraction of settled larvae in each lab treatment that survive to 1 yr outplanting
m0 <- c(0, 0) # density dependent mortality rate of larvae in each lab treatment that get outplanted immediately
m1 <- c(0, 0) # density dependent mortality rate of larvae in each lab treatment that get outplanted after 1 year
sett_props <- list(T1 = 0.95, T2 = 0.95) # proportion of larvae that settle on each tile type
size_props <- matrix(NA, nrow = length(lab_treatments), ncol = n) # matrix for fractions of retained recruits in each size class at the end of their year in the lab
size_props[1, ] <- c(1, 0, 0, 0, 0) # first lab treatment (0_T1)
size_props[2, ] <- c(1, 0, 0, 0, 0) # second lab treatment (0_T2)
lab_pars <- list(s0 = s0, s1 = s1, m0 = m0, m1 = m1, sett_props = sett_props, size_props = size_props)

# restoration strategy parameters
tile_props <- list(T1 = 0.5, T2 = 0.5) # proportion of tiles that are each type
orchard_yield <- 0 # percent of new orchard babies successfully collected
reef_yield <- 0.001 # percent of new reef babies successfully collected and fertilized
reef_prop <- c(0.5, 0.5) # proportion of lab recruits from each lab treatment outplanted to reef (1-proportion outplanted to orchard)
reef_out_props <- matrix(NA, nrow = length(lab_treatments), ncol = length(reef_treatments)) # proportion recruits going from each lab treatment to each reef treatment (row = origin lab treatment, column = destination reef treatment)
reef_out_props[1,] <- c(1) # from first lab treatment to each reef treatment (here there's just on intervention reef so they all go there)
reef_out_props[2,] <- c(1) # from second lab treatment to each reef treatment 

orchard_out_props <- matrix(NA, nrow = length(lab_treatments), ncol = length(orchard_treatments)) # proportion recruits going from each lab treatment to each orchard treatment
orchard_out_props[1,] <- c(1) # from first lab treatment to each orchard 
orchard_out_props[2,] <- c(1) # from second lab treatment to each orchard 

# sizes allocated to each treatment
reef_areas <- c(A_reef)*10000 # m^2 given to each post-outplanting reef treatment/subpopulation, convert to cm^2 by multiplying by 10,000
lab_max <- c(2500) # total number of settlers that the lab can accommodate
lab_retain_max <- c(1200) # total number of settlers that the lab can keep for a year
orchard_size <- c(200) # number of individuals each orchard treatment has space for

# coral transplanting
transplant <- rep(0, years) # years in which corals get transplanted from the orchard
#transplant[10] <- 1 # in 10th year, move corals from orchard to reef

null_mat <- matrix(0, nrow = years, ncol = n)
#rest_pars$trans_mats[[1]][[1]][which(transplant!=0)[1],] <- c(0, 0, 0, 2,0)# max number of colonies of each size class to move to reef in first transplant event
trans_mats <- list()
trans_mats[[1]] <- list() # first orchard 
trans_mats[[1]][[1]] <- null_mat # number of corals in each size class that originated from the first source of colonies for this orchard to transplant at each timepoint
trans_mats[[1]][[2]] <- null_mat # number of corals in each size class that originated from the second source of colonies for this orchard to transplant at each timepoint

# trans_mats[[2]] <- list() # second orchard
# trans_mats[[2]][[1]] <- null_mat # first source
# trans_mats[[2]][[2]] <- null_mat # second source

# where on the reef to put the transplants
#rest_pars$trans_reef[[1]][[1]][which(transplant!=0)[1],] <- c(1, 1)# reef and source subpop to transport the corals to
trans_reef <- list()
trans_reef[[1]] <- list() # first orchard
trans_reef[[1]][[1]] <- matrix(c(1, 2), nrow = years, ncol = 2, byrow = T) # first source corals in this orchard get transplanted to the second source of corals in the first reef
trans_reef[[1]][[2]] <- matrix(c(1, 3), nrow = years, ncol = 2, byrow = T) # second source corals in this orchard get transplanted to the third source of corals in the second reef

rest_pars <- list(tile_props = tile_props, orchard_yield = orchard_yield, reef_yield = reef_yield, reef_prop = reef_prop, reef_out_props = reef_out_props, orchard_out_props = orchard_out_props, reef_areas = reef_areas, lab_max = lab_max, lab_retain_max = lab_retain_max, orchard_size = orchard_size, transplant = transplant, trans_mats = trans_mats, trans_reef = trans_reef)


sim1 <- rse_mod1(years, n, A_mids, surv_pars.r, dens_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, surv_pars.o, dens_pars.o, growth_pars.o, shrink_pars.o, frag_pars.o, fec_pars.o, lambda, lambda_R, sigma_s, sigma_f, ext_rand, seeds, dist_yrs, dist_pars.r, dist_effects.r, dist_pars.o, dist_effects.o, orchard_treatments, reef_treatments, lab_treatments, lab_pars, rest_pars, N0.r, N0.o, N0.l)

# calculate lambda
mats <- mat_pars_fun(years, n, surv_pars.r[[1]][[1]], growth_pars.r[[1]][[1]], shrink_pars.r[[1]][[1]], frag_pars.r[[1]][[1]], fec_pars.r[[1]][[1]], sigma_s, sigma_f, seeds, dist_yrs, dist_pars.r[[1]][[1]],dist_effects.r[[1]][[1]])

Tmat.i <- mats$growth[[1]]
Fmat.i <- mats$fragmentation[[1]]


# calculate lambda (population growth rate)
pop_mat <-  (Tmat.i + Fmat.i) * matrix(surv_pars.r[[1]][[1]], nrow = n, ncol = n, byrow = T)
lambda_1 <- Re(eigen(pop_mat)$values[1]) # leading eigenvalue

lambda_1

# add confidence intervals
# highest survival, growth, and fragmentation; lowest shrinkage
all_pars <- default_pars_fun(n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), summ_metric_list = list(field_surv = "Q95", field_growth = "Q95", field_shrink = "Q05", field_frag = "Q95", nurs_surv = "Q95", nurs_growth = "Q95", nurs_shrink = "Q05"), field_surv1, field_growth1, nurs_surv1, nurs_growth1, apal_frag_summ)

# reef parameters
surv_pars.r <- all_pars$surv_pars.r
growth_pars.r <- all_pars$growth_pars.r
shrink_pars.r <- all_pars$shrink_pars.r
frag_pars.r <- all_pars$frag_pars.r

# orchard parameters
surv_pars.o <- all_pars$surv_pars.o
growth_pars.o <- all_pars$growth_pars.o
shrink_pars.o <- all_pars$shrink_pars.o
frag_pars.o <- all_pars$frag_pars.o


sim1_up <- rse_mod1(years, n, A_mids, surv_pars.r, dens_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, surv_pars.o, dens_pars.o, growth_pars.o, shrink_pars.o, frag_pars.o, fec_pars.o, lambda, lambda_R, sigma_s, sigma_f, ext_rand, seeds, dist_yrs, dist_pars.r, dist_effects.r, dist_pars.o, dist_effects.o, orchard_treatments, reef_treatments, lab_treatments, lab_pars, rest_pars, N0.r, N0.o, N0.l)

# calculate lambda
mats <- mat_pars_fun(years, n, surv_pars.r[[1]][[1]], growth_pars.r[[1]][[1]], shrink_pars.r[[1]][[1]], frag_pars.r[[1]][[1]], fec_pars.r[[1]][[1]], sigma_s, sigma_f, seeds, dist_yrs, dist_pars.r[[1]][[1]],dist_effects.r[[1]][[1]])

Tmat.i <- mats$growth[[1]]
Fmat.i <- mats$fragmentation[[1]]


# calculate lambda (population growth rate)
pop_mat <-  (Tmat.i + Fmat.i) * matrix(surv_pars.r[[1]][[1]], nrow = n, ncol = n, byrow = T)
lambda_1 <- Re(eigen(pop_mat)$values[1]) # leading eigenvalue

lambda_1


# lowest survival, growth, and fragmentation; highest shrinkage
all_pars <- default_pars_fun(n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), summ_metric_list = list(field_surv = "Q05", field_growth = "Q05", field_shrink = "Q95", field_frag = "Q05", nurs_surv = "Q05", nurs_growth = "Q05", nurs_shrink = "Q95"), field_surv1, field_growth1, nurs_surv1, nurs_growth1, apal_frag_summ)

# reef parameters
surv_pars.r <- all_pars$surv_pars.r
growth_pars.r <- all_pars$growth_pars.r
shrink_pars.r <- all_pars$shrink_pars.r
frag_pars.r <- all_pars$frag_pars.r

# orchard parameters
surv_pars.o <- all_pars$surv_pars.o
growth_pars.o <- all_pars$growth_pars.o
shrink_pars.o <- all_pars$shrink_pars.o
frag_pars.o <- all_pars$frag_pars.o


sim1_low <- rse_mod1(years, n, A_mids, surv_pars.r, dens_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, surv_pars.o, dens_pars.o, growth_pars.o, shrink_pars.o, frag_pars.o, fec_pars.o, lambda, lambda_R, sigma_s, sigma_f, ext_rand, seeds, dist_yrs, dist_pars.r, dist_effects.r, dist_pars.o, dist_effects.o, orchard_treatments, reef_treatments, lab_treatments, lab_pars, rest_pars, N0.r, N0.o, N0.l)

# calculate lambda
mats <- mat_pars_fun(years, n, surv_pars.r[[1]][[1]], growth_pars.r[[1]][[1]], shrink_pars.r[[1]][[1]], frag_pars.r[[1]][[1]], fec_pars.r[[1]][[1]], sigma_s, sigma_f, seeds, dist_yrs, dist_pars.r[[1]][[1]],dist_effects.r[[1]][[1]])

Tmat.i <- mats$growth[[1]]
Fmat.i <- mats$fragmentation[[1]]
#mats

# calculate lambda (population growth rate)
pop_mat <-  (Tmat.i + Fmat.i) * matrix(surv_pars.r[[1]][[1]], nrow = n, ncol = n, byrow = T)
lambda_1 <- Re(eigen(pop_mat)$values[1]) # leading eigenvalue

lambda_1

pop_lambda_fun(surv_pars.r[[1]][[1]], growth_pars.r[[1]][[1]], shrink_pars.r[[1]][[1]], frag_pars.r[[1]][[1]])



```


```{r}


# plot results

reef1_tot <- apply(sim1$reef_pops_pre[[1]][[1]], MARGIN = 2, sum) + apply(sim1$reef_pops_pre[[1]][[2]], MARGIN = 2, sum) + apply(sim1$reef_pops_pre[[1]][[3]], MARGIN = 2, sum)
orch1_tot <- apply(sim1$orchard_pops_pre[[1]][[1]], MARGIN = 2, sum) + apply(sim1$orchard_pops_pre[[1]][[2]], MARGIN = 2, sum)

reef1_up <- apply(sim1_up$reef_pops_pre[[1]][[1]], MARGIN = 2, sum) + apply(sim1_up$reef_pops_pre[[1]][[2]], MARGIN = 2, sum) + apply(sim1_up$reef_pops_pre[[1]][[3]], MARGIN = 2, sum)
orch1_tot <- apply(sim1_up$orchard_pops_pre[[1]][[1]], MARGIN = 2, sum) + apply(sim1_up$orchard_pops_pre[[1]][[2]], MARGIN = 2, sum)

reef1_low <- apply(sim1_low$reef_pops_pre[[1]][[1]], MARGIN = 2, sum) + apply(sim1_low$reef_pops_pre[[1]][[2]], MARGIN = 2, sum) + apply(sim1_low$reef_pops_pre[[1]][[3]], MARGIN = 2, sum)
orch1_tot <- apply(sim1_low$orchard_pops_pre[[1]][[1]], MARGIN = 2, sum) + apply(sim1_low$orchard_pops_pre[[1]][[2]], MARGIN = 2, sum)

plot(x = c(1:years), y = reef1_tot, type = "l", las = 1, ylim = c(0, max(max(reef1_up[1:10]), max(orch1_tot))), xlab = "year", ylab = "Pop'n size", lwd = 2, xlim = c(1, 10))
polygon(x = c(c(1:years), rev(c(1:years))), y = c(reef1_up, rev(reef1_low)), col = adjustcolor("gray", alpha.f = 0.5), border = NA)
lines(x = c(1:years), y = reef1_tot)
#lines(x = c(1:years), y = orch1_tot, type = "l", col = "dodgerblue", lwd = 2, lty = 2)
#legend("topleft", legend = c("reef", "orchard"), col = c("black", "dodgerblue"), lwd = 2, lty = c(1, 2))


```

```{r}
# with summary function

reef1_tot <- model_summ(model_sim = sim1, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

#orch1_tot <- model_summ(model_sim = sim1, location = "orchard", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

reef1_up <- model_summ(model_sim = sim1_up, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

reef1_low <- model_summ(model_sim = sim1_low, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))


plot(x = c(1:years), y = reef1_tot, type = "l", las = 1, ylim = c(0, max(max(reef1_up[1:10]), max(orch1_tot))), xlab = "year", ylab = "Pop'n size", lwd = 2, xlim = c(1, 10))
polygon(x = c(c(1:years), rev(c(1:years))), y = c(reef1_up, rev(reef1_low)), col = adjustcolor("gray", alpha.f = 0.5), border = NA)
lines(x = c(1:years), y = reef1_tot)
#lines(x = c(1:years), y = orch1_tot, type = "l", col = "dodgerblue", lwd = 2, lty = 2)
#legend("topleft", legend = c("reef", "orchard"), col = c("black", "dodgerblue"), lwd = 2, lty = c(1, 2))

reef1_tot <- model_summ(model_sim = sim1, location = "reef", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

reef1_up <- model_summ(model_sim = sim1_up, location = "reef", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

reef1_low <- model_summ(model_sim = sim1_low, location = "reef", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

#sum(sim1$reef_pops_pre[[1]][[1]][,1]*A_mids)/10000 


plot(x = c(1:years), y = reef1_tot, type = "l", las = 1, ylim = c(0, max(max(reef1_up[1:10]), max(orch1_tot))), xlab = "year", ylab = "Area covered", lwd = 2, xlim = c(1, 10))
polygon(x = c(c(1:years), rev(c(1:years))), y = c(reef1_up, rev(reef1_low)), col = adjustcolor("gray", alpha.f = 0.5), border = NA)
lines(x = c(1:years), y = reef1_tot)


reef1_tot <- model_summ(model_sim = sim1, location = "reef", metric = "production", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

reef1_up <- model_summ(model_sim = sim1_up, location = "reef", metric = "production", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

reef1_low <- model_summ(model_sim = sim1_low, location = "reef", metric = "production", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

#sum(sim1$reef_pops_pre[[1]][[1]][,1]*A_mids)/10000 


plot(x = c(1:years), y = reef1_tot, type = "l", las = 1, ylim = c(0, max(max(reef1_up[1:10]), max(orch1_tot))), xlab = "year", ylab = "Reproduction", lwd = 2, xlim = c(1, 10))
polygon(x = c(c(1:years), rev(c(1:years))), y = c(reef1_up, rev(reef1_low)), col = adjustcolor("gray", alpha.f = 0.5), border = NA)
lines(x = c(1:years), y = reef1_tot)

```



random iterations: first make a function for summarizing model results (total population size and total area covered) and calculating lambda (then show histogram of lambda values)


```{r}

# number of iterations
n_sample1 <- 100

# holding vector for population lambdas
lambda_set <- rep(NA, n_sample1)

# holding matrix for population sizes
Nmat <- matrix(NA, nrow = years, ncol = n_sample1)

# get the parameter sets
field_surv1 <- field_surv
field_growth1 <- field_growth
nurs_surv1 <- nurs_surv
nurs_growth1 <- nurs_growth

# nursery data only available for size classes 1-2, so use the field data for larger size classes
nurs_surv1$SC_surv_df <- nurs_surv1$SC_surv_df %>% filter(size_class < 3) %>% rbind(field_surv1$SC_surv_df[field_surv1$SC_surv_df$size_class > 2,])

nurs_growth1$mat_list[[3]] <- field_growth1$mat_list[[3]]
nurs_growth1$mat_list[[4]] <- field_growth1$mat_list[[4]]
nurs_growth1$mat_list[[5]] <- field_growth1$mat_list[[5]]


set.seed(500)
all_pars <- rand_pars_fun(n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), n_sample = n_sample1, field_surv = field_surv1, field_growth = field_growth1, nurs_surv = nurs_surv1, nurs_growth = nurs_growth1, apal_frag)


for(i in 1:n_sample1){
  
  
  sim_i <- rse_mod1(years, n, A_mids, surv_pars.r = all_pars$surv_pars_L.r[[i]], dens_pars.r, growth_pars.r = all_pars$growth_pars_L.r[[i]], shrink_pars.r = all_pars$shrink_pars_L.r[[i]], frag_pars.r = all_pars$frag_pars_L.r[[i]], fec_pars.r, surv_pars.o = all_pars$surv_pars_L.o[[i]], dens_pars.o, growth_pars.o = all_pars$growth_pars_L.o[[i]], shrink_pars.o = all_pars$shrink_pars_L.o[[i]], frag_pars.o = all_pars$frag_pars_L.o[[i]], fec_pars.o, lambda, lambda_R, sigma_s, sigma_f, ext_rand, seeds, dist_yrs, dist_pars.r, dist_effects.r, dist_pars.o, dist_effects.o, orchard_treatments, reef_treatments, lab_treatments, lab_pars, rest_pars, N0.r, N0.o, N0.l)
  
  Nmat[,i] <- model_summ(model_sim = sim_i, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

  lambda_set[i] <- pop_lambda_fun(all_pars$surv_pars_L.r[[i]][[1]][[1]], all_pars$growth_pars_L.r[[i]][[1]][[1]], all_pars$shrink_pars_L.r[[i]][[1]][[1]], all_pars$frag_pars_L.r[[i]][[1]][[1]])
  
}


```


```{r}

par(mfrow = c(2, 1))
layout(matrix(c(1, 2), nrow = 1, ncol = 2, byrow = T), widths = rep(1, 2), heights = rep(1, 1))
#layout.show(2)
par(mar=c(0, 3, 3, 0.1), oma = c(4.5, 1, 2, 0.1))


plot(x = c(1:years), y = Nmat[,1], type = "l", col = adjustcolor("black", alpha.f = 0.25), las = 1, ylim = c(0, max(Nmat)), xlab = NA, ylab = NA)
mtext(side = 1, "Year", line = 2)
mtext(side = 2, "Population size", line = 2.25)
for(i in 2:n_sample1){
  lines(x = c(1:years), y = Nmat[,i], col = adjustcolor("black", alpha.f = 0.25))
}

hist(lambda_set, main = NA, xlab = NA, ylab = NA, las = 1)
mtext(side = 1, "Lambda", line = 2)
mtext(side = 2, "Frequency", line = 2)

```

try using the disturbance function to allow parameters to change each year

```{r}
dist_yrs <- c(1:years) # years when disturbance occurs, can set to > years or NA for none

# effects of disturbance on each reef subpop
dist_effects.r <- list()
dist_effects.r[[1]] <- list() # list where each element is the effects of disturbance on corals from each source in the 1st reef subpop (either "survival" for survival, "Tmat" for growth/survival, "Fmat" for fragmentation, or "fecundity" for reproduction)
dist_effects.r[[1]][[1]] <- list() # effects of disturbances on corals from first source in first reef subpop
dist_effects.r[[1]][[2]] <- list() # first reef, second source
dist_effects.r[[1]][[3]] <- list() # first reef, third source

for(i in 1:length(dist_yrs)){
  dist_effects.r[[1]][[1]][[i]] <- c("survival", "Tmat", "Fmat") # effects of ith disturbance on corals from first source in first reef subpop 
  dist_effects.r[[1]][[2]][[i]] <- c("survival", "Tmat", "Fmat")
  dist_effects.r[[1]][[3]][[i]] <- c("survival", "Tmat", "Fmat")
}


set.seed(1)
all_pars <- rand_pars_fun(n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), n_sample = length(dist_yrs), field_surv = field_surv1, field_growth = field_growth1, nurs_surv = nurs_surv1, nurs_growth = nurs_growth1, apal_frag)

# need to rearrange the list structures
surv_pars.r1 <- list() # first source (external recruits)
surv_pars.r2 <- list() # second source
surv_pars.r3 <- list() # third source

growth_pars.r1 <- list()
growth_pars.r2 <- list()
growth_pars.r3 <- list()

frag_pars.r1 <- list()
frag_pars.r2 <- list()
frag_pars.r3 <- list()

for(i in 1:length(dist_yrs)){
  surv_pars.r1[[i]] <- all_pars$surv_pars_L.r[[i]][[1]][[1]]
  surv_pars.r2[[i]] <- all_pars$surv_pars_L.r[[i]][[1]][[2]]
  surv_pars.r3[[i]] <- all_pars$surv_pars_L.r[[i]][[1]][[3]]
  
  mats1 <- G_fun(years = 1, n = 5, growth_pars = all_pars$growth_pars_L.r[[i]][[1]][[1]], shrink_pars = all_pars$shrink_pars_L.r[[i]][[1]][[1]], frag_pars = all_pars$frag_pars_L.r[[i]][[1]][[1]])
  
  mats2 <- G_fun(years = 1, n = 5, growth_pars = all_pars$growth_pars_L.r[[i]][[1]][[2]], shrink_pars = all_pars$shrink_pars_L.r[[i]][[1]][[2]], frag_pars = all_pars$frag_pars_L.r[[i]][[1]][[2]])
  
  mats3 <- G_fun(years = 1, n = 5, growth_pars = all_pars$growth_pars_L.r[[i]][[1]][[3]], shrink_pars = all_pars$shrink_pars_L.r[[i]][[1]][[3]], frag_pars = all_pars$frag_pars_L.r[[i]][[1]][[3]])
  
  growth_pars.r1[[i]] <- mats1$G_list[[1]]
  growth_pars.r2[[i]] <- mats2$G_list[[1]]
  growth_pars.r3[[i]] <- mats3$G_list[[1]]
  
  frag_pars.r1[[i]] <- mats1$Fr_list[[1]]
  frag_pars.r2[[i]] <- mats2$Fr_list[[1]]
  frag_pars.r3[[i]] <- mats3$Fr_list[[1]]
}


# disturbance parameters for each reef subpop (what happens to the population parameters at each disturbance eve)
dist_pars.r <- list()

# first reef population
dist_pars.r[[1]] <- list() 
# first source to second reef subpop
dist_pars.r[[1]][[1]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[1]], dist_surv0 = surv_pars.r1, dist_Tmat0 = growth_pars.r1, dist_Fmat0 = frag_pars.r1, dist_fec0 = NULL)# list of disturbance parameters for the first source to the first reef population (defaults for each parameter type are NULL unless the disturbance affects them, as specified in dist_effects)
# dist_surv0 = list where ith element is the survival probabilities for the ith disturbance
# second source to second reef subpop
dist_pars.r[[1]][[2]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[2]], dist_surv0 = surv_pars.r2, dist_Tmat0 = growth_pars.r2, dist_Fmat0 = frag_pars.r2, dist_fec0 = NULL)
# third source to second reef subpop
dist_pars.r[[1]][[3]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[3]], dist_surv0 = surv_pars.r3, dist_Tmat0 = growth_pars.r3, dist_Fmat0 = frag_pars.r3, dist_fec0 = NULL)


# disturbance parameters for each orchard 

dist_effects.o <- list()

dist_effects.o[[1]] <- list() 
dist_effects.o[[1]][[1]] <- list() # effects of disturbances on corals from first source in first orchard treatment
dist_effects.o[[1]][[2]] <- list() # second source


for(i in 1:length(dist_yrs)){
  dist_effects.o[[1]][[1]][[i]] <- c("none") # effects of ith disturbance on corals from first source in first orchard
  dist_effects.o[[1]][[2]][[i]] <- c("none")
}

dist_pars.o <- list()

# first orchard treatment
dist_pars.o[[1]] <- list() 
# first source in first orchard treatment
dist_pars.o[[1]][[1]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.o[[1]][[1]], dist_surv0 = NULL, dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL) 
# second source in first orchard treatment
dist_pars.o[[1]][[2]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.o[[1]][[2]], dist_surv0 = NULL, dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)



sim1 <- rse_mod1(years, n, A_mids, surv_pars.r, dens_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, surv_pars.o, dens_pars.o, growth_pars.o, shrink_pars.o, frag_pars.o, fec_pars.o, lambda, lambda_R, sigma_s, sigma_f, ext_rand, seeds, dist_yrs, dist_pars.r, dist_effects.r, dist_pars.o, dist_effects.o, orchard_treatments, reef_treatments, lab_treatments, lab_pars, rest_pars, N0.r, N0.o, N0.l)

```


```{r}
reef1_tot <- model_summ(model_sim = sim1, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

#orch1_tot <- model_summ(model_sim = sim1, location = "orchard", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

plot(x = c(1:years), y = reef1_tot, type = "l", las = 1, ylim = c(0, max(max(reef1_tot), max(orch1_tot))), xlab = "year", ylab = "Pop'n size", lwd = 2)
#lines(x = c(1:years), y = orch1_tot, type = "l", col = "dodgerblue", lwd = 2, lty = 2)
#legend("topleft", legend = c("reef", "orchard"), col = c("black", "dodgerblue"), lwd = 2, lty = c(1, 2))

```

