---
title: "rest_pars"
output: html_document
date: "2026-01-30"
---

README: estimating restoration parameters from Fundemar's data

```{r setup, include=FALSE}

# command + option + i for new code chunk
# command + shift + enter to run entire code chunk
# command + enter to run single line of code

library(tidyverse) # command + shift + m for %>%
library(readxl)
library(janitor)

```


from the original "Adicion_Sustratos" matrix (which was just for DLAB in 2025), it looks like they put ~100 tiles in each tank (divided among substrate types) and had 40 tanks

Goal is to know 1) how many embryos they put in each tank, 2) how many tiles per tank, and 3) settlement rates on the tiles. Want to know 1) and 2) to estimate how many tiles they would use for a given number of embryos, and need to know 3) for model parameters


Adicion_Sustratos_Unificada = spawning event, number of embryos, Tank ID, substrate type, and number of substrates

Matriz_Asentamiento_Unificada = number of settlers on the substrates (Tank ID should match A_S_U)

Matriz_Supervivencia_Reclutas_Unificada = recruit survival, but comparing time at T0 to number of settlers in  M_A_U matrix should give survival in the lab?)


```{r}
# read in the data

add_subs <- read_excel("data/Fundemar/Adicion_Sustratos_Unificada.xlsx", sheet = 1) %>% clean_names()

add_subs_APAL <- add_subs %>% filter(species == "APAL")

#View(add_subs_APAL)

sett_mat <- read_excel("data/Fundemar/Matriz_Asentamiento_Unificada.xlsx", sheet = 1) %>% clean_names()

sett_mat_APAL <- sett_mat %>% filter(species == "APAL")

#View(sett_mat_APAL)

surv_mat <- read_excel("data/Fundemar/Matriz_Supervivencia_Reclutas_Unificada.xlsx", sheet = 1) %>% clean_names()

#unique(surv_mat$coral_species)

surv_mat_APAL <- surv_mat %>% filter(coral_species == "Acropora palmata")

# View(surv_mat_APAL) # doesn't have 2025 yet

```


```{r}
# estimating embryos per tank and number of tiles per tank

#View(add_subs_APAL)

#add_subs_APAL %>% group_by(species, spawning_event, spawning_day, spawning_month, spawning_year, year_of_substrate_addition, month_of_substrate_addition, day_of_substrate_addition, tank_id) %>% summarize(n_substrates = sum(amount_of_substrates)) %>% ungroup() %>% arrange(tank_id) %>% view()


#add_subs_APAL %>% group_by(species, spawning_event, spawning_day, tank_id) %>% summarize(n_substrates = sum(amount_of_substrates), embryos_ml_mn = min(embryos_ml), embryos_ml_mx = min(embryos_ml)) %>% ungroup() %>% arrange(spawning_day) %>% mutate(emb_diff = embryos_ml_mx - embryos_ml_mn) %>% view()


tank_counts <- add_subs_APAL %>% group_by(species, spawning_event, spawning_day, tank_id) %>% summarize(n_substrates = sum(amount_of_substrates), embryos_ml = min(embryos_ml), n_embryos = min(no_embryos), vol_embryos_ml = min(volume_of_embryos_ml)) %>% ungroup() %>% arrange(spawning_day)# %>% view()

#View(tank_counts)

mean(tank_counts$embryos_ml) # mean volume put in the tanks = 7293
min(tank_counts$embryos_ml) # 7250
max(tank_counts$embryos_ml) # 7330

mean(tank_counts$vol_embryos_ml) # mean embryos per mL = 1.9
min(tank_counts$vol_embryos_ml) # min = 1.5
max(tank_counts$vol_embryos_ml) # max = 2.07

mean(tank_counts$n_substrates)

#unique(add_subs$embryos_ml[which(add_subs$species %in% c("APAL", "DLAB"))])

#hist(add_subs$embryos_ml)

```

So on average, they are putting about 7,300 mLs in a tank, and on average there are 2 embryos per mL, and they have about 100 substrates per tank

-- so to calculate the number of substrates to use, divide the total mLs collected by 7,300*2 ( = 14600) to get the number of tanks, then multiply that by 100 to get the number of substrates needed (point of this is to avoid scenarios where very little spawn was collected but all model still uses max capacity of tiles, which would mean each tile might have unrealistically low numbers of recruits)

```{r}
# now estimate settlement rates

#unique(add_subs_APAL$spawning_year) # all 2025

#View(sett_mat_APAL[which(sett_mat_APAL$id_rearing_container == "11A"),])

#colnames(sett_mat_APAL)

#sett_mat_APAL_summ <- sett_mat_APAL %>% filter(year_scored == 2025) %>% group_by(month_scored, monitoring_or_random_count, id_rearing_container, substrate_type) %>%  summarize(max_settlers = max(as.numeric(number_of_total_settlers)), n_substrates_sett = n()) #%>% ungroup() %>% group_by(month_scored, id_rearing_container) %>%  summarize(n_substrates_sett = n(), n_settlers = sum(as.numeric(max_settlers))) %>% ungroup() %>% rename(tank_id = id_rearing_container)
  
#sett_mat_APAL_summ <- sett_mat_APAL %>% filter(year_scored == 2025) %>% group_by(month_scored, day_scored, id_rearing_container) %>%  summarize(n_substrates_sett = n(), n_settlers = sum(as.numeric(number_of_total_settlers))) %>% ungroup() %>% rename(tank_id = id_rearing_container)


sett_mat_APAL_summ <- sett_mat_APAL %>% filter(year_scored == 2025) %>% group_by(month_scored, day_scored, id_rearing_container) %>%  summarize(n_substrates_sett = n(), n_settlers = mean(as.numeric(number_of_total_settlers))) %>% ungroup() %>% rename(tank_id = id_rearing_container)


#View(sett_mat_APAL_summ)

#mean(sett_mat_APAL_summ$n_settlers)

sett_rates_APAL <- tank_counts %>% left_join(sett_mat_APAL_summ, by = "tank_id")

#View(sett_rates_APAL)


# day scored = day when counting took place; just use the day with the max number observed for that tile since interested in settlement and not post-settlement mortality? 
# QUESTION: are they removing the substrates after they have been counted, so there is no double counting on different days scored? Or are they recounting substrates that were previously counted to see if there were any new settlers?

# since we don't know, just look at average number of settlers per tile? But also don't know how many substrates there were that had settlers on them...

# for now, assume there is no recounting

# really only care about total number of settlers in the tank? Or number of settlers in the tank and number of substrates with at least one settler?

# maybe they only count a subset of the substrates in a tank?


# in the report they say that the goal is ~30 recruits per substrate

mean(as.numeric(sett_mat_APAL$number_of_total_settlers[which(sett_mat_APAL$year_scored==2025 & sett_mat_APAL$substrate_type=="Cookie cement")]))
mean(as.numeric(sett_mat_APAL$number_of_total_settlers[which(sett_mat_APAL$year_scored==2025 & sett_mat_APAL$substrate_type=="Gear")]))

mean(as.numeric(sett_mat_APAL$number_of_total_settlers[which(sett_mat_APAL$year_scored==2025)]))

median(as.numeric(sett_mat_APAL$number_of_total_settlers[which(sett_mat_APAL$year_scored==2025)]))

hist(as.numeric(sett_mat_APAL$number_of_total_settlers[which(sett_mat_APAL$year_scored==2025)]))

#mean(sett_rates_APAL$n_settlers/sett_rates_APAL$n_substrates_sett)

# have about 20 on average

# multiply average number of settlers per tile (mean = ~20) by the number of tiles per tank (~100) to get total number of settlers in the tank, and divide the total number of embryos in the tank by the number of settlers in the tank to get an estimate of settlement rate

sett_rates_APAL <- sett_rates_APAL %>% mutate(sett_rate = n_settlers*n_substrates/n_embryos)

mean(sett_rates_APAL$sett_rate) # 0.1467558


```

so the average settlement rate is about 15% (might be higher; this is using number of embryos as the initial number when in reality not all of those embryos become larvae)


```{r}



# try replicating the box plot in fundemar's report

# %>% group_by(id_rearing_container, substrate_type, day_scored, month_scored) %>% mutate(number_of_total_settlers = mean(as.numeric(number_of_total_settlers))) %>% ungroup()

sett_mat_APAL %>% filter(year_scored == 2025) %>% 
ggplot(aes(x=as.factor(substrate_type), y=as.numeric(number_of_total_settlers))) + 
    geom_boxplot() +
   #facet_wrap(~substrate_type, scales = "free")+
  labs(title = "Larval settlement",
       y = "Average settlement", x = "Substrate") +
  theme(legend.position = "none")

# pretty close but not an exact match with Fig. 2



```

check ratio of cement to ceramic (gear) substrates

```{r}

#View(sett_mat_APAL[which(sett_mat_APAL$year_scored == 2025),])

length(which(sett_mat_APAL$year_scored == 2025 & sett_mat_APAL$substrate_type=="Cookie cement"))

length(which(sett_mat_APAL$year_scored == 2025 & sett_mat_APAL$substrate_type=="Gear"))

250/(250 + 244)

244/(250 + 244)

```




Density dependence

estimating density dependent mortality rates from Fundemar's outplanting data


```{r, warning=FALSE}
# settler survival
fundR_dt0 <- read_excel("data/Fundemar/Matriz_Supervivencia_Reclutas_Unificada.xlsx", sheet = 1)
#View(fundR_dt0)

fundR_dt0 <- fundR_dt0 %>% clean_names()


fundR_dt <- fundR_dt0 %>% filter(coral_species=="Acropora palmata") %>% filter(number_settlers != "NA") %>% mutate(number_settlers = as.numeric(number_settlers))# %>% filter(is.na(number_settlers)==F)

#View(fundR_dt0)

#View(fundR_dt)

#unique(fundR_dt$outplanting_site)
#fundR_dt %>% filter(outplanting_site == "Mesa de sustrato 2") %>% view() # looks like there's no data on initial outplant numbers at this site

# make data frame with the initial numbers of settlers
fundR_N0 <- fundR_dt %>% filter(months_after_outplanting==0) %>% mutate(N0 = number_settlers) %>% select(-"number_settlers", -"survey_date_year", -"survey_date_month", -"survey_date_day", -"survey_date_day", -"months_after_outplanting", -"remarks", -"time_point")

# now join back together with the main dataset
fundR_dt1 <- fundR_dt %>% left_join(fundR_N0, by = c("coral_species", "outplanting_site", "settlement_technique", "settlement_container_id", "transect", "outplant_date_year", "outplant_date_month", "outplant_date_day", "treatment_1", "treatment_2", "plot_tag_number", "substrate_number_temporary_flag"), relationship="many-to-many") %>% mutate(survival = number_settlers/N0)

#View(fundR_dt1)

# check this worked
#fundR_dt1 %>% filter(plot_tag_number==8, substrate_number_temporary_flag=="1") %>% view()


# now flag cases where the number of settlers increases over time (probably due to natural recruits or miscounting?)
fundR_dt2 <- fundR_dt1 %>% mutate(outplant_date = as.Date(paste(outplant_date_month, outplant_date_day, outplant_date_year, sep = "/"), format = '%m/%d/%Y')) %>% group_by(settlement_technique, transect, treatment_2, outplanting_site, outplant_date, treatment_1, plot_tag_number, substrate_number_temporary_flag) %>% arrange(months_after_outplanting) %>% mutate(survival1 = dplyr::lag(survival, n = 1L)) %>% ungroup() %>% mutate(new_recruits = if_else(survival - survival1 > 0, "Y", "N")) %>% group_by(settlement_technique, transect, treatment_2, outplanting_site, outplant_date, treatment_1, plot_tag_number, substrate_number_temporary_flag) %>% mutate(new_recruits_tot = if_else("Y" %in% new_recruits, "Y", "N")) %>% ungroup() # negative values of survival - survival1 means less recruits at current timepoint, positive values means more recruits (i.e., somehow a new recruit appeared). new_recruits_tot indicates whether a new recruit ever appeared on that tile
fundR_dt2$new_recruits <- ifelse(is.na(fundR_dt2$new_recruits)==T, "N", fundR_dt2$new_recruits)

# check this worked
#fundR_dt2 %>% filter(plot_tag_number==8, substrate_number_temporary_flag=="1") %>% view()
#fundR_dt2 %>% filter(plot_tag_number==2, substrate_number_temporary_flag=="60") %>% view()
#fundR_dt2 %>% filter(plot_tag_number==5, substrate_number_temporary_flag=="16") %>% view()

# select substrates that never had an increase in number of recruits and look at survival one year after outplanting

#noaa_dt$Date <- as.Date(noaa_dt$Date, format = '%m/%d/%Y')

fundR_dt3.2 <- fundR_dt2 %>% mutate(outplant_date = as.Date(paste(outplant_date_month, outplant_date_day, outplant_date_year, sep = "/"), format = '%m/%d/%Y'), survey_date = as.Date(paste(survey_date_month, survey_date_day, survey_date_year, sep = "/"), format = '%m/%d/%Y')) %>% filter(new_recruits_tot == "N") %>% select(outplanting_site, months_after_outplanting, settlement_technique, settlement_container_id, transect, outplant_date, survey_date, survey_date_year, treatment_1, treatment_2, plot_tag_number, substrate_number_temporary_flag, number_settlers, N0)
```


```{r}
#View(fundR_dt3)

#hist(fundR_dt2$months_after_outplanting)

#unique(fundR_dt3.2$months_after_outplanting)

#fundR_dt3.2 %>% group_by(months_after_outplanting, outplant_date) %>% summarize(n = n()) %>% view()

# just use data from one month after outplanting (less time for other factors to influence mortality); all are from the same outplant date
fundDD <- fundR_dt3.2 %>% mutate(prop_surv = number_settlers/N0) %>%filter(months_after_outplanting == 1) %>% mutate(n_dead = N0-number_settlers) #%>% mutate(prop_surv_yr = prop_surv^12) #filter(months_after_outplanting > 0 & months_after_outplanting < 2)

# fundDD %>% group_by(treatment_1) %>% summarize(n = n()) %>% view() # almost all are cookie or tetrapod cement


# fit nls model?
# dd_mod <- nls(prop_surv ~ exp(m*N0), data = fundDD, start = list(m = -0.1))
# 
# summary(dd_mod)
# 

# plot the raw data
plot(x = fundDD$N0, y = fundDD$prop_surv, las = 1, pch = 16, col = adjustcolor("black", alpha.f = 0.5))
# lines(x = seq(from = 0, to = 80, length.out = 1000), y = exp(-0.97*seq(from = 0, to = 80, length.out = 1000)), type = "l")


# should first convert prop surviving over one month to proportion surviving after 1 yr?

# plot with y on the log scale and fit a line to this? but have issue of the 0's
# plot(x = fundDD$N0, y = log(fundDD$prop_surv + 1), las = 1, pch = 16, col = adjustcolor("black", alpha.f = 0.5))

```

tile sizes (from "FW: Data for RSE" email sent on 9/15/25): "for Tetrapod cement we use 130 cm2, the cookies are not standardized", email also included some diagrams for gear


problem with using proportions: for low densities, it's easier to get 0% or 100% mortality because there aren't very many individuals...


```{r}

# try making it binary?

# fundDD

# plot(x = fundDD$N0, y = rep(0, length(fundDD$N0)), ylim = c(0, 1), col = NA, las = 1)
# for(i in 1:nrow(fundDD)){
#   points(x = rep(fundDD$N0[i], fundDD$number_settlers[i]), y = rep(1, fundDD$number_settlers[i]), pch = 16, col = adjustcolor("black", alpha.f = 0.15))
#   points(x = rep(fundDD$N0[i], fundDD$n_dead[i]), y = rep(0, fundDD$n_dead[i]), pch = 16, col = adjustcolor("black", alpha.f = 0.15))
# }


```


Solution: group into density bins (e.g., 0-10, 10-20, etc.) and calculate prop surviving in each bin, then fit model to the values from each bin. Check what bin size gives reasonable numbers per bin in the lower sizes


```{r}

# make data frame with 1's and 0's
# can use a for loop: for each row, make a data frame with number of rows = N0, then fill in with 1's and 0's, then join them all together

#dd_bounds <- c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90) 

dd_bounds <- c(0, 15, 30, 45, 60, 75, 90) 

# means in each of these regions
dd_means <- rep(NA, length(dd_bounds)-1)

for(j in 2:length(dd_bounds)){
  dd_means[j-1] <- mean(fundDD$N0[which(fundDD$N0 >= dd_bounds[j-1] & fundDD$N0 < dd_bounds[j])])
}

for(i in 1:nrow(fundDD)){
  
  dt_sub <- fundDD[i,]
  
  #dt_sub$N_class <- ifelse(dt_sub$N0 >= 0 & dt_sub$N0 < 10, dd_means[1], ifelse(dt_sub$N0 >= 10 & dt_sub$N0 < 20, dd_means[2], ifelse(dt_sub$N0 >= 20 & dt_sub$N0 < 30, dd_means[3], ifelse(dt_sub$N0 >= 30 & dt_sub$N0 < 40, dd_means[4], ifelse(dt_sub$N0 >= 40 & dt_sub$N0 < 50, dd_means[5], ifelse(dt_sub$N0 >= 50 & dt_sub$N0 < 60, dd_means[6], ifelse(dt_sub$N0 >= 60 & dt_sub$N0 < 70, dd_means[7], ifelse(dt_sub$N0 >= 70 & dt_sub$N0 < 80, dd_means[8], ifelse(dt_sub$N0 >= 80, dd_means[9], NA)))))))))
  
  dt_sub$N_class <- ifelse(dt_sub$N0 >= 0 & dt_sub$N0 < 15, dd_means[1], ifelse(dt_sub$N0 >= 15 & dt_sub$N0 < 30, dd_means[2], ifelse(dt_sub$N0 >= 30 & dt_sub$N0 < 45, dd_means[3], ifelse(dt_sub$N0 >= 45 & dt_sub$N0 < 60, dd_means[4], ifelse(dt_sub$N0 >= 60 & dt_sub$N0 < 75, dd_means[5], ifelse(dt_sub$N0 >= 75 & dt_sub$N0 < 90, dd_means[6], NA))))))
  
  fundDD_ind_i <- data.frame(
    N0 = rep(dt_sub$N0, dt_sub$N0),
    survived = c(rep(0, dt_sub$n_dead), rep(1, dt_sub$number_settlers)),
    N_class = dt_sub$N_class
  )
  
  if(i == 1){
    fundDD_ind <- fundDD_ind_i
  } else{
    
    fundDD_ind <- rbind(fundDD_ind, fundDD_ind_i)
  }
  
}

#View(fundDD_ind)

#sort(unique(fundDD_ind$N0))
#View(fundDD_ind)

fundDD_summ <- fundDD_ind %>% group_by(N_class) %>% summarize(N_tot = n(), N_survived = sum(survived[survived == 1])) %>% mutate(prop_survived = N_survived/N_tot) %>% mutate(prop_survived_yr = prop_survived^12)# %>% view()

#View(fundDD_summ)
  
```


```{r}

#dd_mod3 <- nls(prop_survived ~ exp(m*N_class), data = fundDD_summ, start = list(m = -0.1))

#summary(dd_mod3)

#dd_mod3 <- lm(log(prop_survived + 1) ~ N_class, data = fundDD_summ)


# dd_mod3 <- lm(log(prop_survived_yr) ~ N_class, data = fundDD_summ)
# 
# 
# summary(dd_mod3)
# 
# plot(x = fundDD_summ$N_class, y = log(fundDD_summ$prop_survived_yr), las = 1, pch = 16)
# abline(a = dd_mod3$coefficients[1], b = dd_mod3$coefficients[2], col = "blue")

fundDD_summ

dd_mod3 <- lm(log(prop_survived) ~ N_class, data = fundDD_summ)

plot(x = fundDD_summ$N_class, y = log(fundDD_summ$prop_survived), las = 1, pch = 16)
abline(a = dd_mod3$coefficients[1], b = dd_mod3$coefficients[2], col = "blue")

summary(dd_mod3)

```

slope is approx. -0.026

y = exp(mx)*exp(b)

log(y) = mx + b


```{r}

# plot(x = fundDD_summ$N_class, y = fundDD_summ$prop_survived_yr, las = 1, pch = 16)
# lines(x = seq(from = 0, to = 80, length.out = 1000), y = exp(dd_mod3$coefficients[1])*exp(dd_mod3$coefficients[2]*seq(from = 0, to = 80, length.out = 1000)), type = "l", col = "blue")

plot(x = fundDD_summ$N_class, y = fundDD_summ$prop_survived, las = 1, pch = 16, xlab = "Density class", ylab = "Proportion surviving", ylim = c(0, 0.18), xlim = c(3, 85))
lines(x = seq(from = 0, to = 80, length.out = 1000), y = exp(dd_mod3$coefficients[1])*exp(dd_mod3$coefficients[2]*seq(from = 0, to = 80, length.out = 1000)), type = "l", col = "blue")
text(x = fundDD_summ$N_class + 2, y = fundDD_summ$prop_survived + 0.01, paste("n = ", as.character(fundDD_summ$N_tot), sep = ""))
text(x = 70, y = 0.15, "y = exp(-0.026x)exp(-1.86)", col = "blue")

```




density dependence is formulated as

```{r}
# S_i[1] <- S_i[1]*exp(-m*sum(N_all)); want to estimate m
```

 

Fecundity:

```{r}

# without additional information (Fundemar's report only): c(0, 0, rep(1255111/26, 3)) 

# size class sizes
SC1 <- 0
SC2 <- 10
SC3 <- 100
SC4 <- 900
SC5 <- 4000

A_mids <- c((SC1 + SC2)/2, (SC2 + SC3)/2, (SC3 + SC4)/2, (SC4 + SC5)/2, 9325) 

# fecundity probabilities
p_fec <- c(0, 0, 0.1, 0.43, 0.88)

# oocyte density
d_ooc <- 63.6

A_mids*d_ooc*p_fec

# estimating fecundity parameters
# Fundemar's report: 26 colonies, 1255111 embryos collected
# assume they were all size class 4, what is the yield?

1255111/(A_mids[4]*p_fec[4]*d_ooc*26)

```










