---
title: "rest_pars"
output: html_document
date: "2026-01-30"
---

README: estimating restoration parameters from Fundemar's data

```{r setup, include=FALSE}

# command + option + i for new code chunk
# command + shift + enter to run entire code chunk
# command + enter to run single line of code

library(tidyverse) # command + shift + m for %>%
library(readxl)
library(janitor)

```


from the original "Adicion_Sustratos" matrix (which was just for DLAB in 2025), it looks like they put ~100 tiles in each tank (divided among substrate types) and had 40 tanks

Goal is to know 1) how many embryos they put in each tank, 2) how many tiles per tank, and 3) settlement rates on the tiles. Want to know 1) and 2) to estimate how many tiles they would use for a given number of embryos, and need to know 3) for model parameters


Adicion_Sustratos_Unificada = spawning event, number of embryos, Tank ID, substrate type, and number of substrates

Matriz_Asentamiento_Unificada = number of settlers on the substrates (Tank ID should match A_S_U)

Matriz_Supervivencia_Reclutas_Unificada = recruit survival, but comparing time at T0 to number of settlers in  M_A_U matrix should give survival in the lab?)


```{r}
# read in the data

add_subs <- read_excel("data/Fundemar/Adicion_Sustratos_Unificada.xlsx", sheet = 1) %>% clean_names()

add_subs_APAL <- add_subs %>% filter(species == "APAL")

#View(add_subs_APAL)

sett_mat <- read_excel("data/Fundemar/Matriz_Asentamiento_Unificada.xlsx", sheet = 1) %>% clean_names()

sett_mat_APAL <- sett_mat %>% filter(species == "APAL")

#View(sett_mat_APAL)

surv_mat <- read_excel("data/Fundemar/Matriz_Supervivencia_Reclutas_Unificada.xlsx", sheet = 1) %>% clean_names()

#unique(surv_mat$coral_species)

surv_mat_APAL <- surv_mat %>% filter(coral_species == "Acropora palmata")

# View(surv_mat_APAL) # doesn't have 2025 yet

```


```{r}
# estimating embryos per tank and number of tiles per tank

#View(add_subs_APAL)

#add_subs_APAL %>% group_by(species, spawning_event, spawning_day, spawning_month, spawning_year, year_of_substrate_addition, month_of_substrate_addition, day_of_substrate_addition, tank_id) %>% summarize(n_substrates = sum(amount_of_substrates)) %>% ungroup() %>% arrange(tank_id) %>% view()


#add_subs_APAL %>% group_by(species, spawning_event, spawning_day, tank_id) %>% summarize(n_substrates = sum(amount_of_substrates), embryos_ml_mn = min(embryos_ml), embryos_ml_mx = min(embryos_ml)) %>% ungroup() %>% arrange(spawning_day) %>% mutate(emb_diff = embryos_ml_mx - embryos_ml_mn) %>% view()


tank_counts <- add_subs_APAL %>% group_by(species, spawning_event, spawning_day, tank_id) %>% summarize(n_substrates = sum(amount_of_substrates), embryos_ml = min(embryos_ml), n_embryos = min(no_embryos), vol_embryos_ml = min(volume_of_embryos_ml)) %>% ungroup() %>% arrange(spawning_day)# %>% view()

#View(tank_counts)

mean(tank_counts$embryos_ml) # mean volume put in the tanks = 7293
min(tank_counts$embryos_ml) # 7250
max(tank_counts$embryos_ml) # 7330

mean(tank_counts$vol_embryos_ml) # mean embryos per mL = 1.9
min(tank_counts$vol_embryos_ml) # min = 1.5
max(tank_counts$vol_embryos_ml) # max = 2.07

mean(tank_counts$n_substrates)

#unique(add_subs$embryos_ml[which(add_subs$species %in% c("APAL", "DLAB"))])

#hist(add_subs$embryos_ml)

```

So on average, they are putting about 7,300 mLs in a tank, and on average there are 2 embryos per mL, and they have about 100 substrates per tank

```{r}
# now estimate settlement rates

#unique(add_subs_APAL$spawning_year) # all 2025

sett_mat_APAL_summ <- sett_mat_APAL %>% filter(year_scored == 2025) %>% group_by(month_scored, day_scored, id_rearing_container) %>%  summarize(n_substrates_sett = n(), n_settlers = sum(as.numeric(number_of_total_settlers))) %>% ungroup() %>% rename(tank_id = id_rearing_container)

View(sett_mat_APAL_summ)

sett_rates_APAL <- tank_counts %>% left_join(sett_mat_APAL_summ, by = "tank_id")

View(sett_rates_APAL)


# CHECK what the day scored means

```






