---
title: "rse_new_scenario_analyses"
output: html_document
date: "2026-02-13"
---

README: Primary model analyses (evaluating different management scenarios).

- MODEL FUNCTION: rse_mod1() from rse_funs.R (current model with density dependence)
- DEPENDS ON: coral_demographic_funs.R, rse_funs.R, Detmer-2025-coral-parameters repo
- STATUS: Current — primary analysis file for the manuscript
- ANALYSES (all in development): (1) Orchard expansion — different proportions of substrates outplanted to orchard vs. reef, with/without disturbance, sensitivity to 6+ parameters. (2) Lab grow-out — keeping substrates in the lab for a year. (3) Orchard transplanting - moving adult corals from the orchard to the reef
- SUPERSEDES: rse_scenario_analyses.Rmd (initial scenario analyses with older version of the rse model)


```{r setup, include=FALSE}

# command + option + i for new code chunk
# command + shift + enter to run entire code chunk
# command + enter to run single line of code

# load packages
library(tidyverse) # command + shift + m for %>%
library(tictoc)

# load the model functions
source("coral_demographic_funs.R")
source("rse_funs.R")
```


# model set up

read in the parameter data and set default parameter values

```{r}
# NOTE: Data files are stored in the separate Detmer-2025-coral-parameters repository.
# Update DATA_PATH to point to your local copy of that repository.
DATA_PATH <- "../Detmer-2025-coral-parameters"

# fragmentation data
apal_frag_summ <- read.csv(file.path(DATA_PATH, "05_data/standardized/apal_fragmentation_summ.csv"))
apal_frag <- read.csv(file.path(DATA_PATH, "05_data/standardized/apal_fragmentation.csv"))


# survival data
field_surv <- readRDS(file.path(DATA_PATH, "parameter_lists/field_surv_pars.rds"))
nurs_surv <- readRDS(file.path(DATA_PATH, "parameter_lists/nurs_surv_pars.rds"))
lab_surv <- readRDS(file.path(DATA_PATH, "parameter_lists/lab_surv_pars.rds"))

# recruit survival
recruit_surv <- readRDS(file.path(DATA_PATH, "parameter_lists/recruit_surv_pars.rds"))

# growth data
field_growth <- readRDS(file.path(DATA_PATH, "parameter_lists/field_growth_pars.rds"))
nurs_growth <- readRDS(file.path(DATA_PATH, "parameter_lists/nurs_growth_pars.rds"))
lab_growth <- readRDS(file.path(DATA_PATH, "parameter_lists/lab_growth_pars.rds"))

# short-term lab survival data
apal_lab_short_surv <- read.csv(file.path(DATA_PATH, "05_data/standardized/apal_surv_lab_short.csv"))


# change 1: format the growth data to be formatted as transition matrices and add these to the growth lists
field_Tmats <- growth_format(growth_dt = field_growth)
field_growth$mat_list <- field_Tmats$mat_list
field_growth$summ_list <- field_Tmats$summ_list

# nurs_Tmats <- growth_format(growth_dt = nurs_growth)
# nurs_growth$mat_list <- nurs_Tmats$mat_list
# nurs_growth$summ_list <- nurs_Tmats$summ_list

# TEMPORARY NEED TO FIX: currently don't have a trans_summary for the nursery data so just sub the field data for now
nurs_growth$mat_list <- field_Tmats$mat_list
nurs_growth$summ_list <- field_Tmats$summ_list


# update formatting of the recruit survival to match field survival data
recruit_surv$SC_surv_df <- data.frame(
  prop_survived = recruit_surv$survival_boot,
  size_class = rep(1, length(recruit_surv$survival_boot)),
  replicate = c(1:length(recruit_surv$survival_boot))
)

recruit_surv$SC_surv_summ_df <- recruit_surv$survival_summary %>% rename(mean = mean_survival, sd = sd_survival) %>% select(-ci_lower, -ci_upper)
size_class_col <- data.frame(
  size_class = rep(1, nrow(recruit_surv$SC_surv_summ_df))
)

recruit_surv$SC_surv_summ_df <- cbind(size_class_col, recruit_surv$SC_surv_summ_df)


```


parameter sets

```{r}

# set upt the orchard, reef, and lab treatments
orchard_treatments <- c("orchard1") # orchards (could have multiple orchards and/or orchards with different post-outplanting treatments)
reef_treatments <- c("reef1") # reef treatments/subpopulations (e.g., could have "urchin outplanting" or "algal removal" or other postoutplanting treatments)
lab_treatments <- c("0_T1", "1_T1") # lab treatments "TX" is tile type/treatment, "X_" indicates whether recruits are outplanted immediately (0_) or the next year (0_1)


# demographic parameters for each orchard and reef
field_surv1 <- field_surv
recruit_surv1 <- recruit_surv # recruit survival rates in field (= reef)
field_growth1 <- field_growth
nurs_surv1 <- nurs_surv
nurs_growth1 <- nurs_growth

# calculate mean parameter values (survival, growth, shrinkage, fragmentation) for each 
all_pars <- default_pars_fun(n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), summ_metric_list = list(field_surv = "mean", recruit_surv = "mean", field_growth = "mean", field_shrink = "mean", field_frag = "mean", nurs_surv = "mean", nurs_growth = "mean", nurs_shrink = "mean"), field_surv1, recruit_surv1, field_growth1, nurs_surv1, nurs_growth1, apal_frag_summ)


# calculate random parameter sets
n_sample1 <- 100 # number of parameter sets

set.seed(500)
all_pars_R <- rand_pars_fun(n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), n_sample = n_sample1, field_surv = field_surv1, recruit_surv = recruit_surv1, field_growth = field_growth1, nurs_surv = nurs_surv1, nurs_growth = nurs_growth1, apal_frag)

```






```{r}

years <- 51 # number of years in simulation
n <- 5 # number of size classes

# define size class boundaries
SC1 <- 0
SC2 <- 10
SC3 <- 100
SC4 <- 900
SC5 <- 4000

A_mids <- c((SC1 + SC2)/2, (SC2 + SC3)/2, (SC3 + SC4)/2, (SC4 + SC5)/2, 9325) # midpoint areas of each size class (cm^2); last value is the 50% quantile of observations greater than SC5 in the standardized data from the literature

A_reef <- mean(c(4492.011, 10076.67, 8943.55)) # area of the reef (very rough estimates from kml file of Fundemar's sites)


# reef parameters: survival
surv_pars.r <- all_pars$surv_pars.r
surv_pars.rc <- all_pars$surv_pars.rc
# reef parameters: growth
growth_pars.r <- all_pars$growth_pars.r
# reef parameters: shrinkage
shrink_pars.r <- all_pars$shrink_pars.r
# reef parameters: fragmentation
frag_pars.r <- all_pars$frag_pars.r

# density dependent survival (within first year post-outplanting only)
dens_pars.r <- list()
dens_pars.r[[1]] <- list() # first reef treatment/subpop
dens_pars.r[[1]][[1]] <- 0.02 # first source to first reef (external recruits)
dens_pars.r[[1]][[2]] <- dens_pars.r[[1]][[1]] # second source to first reef (tiles outplanted immediately)
dens_pars.r[[1]][[3]] <- dens_pars.r[[1]][[1]] # third source to first reef (tiles retained in lab)

# reef fecundity
fec_pars.r <- list()
fec_pars.r[[1]] <- list()
fec_pars.r[[1]][[1]] <- c(0, 0, rep(1255111/26, 3)) 
fec_pars.r[[1]][[2]] <- fec_pars.r[[1]][[1]]
fec_pars.r[[1]][[3]] <- fec_pars.r[[1]][[1]]

# orchard parameters: survival
surv_pars.o <-all_pars$surv_pars.o 
# orchard parameters: growth
growth_pars.o <- all_pars$growth_pars.o
# orchard parameters: shrinkage
shrink_pars.o <- all_pars$shrink_pars.o
# orchard parameters: fragmentation
frag_pars.o <- all_pars$frag_pars.o

# density dependent survival (smallest size class only for now)
dens_pars.o <- list()
dens_pars.o[[1]] <- list() # first orchard
dens_pars.o[[1]][[1]] <- 0.02
dens_pars.o[[1]][[2]] <- 0.02

# orchard parameters: fecundity
fec_pars.o <- list()
fec_pars.o[[1]] <- list() # first treatment
fec_pars.o[[1]][[1]] <- c(0, 0, rep(1255111/26, 3)) # first source
fec_pars.o[[1]][[2]] <- c(0, 0, rep(1255111/26, 3)) # second source

lambda <- 0 # external recruitment to reef

lambda_R <- 1255111 # babies collected from the reference reef each year; from Table 1 of Fundemar's report (total embryos collected in 2025)

# stochasticity parameters
sigma_s <- 0 # sd in in survival
sigma_f <- 0 # sd in fecundity
ext_rand <- c(FALSE, FALSE) # whether 1) external recruitment and 2) reference reef reproduction is stochastic 
seeds <- c(1000, 5000, 10000, 40000)


# disturbance parameters
dist_yrs <- NA# years when disturbance occurs

# effects of disturbance on each reef subpop
dist_effects.r <- list()
dist_effects.r[[1]] <- list() # list where each element is the effects of disturbance on corals from each source in the 1st reef subpop (either "survival" for survival, "Tmat" for growth/survival, "Fmat" for fragmentation, or "fecundity" for reproduction)
dist_effects.r[[1]][[1]] <- list() # effects of disturbances on corals from first source in first reef subpop
dist_effects.r[[1]][[1]][[1]] <- c("survival") # [[1]][[1]][[x]] = effects of xth disturbance on corals from first source in first reef subpop
dist_effects.r[[1]][[2]] <- list() # second source
dist_effects.r[[1]][[2]][[1]] <- c("survival")
dist_effects.r[[1]][[3]] <- list() # third source
dist_effects.r[[1]][[3]][[1]] <- c("survival")



# disturbance severity (survival parameters in disturbance year)
DS.r <- surv_pars.r
DS.r[[1]][[1]] <- DS.r[[1]][[1]]*0.1
DS.r[[1]][[2]] <- DS.r[[1]][[2]]*0.1
DS.r[[1]][[3]] <- DS.r[[1]][[3]]*0.1

DS.rc <- surv_pars.rc
DS.rc[[1]][[1]] <- DS.rc[[1]][[1]]*0.1
DS.rc[[1]][[2]] <- DS.rc[[1]][[2]]*0.1
DS.rc[[1]][[3]] <- DS.rc[[1]][[3]]*0.1

DS.o <- surv_pars.o
DS.o[[1]][[1]] <- DS.o[[1]][[1]]*0.1
DS.o[[1]][[2]] <- DS.o[[1]][[2]]*0.1

# disturbance parameters for each reef subpop (what happens to the population parameters at each disturbance eve)
dist_pars.r <- list()

# first reef population
dist_pars.r[[1]] <- list() 
# first source to first reef subpop
dist_pars.r[[1]][[1]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[1]], dist_surv0 = list(DS.r[[1]][[1]]), dist_surv_rc0 = list(DS.rc[[1]][[1]]), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL) # list of disturbance parameters for the first source to the first reef population (defaults for each parameter type are NULL unless the disturbance affects them, as specified in dist_effects)
# second source to first reef subpop (lab tiles)
dist_pars.r[[1]][[2]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[2]], dist_surv0 = list(DS.r[[1]][[2]]), dist_surv_rc0 = list(DS.rc[[1]][[2]]), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)
dist_pars.r[[1]][[3]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[3]], dist_surv0 = list(DS.r[[1]][[3]]), dist_surv_rc0 = list(DS.rc[[1]][[3]]), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)


# disturbance effects for each orchard subpop
dist_effects.o <- list()

dist_effects.o[[1]] <- list() 
dist_effects.o[[1]][[1]] <- list() 
dist_effects.o[[1]][[1]][[1]] <- c("survival") 
dist_effects.o[[1]][[2]] <- list() 
dist_effects.o[[1]][[2]][[1]] <- c("survival") 

# disturbance parameters for each orchard subpop
dist_pars.o <- list()

# first orchard treatment
dist_pars.o[[1]] <- list() 
# first source in first orchard treatment
dist_pars.o[[1]][[1]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.o[[1]][[1]], dist_surv0 = list(DS.o[[1]][[1]]), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL) 
# second source in first orchard
dist_pars.o[[1]][[2]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.o[[1]][[2]], dist_surv0 = list(DS.o[[1]][[2]]), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL) 


# lab parameters
# fractions of settled larvae in each lab treatment that survive to immediate outplanting each year
s0 <- matrix(NA, nrow = years, ncol = 2)
s0[,1] <- rep(0.95, years)
s0[,2] <- rep(0.95, years)

#s0[10,1] <- 0.05
#s0[10,2] <- 0.05

# fractions of settled larvae in each lab treatment that survive to 1 yr outplanting
s1 <- matrix(NA, nrow = years, ncol = 2)
s1[,1] <- rep(0.7, years)
s1[,2] <- rep(0.7, years)

m0 <- c(0.02, 0.02) # density dependent mortality rate of larvae in each lab treatment that get outplanted immediately
m1 <- c(0.02, 0.02) # density dependent mortality rate of larvae in each lab treatment that get outplanted after 1 year
#sett_props <- list(T1 = 0.15, T2 = 0.15) # proportion of larvae that settle on each tile type
sett_props <- list(T1 = 0.15) # proportion of larvae that settle on each tile type
size_props <- matrix(NA, nrow = length(lab_treatments), ncol = n) # matrix for fractions of retained recruits in each size class at the end of their year in the lab
size_props[1, ] <- c(1, 0, 0, 0, 0) # first lab treatment 
#size_props[2, ] <- c(1, 0, 0, 0, 0) # second lab treatment 
size_props1 <- matrix(NA, nrow = length(lab_treatments), ncol = n) # matrix for fractions of retained recruits in each size class at the end of their year in the lab
#size_props1[1, ] <- c(1, 0, 0, 0, 0) # first lab treatment 
size_props1[2, ] <- c(1, 0, 0, 0, 0) # second lab treatment 

lab_pars <- list(s0 = s0, s1 = s1, m0 = m0, m1 = m1, sett_props = sett_props, size_props = size_props, size_props1 = size_props1)

# restoration strategy parameters
tile_props <- list(T1 = 1) # proportion of tiles that are each type
orchard_yield <- 1 # percent of new orchard babies successfully collected
reef_yield <- 1 # percent of new reef babies successfully collected and fertilized
spawn_target <- 2000000 # target number of embryos to collect from orchard each year (note: if this is set to zero, there won't be any settlers)
reef_prop <- c(0.5, 1) # proportion of lab recruits from each lab treatment outplanted to reef (1-proportion outplanted to orchard)
reef_out_props <- matrix(NA, nrow = length(lab_treatments), ncol = length(reef_treatments)) # proportion recruits going from each lab treatment to each reef treatment (row = origin lab treatment, column = destination reef treatment)
reef_out_props[1,] <- c(1) # from first lab treatment to each reef treatment (here there's just on intervention reef so they all go there)
reef_out_props[2,] <- c(1) # from second lab treatment to each reef treatment 

orchard_out_props <- matrix(NA, nrow = length(lab_treatments), ncol = length(orchard_treatments)) # proportion recruits going from each lab treatment to each orchard treatment
orchard_out_props[1,] <- c(1) # from first lab treatment to each orchard 
orchard_out_props[2,] <- c(1) # from second lab treatment to each orchard 

# sizes allocated to each treatment
reef_areas <- c(A_reef)*10000 # m^2 given to each post-outplanting reef treatment/subpopulation, convert to cm^2 by multiplying by 10,000
lab_max <- 4000 # total number of tiles that the lab can accommodate 
lab_retain_max <- 0 # 3100*0.5 # total number of tiles that the lab can keep for a year (must be less than or equal to lab_max; if > 0 then must have at least one "1_" lab treatment, if equal to lab max then must have all treatments be "1_")
tank_min <- 14600 # min number of embryos to put in a tank 
tank_max <- 33333 # max number of embryos to put in a tank

orchard_size <- c(30*500) # number of tiles each orchard has space for (500 stars x 30 tiles per star)

# coral transplanting
transplant <- rep(0, years) # years in which corals get transplanted from the orchard
#transplant[10] <- 1 # in 10th year, move corals from orchard to reef

# null_mat <- matrix(0, nrow = years, ncol = n)
null_mat <- matrix(c(0, 0.75*orchard_size, 0, 0, 0), nrow = years, ncol = n, byrow = T)
#trans_mats[[1]][[1]][which(transplant!=0)[1],] <- c(0, 0, 0, 2,0)# max number of colonies of each size class to move to reef in first transplant event
trans_mats <- list()
trans_mats[[1]] <- list() # first orchard 
trans_mats[[1]][[1]] <- null_mat # number of corals in each size class that originated from the first source of colonies for this orchard to transplant at each timepoint
trans_mats[[1]][[2]] <- null_mat # number of corals in each size class that originated from the second source of colonies for this orchard to transplant at each timepoint

# where on the reef to put the transplants
#trans_reef[[1]][[1]][which(transplant!=0)[1],] <- c(1, 1)# reef and source subpop to transport the corals to
trans_reef <- list()
trans_reef[[1]] <- list() # first orchard
trans_reef[[1]][[1]] <- matrix(c(1, 2), nrow = years, ncol = 2, byrow = T) # first source corals in this orchard get transplanted to the second source of corals in the first reef
trans_reef[[1]][[2]] <- matrix(c(1, 3), nrow = years, ncol = 2, byrow = T) # second source corals in this orchard get transplanted to the third source of corals in the second reef

trans_surv <- c(1, 1, 1, 1, 1) # proportion of colonies of each size class that survival transplanting

rest_pars <- list(tile_props = tile_props, orchard_yield = orchard_yield, reef_yield = reef_yield, spawn_target = spawn_target, reef_prop = reef_prop, reef_out_props = reef_out_props, orchard_out_props = orchard_out_props, reef_areas = reef_areas, lab_max = lab_max, lab_retain_max = lab_retain_max, tank_min = tank_min, tank_max = tank_max, orchard_size = orchard_size, transplant = transplant, trans_mats = trans_mats, trans_reef = trans_reef, trans_surv = trans_surv)


# initial conditions in each reef subpopulation
N0.r <- list()
N0.r[[1]] <- list() # second reef subpop
N0.r[[1]][[1]] <- rep(0, n) # first source (external recruits)
N0.r[[1]][[2]] <- rep(0, n) # second source
N0.r[[1]][[3]] <- rep(0, n) # third source

# initial conditions in each orchard subpopulation
N0.o <- list()
N0.o[[1]] <- list() # first orchard treatment
N0.o[[1]][[1]] <- rep(0, n) # first source
N0.o[[1]][[2]] <- rep(0, n) # second source

N0.l <- list()
N0.l[[1]] <- 1000 # start with some babies in the lab
N0.l[[2]] <- 0 

sim1 <- rse_mod1(years, n, A_mids, surv_pars.rc, surv_pars.r, dens_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, surv_pars.o, dens_pars.o, growth_pars.o, shrink_pars.o, frag_pars.o, fec_pars.o, lambda, lambda_R, sigma_s, sigma_f, ext_rand, seeds, dist_yrs, dist_pars.r, dist_effects.r, dist_pars.o, dist_effects.o, orchard_treatments, reef_treatments, lab_treatments, lab_pars, rest_pars, N0.r, N0.o, N0.l)

```


check output

```{r}


#sim1$reef_out[[1]][[2]]

#(sim1$reef_out[[1]][[2]])/sim1$tiles_out_tot


#sim1$tiles_out_tot

reef1_tot <- model_summ(model_sim = sim1, location = "reef", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

orch1_tot <- model_summ(model_sim = sim1, location = "orchard", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))


plot(x = c(1:years), y = reef1_tot, type = "l", las = 1, ylim = c(0, max(max(reef1_tot), max(orch1_tot))), xlab = "year", ylab = "Area covered", lwd = 2)
lines(x = c(1:years), y = orch1_tot, type = "l", col = "dodgerblue", lwd = 2, lty = 2)
legend("topleft", legend = c("reef", "orchard"), col = c("black", "dodgerblue"), lwd = 2, lty = c(1, 2))


reef1_tot <- model_summ(model_sim = sim1, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

orch1_tot <- model_summ(model_sim = sim1, location = "orchard", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))


plot(x = c(1:years), y = reef1_tot, type = "l", las = 1, ylim = c(0, max(max(reef1_tot), max(orch1_tot))), xlab = "year", ylab = "Population size", lwd = 2)
lines(x = c(1:years), y = orch1_tot, type = "l", col = "dodgerblue", lwd = 2, lty = 2)
legend("topleft", legend = c("reef", "orchard"), col = c("black", "dodgerblue"), lwd = 2, lty = c(1, 2))


plot(x = c(1:years), y = reef1_tot, type = "l", las = 1, ylim = c(0, max(reef1_tot)), xlab = "year", ylab = "Area covered", lwd = 2)

plot(x = c(1:years), y = sim1$reef_out[[1]][[2]], type = "l", ylim = c(0, max(max(sim1$reef_out[[1]][[2]], na.rm = T), max(sim1$orchard_out[[1]][[1]], na.rm = T))))
lines(x = c(1:years), y = sim1$orchard_out[[1]][[1]], col = "dodgerblue")

#sim1$reef_out

```


set up default disturbance parameters 

```{r}

dist_yrs3 <- c(1:max(years))[c(F, T,F)] # years when disturbance occurs: once every 3 years

# list where elements are for each group, each element is matrix where rows are time points with disturbances and columns are size classes

dist_mat.r <- matrix(0.2, nrow = length(dist_yrs3), ncol = 5)
dist.r <- list(dist_mat.r, dist_mat.r, dist_mat.r) # disturbance impacts on survival for each source on the reef

#dist_mat.o <- matrix(0.4, nrow = length(dist_yrs3), ncol = 5)
dist_mat.o <- matrix(1, nrow = length(dist_yrs3), ncol = 5)
dist.o <- list(dist_mat.o, dist_mat.o) # disturbance impacts on survival for each source on the orchard

dist_pars_list3 <- list(dist.r = dist.r, dist.o = dist.o)


# repeat for disturbance every 5 years
dist_yrs5 <- c(1:max(years))[c(F, T, F, F, F)] # years when disturbance occurs: once every 5 years but with first disturbance in year 3

# list where elements are for each group, each element is matrix where rows are time points with disturbances and columns are size classes

dist_mat.r <- matrix(0.2, nrow = length(dist_yrs5), ncol = 5)
dist.r <- list(dist_mat.r, dist_mat.r, dist_mat.r) # disturbance impacts on survival for each source on the reef

dist_mat.o <- matrix(1, nrow = length(dist_yrs5), ncol = 5)
dist.o <- list(dist_mat.o, dist_mat.o) # disturbance impacts on survival for each source on the orchard

dist_pars_list5 <- list(dist.r = dist.r, dist.o = dist.o)


# repeat for disturbance every 7 years
dist_yrs7 <- c(1:max(years))[c(F, T,F, F, F, F, F)] # years when disturbance occurs: once every 7 years, with first disturbance in year 3

# list where elements are for each group, each element is matrix where rows are time points with disturbances and columns are size classes

dist_mat.r <- matrix(0.2, nrow = length(dist_yrs7), ncol = 5)
dist.r <- list(dist_mat.r, dist_mat.r, dist_mat.r) # disturbance impacts on survival for each source on the reef

dist_mat.o <- matrix(1, nrow = length(dist_yrs7), ncol = 5)
dist.o <- list(dist_mat.o, dist_mat.o) # disturbance impacts on survival for each source on the orchard

dist_pars_list7 <- list(dist.r = dist.r, dist.o = dist.o)


```



# functions

Note these functions are defined here rather than in rse_funs.R because they depend on variables from this analysis environment (rest_pars, years, etc.)

## SC1 scenarios

```{r}


# function for running orchard expansion simulations with different restoration parameter values
# prop_set = vector with proportions of substrates to outplant to orchard
# dem_pars = set of demographic parameters to use
# par_set2 = values of the second parameter to iterate over
# par2_name = name of the second parameter to iterate over (NOTE: function currently only allows for parameters that are in rest_pars list, lab_pars list, or "lambda_R", "frag_pars.o", fec_pars.o, "surv_pars.o_1", and "surv_pars.r_1")
# dist = whether or not there is a disturbance regime (T or F)
orch_exp_fun1 <- function(dem_pars, prop_set, par_set2 = NULL, par2_name = "none", dist = F, dist_yrs = NULL, dist_pars_list = NULL){
  
  out_list <- list()


# outer loop = demographic parameter values
# inner loop = proportion to outplant to the reef


L_outer <- ifelse(par2_name == "none", 1, length(par_set2))  

for(i in 1:L_outer){ # for each parameter set
  
  # set up the parameters
  
  surv_pars.r <- dem_pars$surv_pars.r
  surv_pars.rc <- dem_pars$surv_pars.rc
  growth_pars.r <- dem_pars$growth_pars.r
  shrink_pars.r <- dem_pars$shrink_pars.r
  frag_pars.r <- dem_pars$frag_pars.r
  surv_pars.o <- dem_pars$surv_pars.o
  growth_pars.o <- dem_pars$growth_pars.o
  shrink_pars.o <- dem_pars$shrink_pars.o
  frag_pars.o <- dem_pars$frag_pars.o
  
 if(par2_name == "frag_pars.o"){
frag_pars.o[[1]][[1]][[4]] <- frag_pars.r[[1]][[1]][[4]]*par_set2[i]

frag_pars.o[[1]][[1]][[5]] <- frag_pars.r[[1]][[1]][[5]]*par_set2[i]

 }
  
  if(par2_name == "surv_pars.o_1"){
surv_pars.o[[1]][[1]][1] <- surv_pars.r[[1]][[1]][1]*par_set2[i]

  }
  
  if(par2_name == "surv_pars.r_1"){
surv_pars.r[[1]][[1]][1] <- surv_pars.r[[1]][[1]][1]*par_set2[i]
surv_pars.r[[1]][[2]][1] <- surv_pars.r[[1]][[1]][1]*par_set2[i]

  }
  
  if(par2_name == "fec_pars.o"){
    fec_pars.o <- list()
    fec_pars.o[[1]] <- list() # first treatment
    fec_pars.o[[1]][[1]] <- c(0, 0, rep(1255111/26, 3))*par_set2[i]
  }
  
  # set up disturbance
  if(dist == F){
    
    dist_yrs <- NA
    
  } else{
    # update with disturbance effects
  
# effects of disturbance on each reef subpop
dist_effects.r <- list()
dist_effects.r[[1]] <- list() 
dist_effects.r[[1]][[1]] <- list() # effects of disturbances on corals from first source in first reef subpop
dist_effects.r[[1]][[1]] <- as.list(rep("survival", length(dist_yrs))) # effects of each disturbance on corals from first source in first reef subpop 
dist_effects.r[[1]][[2]] <- list() # second source
dist_effects.r[[1]][[2]] <- as.list(rep("survival", length(dist_yrs)))
dist_effects.r[[1]][[3]] <- list() # third source
dist_effects.r[[1]][[3]] <- as.list(rep("survival", length(dist_yrs)))


# disturbance effects for each orchard subpop
dist_effects.o <- list()

dist_effects.o[[1]] <- list() 
dist_effects.o[[1]][[1]] <- list() # effects of disturbances on corals from first source in first orchard treatment
dist_effects.o[[1]][[1]] <- as.list(rep("survival", length(dist_yrs))) 
dist_effects.o[[1]][[2]] <- list() # second source
dist_effects.o[[1]][[2]] <- as.list(rep("survival", length(dist_yrs))) 


surv_pars.r1 <- list(list(), list(), list()) # holding list for survival following disturbance events for each source to the reef
surv_pars.rc1 <- list(list(), list(), list()) # same for recruit survival
surv_pars.o1 <- list(list(), list()) # holding list for survival following disturbance events for each source to the orchard

for(jj in 1:length(dist_yrs)){ # for each year with a disturbance
  
  dist_pars_list$dist.r[[1]][jj,]
  
  # update the survival parameters
  surv_pars.r1[[1]][[jj]] <- surv_pars.r[[1]][[1]]*dist_pars_list$dist.r[[1]][jj,] # note this is the same as the proportional reduction in reference reef production
  surv_pars.r1[[2]][[jj]] <- surv_pars.r[[1]][[2]]*dist_pars_list$dist.r[[2]][jj,]
  surv_pars.r1[[3]][[jj]] <- surv_pars.r[[1]][[3]]*dist_pars_list$dist.r[[3]][jj,]
  
  surv_pars.rc1[[1]][[jj]] <- surv_pars.rc[[1]][[1]]*dist_pars_list$dist.rc[[1]][jj] # note this is the same as the proportional reduction in reference reef production
  surv_pars.rc1[[2]][[jj]] <- surv_pars.rc[[1]][[2]]*dist_pars_list$dist.rc[[2]][jj]
  surv_pars.rc1[[3]][[jj]] <- surv_pars.rc[[1]][[3]]*dist_pars_list$dist.rc[[3]][jj]
  
  surv_pars.o1[[1]][[jj]] <- surv_pars.o[[1]][[1]]*dist_pars_list$dist.o[[1]][jj,]
  surv_pars.o1[[2]][[jj]] <- surv_pars.o[[1]][[2]]*dist_pars_list$dist.o[[2]][jj,]

}


# disturbance parameters for each reef subpop (what happens to the population parameters at each disturbance event)
dist_pars.r <- list()
dist_pars.r[[1]] <- list() 
dist_pars.r[[1]][[1]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[1]], dist_surv0 = surv_pars.r1[[1]], dist_surv_rc0 = surv_pars.rc1[[1]], dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL) 
dist_pars.r[[1]][[2]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[2]], dist_surv0 = surv_pars.r1[[2]], dist_surv_rc0 = surv_pars.rc1[[2]], dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)
dist_pars.r[[1]][[3]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[3]], dist_surv0 = surv_pars.r1[[3]], dist_surv_rc0 = surv_pars.rc1[[3]], dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)


# disturbance parameters for each orchard subpop
dist_pars.o <- list()
dist_pars.o[[1]] <- list() 
dist_pars.o[[1]][[1]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.o[[1]][[1]], dist_surv0 = surv_pars.o1[[1]], dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)
dist_pars.o[[1]][[2]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.o[[1]][[2]], dist_surv0 = surv_pars.o1[[2]], dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)
  
  
  }
  

  
  # holding list for model simulations with this parameter set
  mod_list <- list()
  
  for(j in 1:length(prop_set)){ # for each proportion outplanted to reef
    
    lab_pars_ij <- lab_pars
    rest_pars_ij <- rest_pars
    
    rest_pars_ij$reef_prop <- c(prop_set[j], 1) # proportion of tiles outplanted to reef for each treatment
    
    if(par2_name %in% names(rest_pars_ij)){
      rest_pars_ij[[which(names(rest_pars_ij) == par2_name)]] <- par_set2[i]
    } 
    
    if(par2_name %in% names(lab_pars_ij)){
      lab_pars_ij[[which(names(lab_pars_ij) == par2_name)]] <- par_set2[i]
    } 
    
    
    if(par2_name == "lambda_R"){
      lambda_R <- par_set2[i]
    } 
    
    # run the model
    sim_ij <- rse_mod1(years, n, A_mids, surv_pars.rc, surv_pars.r, dens_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, surv_pars.o, dens_pars.o, growth_pars.o, shrink_pars.o, frag_pars.o, fec_pars.o, lambda, lambda_R, sigma_s, sigma_f, ext_rand, seeds, dist_yrs, dist_pars.r, dist_effects.r, dist_pars.o, dist_effects.o, orchard_treatments, reef_treatments, lab_treatments, lab_pars_ij, rest_pars_ij, N0.r, N0.o, N0.l)
    
    # save the output
    mod_list[[j]] <- sim_ij
    
  } # end of iterations over outplanting proportions
  
  
  out_list[[i]] <- mod_list
  
} # end of iterations over parameter sets

return(out_list)
  
}


# wrapper function for iterating over multiple demographic parameter sets
# dem_pars_all = list with all sets of demographic parameters
# n_pars = number of parameter sets in dem_pars_all
orch_exp_fun2 <- function(dem_pars_all, n_pars, prop_set, par_set2 = NULL, par2_name = "none", dist = F, dist_yrs = NULL, dist_pars_list = NULL){
  
  out_list2 <- list() # holding list
  
  for(i in 1:n_pars){
    
    dem_pars <- sapply(dem_pars_all, "[[", i) # select the i^th element from each component of the full list of demographic parameters
    
    dem_pars <- list(surv_pars.r = list(dem_pars$surv_pars_L.r), surv_pars.rc = list(dem_pars$surv_pars_L.rc), growth_pars.r = list(dem_pars$growth_pars_L.r), shrink_pars.r = list(dem_pars$shrink_pars_L.r), frag_pars.r = list(dem_pars$frag_pars_L.r), surv_pars.o = list(dem_pars$surv_pars_L.o), growth_pars.o = list(dem_pars$growth_pars_L.o), shrink_pars.o = list(dem_pars$shrink_pars_L.o), frag_pars.o = list(dem_pars$frag_pars_L.o))
    
    out_list2[[i]] <- orch_exp_fun1(dem_pars, prop_set, par_set2, par2_name, dist, dist_yrs, dist_pars_list)
    
  }
  
  return(out_list2)
  
}


```


```{r}
# https://stackoverflow.com/questions/44176908/r-list-get-first-item-of-each-element
#sapply(all_pars_R, "[[", 100)
```

## SC2 scenarios

```{r}


# function for running orchard expansion simulations with different restoration parameter values
# growth_set = vector with growth of babies kept in lab or in orchard (proportion that transition from size class 1 to 2 after one year)
# surv_set = vector with survival rates after outplanting/transplanting to reef
# out_par_type = either "lab_growth" (growing out babies in lab for a year) or "orch_trans" (transplanting corals from the orchard to the reef)
# dem_pars = set of demographic parameters to use
# par_set2 = values of the second parameter to iterate over
# par2_name = name of the second parameter to iterate over (NOTE: function currently only allows for parameters that are in rest_pars list or "lambda_R", "frag_pars.o", fec_pars.o, "surv_pars.o_1", and "surv_pars.r_1")
# dist = whether or not there is a disturbance regime (T or F)
lab_trans_fun1 <- function(dem_pars, growth_set, surv_set, out_par_type, par_set2 = NULL, par2_name = "none", dist = F, dist_yrs = NULL, dist_pars_list = NULL){
  
  out_list <- list()


# outer loop = demographic parameter values
# inner loop = proportion to outplant to the reef

L_outer <- ifelse(par2_name == "none", 1, length(par_set2))  

for(i in 1:L_outer){ # for each parameter set
  
  # set up the parameters
  
  surv_pars.r <- dem_pars$surv_pars.r
  surv_pars.rc <- dem_pars$surv_pars.rc
  growth_pars.r <- dem_pars$growth_pars.r
  shrink_pars.r <- dem_pars$shrink_pars.r
  frag_pars.r <- dem_pars$frag_pars.r
  surv_pars.o <- dem_pars$surv_pars.o
  growth_pars.o <- dem_pars$growth_pars.o
  shrink_pars.o <- dem_pars$shrink_pars.o
  frag_pars.o <- dem_pars$frag_pars.o
  
 if(par2_name == "frag_pars.o"){
frag_pars.o[[1]][[1]][[4]] <- frag_pars.r[[1]][[1]][[4]]*par_set2[i]

frag_pars.o[[1]][[1]][[5]] <- frag_pars.r[[1]][[1]][[5]]*par_set2[i]

 }
  
  if(par2_name == "surv_pars.o_1"){
surv_pars.o[[1]][[1]][1] <- surv_pars.r[[1]][[1]][1]*par_set2[i]

  }
  
  if(par2_name == "surv_pars.r_1"){
surv_pars.r[[1]][[1]][1] <- surv_pars.r[[1]][[1]][1]*par_set2[i]
surv_pars.r[[1]][[2]][1] <- surv_pars.r[[1]][[1]][1]*par_set2[i]

  }
  
  if(par2_name == "fec_pars.o"){
    fec_pars.o <- list()
    fec_pars.o[[1]] <- list() # first treatment
    fec_pars.o[[1]][[1]] <- c(0, 0, rep(1255111/26, 3))*par_set2[i]
  }
  
  # set up disturbance
  if(dist == F){
    
    dist_yrs <- NA
    
  } else{
    # update with disturbance effects
  
# effects of disturbance on each reef subpop
dist_effects.r <- list()
dist_effects.r[[1]] <- list() 
dist_effects.r[[1]][[1]] <- list() # effects of disturbances on corals from first source in first reef subpop
dist_effects.r[[1]][[1]] <- as.list(rep("survival", length(dist_yrs))) # effects of each disturbance on corals from first source in first reef subpop 
dist_effects.r[[1]][[2]] <- list() # second source
dist_effects.r[[1]][[2]] <- as.list(rep("survival", length(dist_yrs)))
dist_effects.r[[1]][[3]] <- list() # third source
dist_effects.r[[1]][[3]] <- as.list(rep("survival", length(dist_yrs)))


# disturbance effects for each orchard subpop
dist_effects.o <- list()

dist_effects.o[[1]] <- list() 
dist_effects.o[[1]][[1]] <- list() # effects of disturbances on corals from first source in first orchard treatment
dist_effects.o[[1]][[1]] <- as.list(rep("survival", length(dist_yrs))) 
dist_effects.o[[1]][[2]] <- list() # second source
dist_effects.o[[1]][[2]] <- as.list(rep("survival", length(dist_yrs))) 


surv_pars.r1 <- list(list(), list(), list()) # holding list for survival following disturbance events for each source to the reef
surv_pars.rc1 <- list(list(), list(), list()) # same for recruits
surv_pars.o1 <- list(list(), list()) # holding list for survival following disturbance events for each source to the orchard

for(jj in 1:length(dist_yrs)){ # for each year with a disturbance
  
  dist_pars_list$dist.r[[1]][jj,]
  
  # update the survival parameters
  surv_pars.r1[[1]][[jj]] <- surv_pars.r[[1]][[1]]*dist_pars_list$dist.r[[1]][jj,] # note this is the same as the proportional reduction in reference reef production
  surv_pars.r1[[2]][[jj]] <- surv_pars.r[[1]][[2]]*dist_pars_list$dist.r[[2]][jj,]
  surv_pars.r1[[3]][[jj]] <- surv_pars.r[[1]][[3]]*dist_pars_list$dist.r[[3]][jj,]
  
  
  surv_pars.rc1[[1]][[jj]] <- surv_pars.rc[[1]][[1]]*dist_pars_list$dist.rc[[1]][jj] # note this is the same as the proportional reduction in reference reef production
  surv_pars.rc1[[2]][[jj]] <- surv_pars.rc[[1]][[2]]*dist_pars_list$dist.rc[[2]][jj]
  surv_pars.rc1[[3]][[jj]] <- surv_pars.rc[[1]][[3]]*dist_pars_list$dist.rc[[3]][jj]
  
  surv_pars.o1[[1]][[jj]] <- surv_pars.o[[1]][[1]]*dist_pars_list$dist.o[[1]][jj,]
  surv_pars.o1[[2]][[jj]] <- surv_pars.o[[1]][[2]]*dist_pars_list$dist.o[[2]][jj,]

}


# disturbance parameters for each reef subpop (what happens to the population parameters at each disturbance event)
dist_pars.r <- list()
dist_pars.r[[1]] <- list() 
dist_pars.r[[1]][[1]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[1]], dist_surv0 = surv_pars.r1[[1]], dist_surv_rc0 = surv_pars.rc1[[1]], dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL) 
dist_pars.r[[1]][[2]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[2]], dist_surv0 = surv_pars.r1[[2]], dist_surv_rc0 = surv_pars.rc1[[2]], dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)
dist_pars.r[[1]][[3]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[3]], dist_surv0 = surv_pars.r1[[3]], dist_surv_rc0 = surv_pars.rc1[[3]], dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)


# disturbance parameters for each orchard subpop
dist_pars.o <- list()
dist_pars.o[[1]] <- list() 
dist_pars.o[[1]][[1]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.o[[1]][[1]], dist_surv0 = surv_pars.o1[[1]], dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)
dist_pars.o[[1]][[2]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.o[[1]][[2]], dist_surv0 = surv_pars.o1[[2]], dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL)
  
  
  }
  

  
  # holding list for model simulations with this parameter set
  mod_list <- list()
  
  for(j in 1:length(growth_set)){ # for each growth rate
    
     mod_list_h <- list()
    
    for(h in 1:length(surv_set)){ # for each survival rate
    
    lab_pars_ij <- lab_pars
    rest_pars_ij <- rest_pars
    
    if(par2_name %in% names(rest_pars_ij)){
      rest_pars_ij[[which(names(rest_pars_ij) == par2_name)]] <- par_set2[i]
    } 
    
    if(par2_name %in% names(lab_pars_ij)){
      lab_pars_ij[[which(names(lab_pars_ij) == par2_name)]] <- par_set2[i]
    } 
    
    if(par2_name == "lambda_R"){
      lambda_R <- par_set2[i]
    } 
    
  
    if(out_par_type == "lab_growth"){
    
    surv_pars.r_j <- surv_pars.r
    surv_pars.rc_j <- surv_pars.rc
    
   #LS_mat <- matrix(NA, nrow = length(LS_set), ncol = 2)
   #LS_mat[,2] <- LS_set # SC2 survival rates
   #LS_mat[,1] <- LS_set*surv_pars.r[[1]][[2]][1]/surv_pars.r[[1]][[2]][2]
   # #surv_pars.r_j[[1]][[3]][c(1,2)] <- LS_mat[h,]
   
   # SC1 survival rates: preserve same ratio between SC1 and SC2 survival as in field parameters
    surv_pars.rc_j[[1]][[3]][1] <- LS_set[h]*surv_pars.rc[[1]][[2]][1]/surv_pars.r[[1]][[2]][2]
    # SC2 survival rates:
    surv_pars.r_j[[1]][[3]][2] <- LS_set[h]
    
    lab_pars_ij$size_props1[2, ] <- c(1-growth_set[j], growth_set[j], 0, 0,0)
      
    sim_ij <- rse_mod1(years, n, A_mids, surv_pars.rc_j, surv_pars.r_j, dens_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, surv_pars.o, dens_pars.o, growth_pars.o, shrink_pars.o, frag_pars.o, fec_pars.o, lambda, lambda_R, sigma_s, sigma_f, ext_rand, seeds, dist_yrs, dist_pars.r, dist_effects.r, dist_pars.o, dist_effects.o, orchard_treatments, reef_treatments, lab_treatments, lab_pars_ij, rest_pars_ij, N0.r, N0.o, N0.l)
    }
    
    if(out_par_type == "orch_trans"){
      
      growth_pars.o_j <- growth_pars.o
      growth_pars.o_j[[1]][[1]][[1]][1] <- growth_set[j]
      
      trans_surv_j <- rep(surv_set[h], 5)
      rest_pars_ij$trans_surv <- trans_surv_j
    
      sim_ij <- rse_mod1(years, n, A_mids, surv_pars.rc, surv_pars.r, dens_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, surv_pars.o, dens_pars.o, growth_pars.o_j, shrink_pars.o, frag_pars.o, fec_pars.o, lambda, lambda_R, sigma_s, sigma_f, ext_rand, seeds, dist_yrs, dist_pars.r, dist_effects.r, dist_pars.o, dist_effects.o, orchard_treatments, reef_treatments, lab_treatments, lab_pars_ij, rest_pars_ij, N0.r, N0.o, N0.l)
    }
    
    
    
    # save the output
    mod_list_h[[h]] <- sim_ij
    
  } # end of iterations over survival rates
    
     # save the output
    mod_list[[j]] <- mod_list_h
     
  } # end of iterations over growth rates
  
  
  out_list[[i]] <- mod_list
  
} # end of iterations over parameter sets

return(out_list)
  
}


# wrapper function for iterating over multiple demographic parameter sets
# dem_pars_all = list with all sets of demographic parameters
# n_pars = number of parameter sets in dem_pars_all
lab_trans_fun2 <- function(dem_pars_all, n_pars, growth_set, surv_set, out_par_type, par_set2 = NULL, par2_name = "none", dist = F, dist_yrs = NULL, dist_pars_list = NULL){
  
  out_list2 <- list() # holding list
  
  for(i in 1:n_pars){
    
    dem_pars <- sapply(dem_pars_all, "[[", i) # select the i^th element from each component of the full list of demographic parameters
    
    dem_pars <- list(surv_pars.r = list(dem_pars$surv_pars_L.r), surv_pars.rc = list(dem_pars$surv_pars_L.rc), growth_pars.r = list(dem_pars$growth_pars_L.r), shrink_pars.r = list(dem_pars$shrink_pars_L.r), frag_pars.r = list(dem_pars$frag_pars_L.r), surv_pars.o = list(dem_pars$surv_pars_L.o), growth_pars.o = list(dem_pars$growth_pars_L.o), shrink_pars.o = list(dem_pars$shrink_pars_L.o), frag_pars.o = list(dem_pars$frag_pars_L.o))
    
    out_list2[[i]] <- lab_trans_fun1(dem_pars, out_set, out_par_type, par_set2, par2_name, dist, dist_yrs, dist_pars_list)
    
  }
  
  return(out_list2)
  
}



# second wrapper function for making the data frames
# wrapper function for combining data frames from multiple parameter sets
# sim_list = list with model simulation output for multiple parameters
# max_yr = end of time interval over which to calculate the metrics
# n_pars = number of demographic parameter sets used in sim_list
# prop_set = vector with proportions of substrates to outplant to orchard
# par_set2 = values of the second parameter to iterate over
# par2_name = name of the second parameter iterated over in sim_list
full_dt_fun2 <- function(sim_list, max_yr, n_pars, growth_set, surv_set, out_par_type, par_set2, par2_name){
  
  for(i in 1:n_pars){
  
  for(j in 1:length(par_set2)){
    
    for(k in 1:length(growth_set)){
      
      for(h in 1:length(surv_set)){
        
        if(n_pars == 1){
          sim_ijk <- sim_list[[j]][[k]][[h]]
        } else{
          sim_ijk <- sim_list[[i]][[j]][[k]][[h]]
        }
      
      dt_ijk <- metrics_dt_fun(sim_ijk, max_yr)
      
      dt_ijk$par_rep <- i
      dt_ijk$growth_par <- growth_set[k]
      dt_ijk$surv_par <- surv_set[h]
      dt_ijk$par2 <- par_set2[j]
        
        if(h == 1){
        dt_h <- dt_ijk
      } else{
        dt_h <- rbind(dt_h, dt_ijk)
      }
      
      } # end of h loop
      
      
      if(k == 1){
        dt_k <- dt_h
      } else{
        dt_k <- rbind(dt_k, dt_h)
      }
    } # end of k loop
    
    if(j == 1){
      dt_j <- dt_k
    } else{
      dt_j <- rbind(dt_j, dt_k)
    }
    
  } # end of j loop
  
  if(i == 1){
    dt_i <- dt_j
  } else{
    dt_i <- rbind(dt_i, dt_j)
  }
  
  } # end of i loop
  
  
  dt_i$par_name2 <- par2_name
  dt_i$out_par_type <- out_par_type
  
  return(dt_i)
}

```


## processing output

function for processing the output

```{r}
# function for extracting performance metrics from the simulations
# sim = model simulation output
# max_yr = end of time interval over which to calculate the metrics
# annual costs = whether or not to also return vector with total costs each year from year 2 to max_yr
metrics_dt_fun <- function(sim, max_yr, annual_costs = F){
  
      r_inf <- 0.056 # inflation rate
      
      # fixed costs
      c_RS <- 84.6*rest_pars$orchard_size/30 + 1200 # reef stars (cost of reef stars plus cost of boat rental for installation); no inflation because assume this is only an initial cost
      
      # fixed costs that are per year (need to account for inflation)
      # c_perm <- 400*length(2:max_yr)
      c_perm <- sum(400*(1 + r_inf)^(2:max_yr)) # permits for collecting babies (400 per year)
      c_boat <- sum(1200*4*(1 + r_inf)^(2:max_yr)) # boat maintenance (1200 every 3 months/4x per year)
      c_log <- sum(500*12*(1 + r_inf)^(2:max_yr)) # logistics/personnel (500 per month/12x per year)
      
      c_fixed <- c_RS + c_perm + c_boat + c_log
  
  
    # make the holding data frame
      L1 <- 1
    dt_i <- data.frame(
  reef_cover_mean = rep(NA, L1), # m2 of coral cover on reef (mean over years 1 to max_yr)
  reef_function_mean = rep(NA, L1), # avg area covered by large/reproductive (size classes 3-5) corals on reef
  reef_tiles_out_tot = rep(NA, L1), # total number of tiles outplanted to reef
  reef_area_out_tot = rep(NA, L1), # total area of reef over which tiles were outplanted 
  reef_recruits_out_tot = rep(NA, L1), # total recruits outplanted to reef
  reef_ROI_mean = rep(NA, L1), # average coral cover on reef per dollar spent, averaged over max_yr years
  reefI_ROI_mean = rep(NA, L1), # average number of corals on reef per dollar spent, averaged over max_yr years
  tot_ROI_mean = rep(NA, L1), # average total ROI (total area covered by corals on reef and orchard per dollar spent)
  totI_ROI_mean = rep(NA, L1), # average total ROI in terms of individuals (total number of corals on reef and orchard per dollar spent)
  orch_function_mean = rep(NA, L1), # avg area covered by large/reproductive (size classes 3-5) corals in orchard
 tot_costs = rep(NA, L1), # total amount spent over length of simulation
 c_fixed = rep(NA, L1), # fixed costs
 c_subs = rep(NA, L1), # costs of substrates
 c_out = rep(NA, L1), # outplanting costs
 c_maint = rep(NA, L1), # orchard maintenance costs
 c_lab = rep(NA, L1), # lab costs
 c_spawn = rep(NA, L1), # costs of collecting spawn
 c_trans = rep(NA, L1) # costs of transplanting corals
)
    
  j <- 1
    
      sim_ij <- sim
      
      # reef cover
      reef_cover_ij <- model_summ(model_sim = sim_ij, location = "reef", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))
      
      # reef individuals
      reef_ind_ij <- model_summ(model_sim = sim_ij, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))
      
      # reef function
      reef_fun_ij <- model_summ(model_sim = sim_ij, location = "reef", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(3, 4, 5))

      # orchard rep
#orch_rep_ij <- model_summ(model_sim = sim_ij, location = "orchard", metric = "production", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))
      
      # orchard cover
    orch_cover_ij <- model_summ(model_sim = sim_ij, location = "orchard", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))
    
    # orchard individuals
    orch_ind_ij <- model_summ(model_sim = sim_ij, location = "orchard", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

      # orchard function
orch_fun_ij <- model_summ(model_sim = sim_ij, location = "orchard", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(3, 4, 5))

      dt_i$reef_cover_mean[j] <- mean(reef_cover_ij[2:max_yr], na.rm = T)
      dt_i$reef_function_mean[j] <- mean(reef_fun_ij[2:max_yr], na.rm = T)
      dt_i$reef_tiles_out_tot[j] <- sum(sim_ij$reef_tiles_out[2:max_yr], na.rm = T)
      dt_i$reef_area_out_tot[j] <- dt_i$reef_tiles_out_tot[j]*5
      dt_i$reef_recruits_out_tot[j] <- sum(sim_ij$reef_out[[1]][[2]][2:max_yr], na.rm = T)
      dt_i$orch_function_mean[j] <- mean(orch_fun_ij[2:max_yr], na.rm = T)
      
    
      
       # variable costs
      # substrate costs
      c_subs_yr <- (sim_ij$reef_tiles_out[2:max_yr] + sim_ij$orchard_tiles_out[2:max_yr]) *(0.5*1 + 0.5*7)*(1 + r_inf)^(2:max_yr) # substrate costs each year
      c_subs <- sum((sim_ij$reef_tiles_out[2:max_yr] + sim_ij$orchard_tiles_out[2:max_yr]) *(0.5*1 + 0.5*7)*(1 + r_inf)^(2:max_yr)) # total substrate costs over time interval
      
      
    # outplanting and orchard maintenance costs
      c_reef <- rep(NA, length(2:max_yr)) # reef
      c_orch <- rep(NA, length(2:max_yr)) # orchard
      c_maint <- rep(NA, length(2:max_yr)) # orchard maintenance
      
      # transplanting costs (ESTIMATED)
      c_trans <- rep(NA, length(2:max_yr)) # transplanting from orchard to reef
      
      for(ii in 2:max_yr){
        #print(ii)
        
 # outplanting costs (assuming you can outplant ~1,250 substrates per day to the reef and 800 per day to the orchard)
        c_reef[ii-1] <- 300*ceiling(sim_ij$reef_tiles_out[ii]/1250) + 200*sim_ij$reef_tiles_out[ii]/1250 + 380/2000*sim_ij$reef_tiles_out[ii]
        
        c_reef[ii-1] <- c_reef[ii-1]*(1 + r_inf)^(ii)
        
       c_orch[ii-1] <- 300*ceiling(sim_ij$orchard_tiles_out[ii]/800) + 200*sim_ij$orchard_tiles_out[ii]/800 + 250/2000 * sim_ij$orchard_tiles_out[ii]
       
       c_orch[ii-1] <- c_orch[ii-1]*(1 + r_inf)^(ii)
        
        # currently occupied reef stars in the orchard:
        RS_ii <- rep(NA, length(orchard_treatments))
        for(kk in 1:length(orchard_treatments)){ # in case there's more than one orchard
          RS_ii[kk] <- sim_ij$orchard_tiles[[kk]][ii]/30 # 30 tiles per reef star
        }
        
        RS_ii <- sum(RS_ii) # filled reef stars in orchard this year
        
        c_maint[ii-1] <- ifelse(RS_ii == 0, 0, (1.5*530-530)/(2*123-123)*RS_ii + 265) 
        c_maint[ii-1] <- c_maint[ii-1]*2*(1 + r_inf)^(ii) # assume maintenance is performed 2x per year, also add inflation
        
        # transplanting costs (ESTIMATED: assume you can collect 1000 corals from the orchard per day and outplant that many to the reef in one more day)
        # trans_colonies_tot
        c_trans[ii-1] <- (300*ceiling(sim_ij$trans_colonies_tot[ii]/1000) + 200*sim_ij$trans_colonies_tot[ii]/1000)*2 # boat costs + diver costs, x2 for collecting from orchard and then outplanting on reef
        
      } # end of loop over 2:max_yr
      
      c_out_yr <- c_reef + c_orch
      c_maint_yr <- c_maint
      c_trans_yr <-  c_trans
      
      c_out <- sum(c_reef) + sum(c_orch)
      c_maint <- sum(c_maint)
      c_trans <-  sum(c_trans)
 
      # lab costs
     # c_lab <- (432*7 + 200*7)*length(2:max_yr) # assume all tiles are only kept in the lab for two weeks
      c_lab <- sum((432*7 + 200*7)*(1 + r_inf)^(2:max_yr))
      c_lab_yr <- (432*7 + 200*7)*(1 + r_inf)^(2:max_yr)
  
      # spawning costs    
      c_spawn <- rep(NA, length(2:max_yr))
      for(ii in 2:max_yr){
        c_spawn[ii-1] <- ifelse(sim_ij$reef_babies_used[ii] > 0, 6500, 0) + ifelse(sim_ij$orchard_babies_used[ii] > 0, 3250, 0)
        c_spawn[ii-1] <- c_spawn[ii-1]*(1 + r_inf)^(ii)
        
      }
     
      c_spawn_yr <- c_spawn
      c_spawn <- sum(c_spawn)
      
      # total costs over length of specified time interval
     tot_costs <- c_fixed + c_subs + c_out + c_maint + c_lab + c_spawn + c_trans
     # total costs each year:
     tot_costs_yr <- rep(c_fixed, length(2:max_yr)) + c_subs_yr + c_out_yr + c_maint_yr + c_lab_yr + c_spawn_yr + c_trans_yr
     
     dt_i$c_fixed[j] <- c_fixed
     dt_i$c_subs[j] <- c_subs
     dt_i$c_out[j] <- c_out
     dt_i$c_maint[j] <- c_maint
     dt_i$c_lab[j] <- c_lab
     dt_i$c_spawn[j] <- c_spawn
     dt_i$c_trans[j] <- c_trans
     dt_i$tot_costs[j] <- tot_costs
      
      #dt_i$tot_dollar_costs[j]
      dt_i$reef_ROI_mean[j] <- mean(reef_cover_ij[2:max_yr]/tot_costs_yr)
      
      dt_i$tot_ROI_mean[j] <- mean((reef_cover_ij[2:max_yr] + orch_cover_ij[2:max_yr])/tot_costs_yr)
      
      dt_i$reefI_ROI_mean[j] <- mean(reef_ind_ij[2:max_yr]/tot_costs_yr)
      
      dt_i$totI_ROI_mean[j] <- mean((reef_ind_ij[2:max_yr] + orch_ind_ij[2:max_yr])/tot_costs_yr)
  
      
  if(annual_costs == F){
    return(dt_i) # return the data frame
  } else{
    return(list(dt_i = dt_i, tot_costs_yr = tot_costs_yr))
  }
  
}



# function that returns mean and error bars for coral cover and orchard function at each time point across demographic parameters
# sim_list = full list of simulations across demographic parameters, orchard outplanting proportions, and an optional second parameter
ts_fun <- function(sim_list, n_pars, max_yr, prop_choice, par2_choice = 1){
  
  # holding matrix for total coral cover at each timepoint (row) and parameter combination (columns)
  reef_cover <- matrix(NA, nrow = max_yr, ncol = n_pars)
  
  # repeat for orchard function
  orch_fun <- matrix(NA, nrow = max_yr, ncol = n_pars)
  
  for(i in 1:n_pars){
    
    sim_i <- sim_list[[i]][[par2_choice]][[prop_choice]]
    
    reef_cover[,i] <- model_summ(model_sim = sim_i, location = "reef", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))[1:max_yr]
    
    orch_fun[,i] <- model_summ(model_sim = sim_i, location = "orchard", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(3, 4, 5))
    
  }
  
  # now calculate the mean, 95%, and 5% quantiles, and standard deviations  across parameter values at each time point
  reef_cover_mn <- apply(reef_cover, 1, function(x) mean(x, na.rm = T))
  reef_cover_low <- apply(reef_cover, 1, function(x) quantile(x, 0.95, na.rm = T))
  reef_cover_up <- apply(reef_cover, 1, function(x) quantile(x, 0.05, na.rm = T))
  
  reef_cover_sd <- apply(reef_cover, 1, function(x) sd(x, na.rm = T))
  
  orch_fun_mn <- apply(orch_fun, 1, function(x) mean(x, na.rm = T))
  orch_fun_low <- apply(orch_fun, 1, function(x) quantile(x, 0.95, na.rm = T))
  orch_fun_up <- apply(orch_fun, 1, function(x) quantile(x, 0.05, na.rm = T))
  orch_fun_sd <- apply(orch_fun, 1, function(x) sd(x, na.rm = T))
  
  
  # return mean, 75% quantile, 25% quantile, and full matrixes for each metric
  
  return(list(reef_cover_mn = reef_cover_mn, reef_cover_low = reef_cover_low, reef_cover_up = reef_cover_up, reef_cover_sd = reef_cover_sd, orch_fun_mn = orch_fun_mn, orch_fun_low = orch_fun_low, orch_fun_up = orch_fun_up, orch_fun_sd = orch_fun_sd, reef_cover_all = reef_cover, orch_fun_all = orch_fun))
  
}




# wrapper function for combining data frames from multiple parameter sets
# sim_list = list with model simulation output for multiple parameters
# max_yr = end of time interval over which to calculate the metrics
# n_pars = number of demographic parameter sets used in sim_list
# prop_set = vector with proportions of substrates to outplant to orchard
# par_set2 = values of the second parameter to iterate over
# par2_name = name of the second parameter iterated over in sim_list
full_dt_fun <- function(sim_list, max_yr, n_pars, prop_set, par_set2, par2_name, annual_costs = F){
  
  costs_list <- list()
  
  for(i in 1:n_pars){
  
    costs_list_i <- list()
    
  for(j in 1:length(par_set2)){
    
    costs_list_j <- list()
    
    for(k in 1:length(prop_set)){
      
      
      if(n_pars == 1){
        sim_ijk <- sim_list[[j]][[k]]
      } else{
        sim_ijk <- sim_list[[i]][[j]][[k]]
      }
      
      
      if(annual_costs == F){
        dt_ijk <- metrics_dt_fun(sim_ijk, max_yr, annual_costs)
      
      dt_ijk$par_rep <- i
      dt_ijk$prop_out <- prop_set[k]
      dt_ijk$par2 <- par_set2[j]
      } else{
        
        metrics_ijk <- metrics_dt_fun(sim_ijk, max_yr, annual_costs)
        
        dt_ijk <- metrics_ijk$dt_i
        
        costs_list_j[[k]] <- metrics_ijk$tot_costs_yr

      }
      
      
      if(k == 1){
        dt_k <- dt_ijk
      } else{
        dt_k <- rbind(dt_k, dt_ijk)
      }
    }
    
    costs_list_i[[j]] <- costs_list_j
    
    if(j == 1){
      dt_j <- dt_k
    } else{
      dt_j <- rbind(dt_j, dt_k)
    }
    
  }
  
  if(i == 1){
    dt_i <- dt_j
  } else{
    dt_i <- rbind(dt_i, dt_j)
  }
  
    
    costs_list[[i]] <- costs_list_i
  }
  
  
  dt_i$par_name2 <- par2_name
  
  if(annual_costs == F){
    return(dt_i)
  } else{
    return(list(dt_i = dt_i, costs_list = costs_list))
  }

}


```


# Figure 2 & SM

Baseline dynamics

```{r}
prop_main <- c(0.5, 1) # orchard investment strategies

lambda_Rs <- c(0, 1255111) # lambda values (combination of lambda_R = 0 and 100% reef investment = no intervention)

# run simulations with lots of different initial conditions, for each determine whether total population size at year 50 is less than the threshold, then for each parameter set determine proportion of runs for which intial conditions meant the population went extinct. 

n_init <- 50 # number of initial conditions to simulate

# holding matrices for coral cover after 50 years
N0_50 <- matrix(NA, nrow = n_sample1, ncol = n_init) # row = parameter rep, col = initial conditions
N0_100 <- matrix(NA, nrow = n_sample1, ncol = n_init)
N0_0 <- matrix(NA, nrow = n_sample1, ncol = n_init)

tic()

set.seed(1000)
for(i in 1:n_init){
N0.r <- list()
N0.r[[1]] <- list() # second reef subpop
N0.r[[1]][[1]] <- rep(0, n) # first source (external recruits)
N0.r[[1]][[2]] <- sample(c(0:20), 5, replace = T) # second source
N0.r[[1]][[3]] <- rep(0, n) # third source

# add disturbance with stochastic frequencies and severities; consider the initial conditions to be part of this (since they would depend on past disturbances) so don't do this factorially with the initial conditions 
n_dist_i <- sample(c(1:round(years/2)), 1, replace = F) # number of disturbances in simulation (max possible is avg of every other year)
dist_yrs_i <- sort(sample(c(1:years), n_dist_i, replace = F)) # years when the disturbances occur

# severity of disturbance on reef
dist_sev.r <- runif(1, min = 0.001, max = 1)
dist_sev.o <- min(2*dist_sev.r, 1)

dist_mat.r <- matrix(dist_sev.r, nrow = length(dist_yrs_i), ncol = 5)
dist.r <- list(dist_mat.r, dist_mat.r, dist_mat.r) # disturbance impacts on survival for each source on the reef

dist_mat.o <- matrix(dist_sev.o, nrow = length(dist_yrs_i), ncol = 5)
dist.o <- list(dist_mat.o, dist_mat.o) # disturbance impacts on survival for each source on the orchard

dist_pars_list_i <- list(dist.r = dist.r, dist.o = dist.o)

  
 # 50% investment in reef
orch_D0_N0_50 <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = prop_main[1], par_set2 = lambda_Rs[2], par2_name = "lambda_R", dist = T, dist_yrs = dist_yrs_i, dist_pars_list = dist_pars_list_i)

# 100% investment in reef
orch_D0_N0_100 <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = prop_main[2], par_set2 = lambda_Rs[2], par2_name = "lambda_R", dist = T, dist_yrs = dist_yrs_i, dist_pars_list = dist_pars_list_i)

# no restoration
orch_D0_N0_0 <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = prop_main[2], par_set2 = lambda_Rs[1], par2_name = "lambda_R", dist = T, dist_yrs = dist_yrs_i, dist_pars_list = dist_pars_list_i)

for(j in 1:n_sample1){
    # for each parameter set, calculate number of corals after 50 years
    
    N0_50[j,i] <- model_summ(model_sim = orch_D0_N0_50[[j]][[1]][[1]], location = "reef", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(1:5))[50]
    
    N0_100[j,i] <- model_summ(model_sim = orch_D0_N0_100[[j]][[1]][[1]], location = "reef", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(1:5))[50]
    
    N0_0[j,i] <- model_summ(model_sim = orch_D0_N0_0[[j]][[1]][[1]], location = "reef", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(1:5))[50]
    
    
  }
}

toc()

# reset initial conditions
N0.r <- list()
N0.r[[1]] <- list() # second reef subpop
N0.r[[1]][[1]] <- rep(0, n) # first source (external recruits)
N0.r[[1]][[2]] <- rep(0, n) # second source
N0.r[[1]][[3]] <- rep(0, n) # third source


```

139.805 sec elapsed

```{r}
ext_thresh <- 100 # 100 m2 of coral

# for each disturbance regime, calculate proportion of parameter combinations the population fell below the extinction threshold

# ext_N0_50 <- rep(NA, n_init)
# ext_N0_100 <- rep(NA, n_init)
# ext_N0_0 <- rep(NA, n_init)
# 
# for(i in 1:n_init){
#   ext_N0_50[i] <- 1-length(which(N0_50[,i] <= ext_thresh))/length(N0_50[,i])
#   ext_N0_100[i] <- 1-length(which(N0_100[,i] <= ext_thresh))/length(N0_100[,i])
#   ext_N0_0[i] <- 1-length(which(N0_0[,i] <= ext_thresh))/length(N0_0[,i])
# }

ext_N0_50 <- length(which(as.vector(N0_50)< ext_thresh))/length(as.vector(N0_50))
ext_N0_100 <- length(which(as.vector(N0_100)< ext_thresh))/length(as.vector(N0_100))
ext_N0_0 <- length(which(as.vector(N0_0)< ext_thresh))/length(as.vector(N0_0))

ext_N0_50
ext_N0_100
ext_N0_0
```

```{r}
# reef cover for the three restoration levels, no disturbance

N0.r <- list()
N0.r[[1]] <- list() # second reef subpop
N0.r[[1]][[1]] <- rep(0, n) # first source (external recruits)
N0.r[[1]][[2]] <- rep(10, n) # second source. Start with some corals
N0.r[[1]][[3]] <- rep(0, n) # third source


orch_D0_50 <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = prop_main[1], par_set2 = lambda_Rs[2], par2_name = "lambda_R", dist = F, dist_yrs = NULL, dist_pars_list = NULL)

orch_D0_100 <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = prop_main[2], par_set2 = lambda_Rs[2], par2_name = "lambda_R", dist = F, dist_yrs = NULL, dist_pars_list = NULL)

orch_D0_0 <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = prop_main[2], par_set2 = lambda_Rs[1], par2_name = "lambda_R", dist = F, dist_yrs = NULL, dist_pars_list = NULL)

# reset initial conditions to default
N0.r <- list()
N0.r[[1]] <- list() # second reef subpop
N0.r[[1]][[1]] <- rep(0, n) # first source (external recruits)
N0.r[[1]][[2]] <- rep(0, n) # second source. Start with some corals
N0.r[[1]][[3]] <- rep(0, n) # third source


```


```{r}
# disturbance heatmap
# just use mean parameter values for now and save the full parameter combinations for the disturbance severity SM figure

# here assume 100% outplanting to reef (looking more closely at disturbance impacts with restoration)

n_dist_sims <- 25

dist_yrs_list <- list() 
dist_yrs_list[[1]] <- c(1:max(years))
dist_yrs_list[[2]] <- c(1:max(years))[c(F, T)]
dist_yrs_list[[3]] <- c(1:max(years))[c(F, F, T)]

for(i in 4:n_dist_sims){
  
  dist_yrs_list[[i]] <- c(1:max(years))[c(c(F, F, T), rep(F, i-3))]
}


# could show coral cover relative to default?

dist_sev.r <- rev(seq(from = 0.001, to = 1, length.out = n_dist_sims))
dist_sev.o <- apply(matrix(c(2*dist_sev.r, rep(1, 25)), ncol = 2, byrow = F), 1, function(x) min(x))
dist_sev.o1 <- dist_sev.o


dist_sims <- list()

tic()
for(i in 1:n_dist_sims){ # severity
  
  dist_sims_j <- list()
  
  for(j in 1:n_dist_sims){ # frequency
    
  dist_mat.r <- matrix(dist_sev.r[i], nrow = length(dist_yrs_list[[j]]), ncol = 5)
dist.r <- list(dist_mat.r, dist_mat.r, dist_mat.r) # disturbance impacts on survival for each source on the reef

dist_mat.o <- matrix(dist_sev.o[i], nrow = length(dist_yrs_list[[j]]), ncol = 5)
dist.o <- list(dist_mat.o, dist_mat.o) # disturbance impacts on survival for each source on the orchard

dist_pars_list_j <- list(dist.r = dist.r, dist.o = dist.o)

dist_sims_j[[j]] <- orch_exp_fun1(dem_pars = all_pars, prop_set = prop_main[2], par_set2 = NULL, par2_name = "none", dist = T, dist_yrs = dist_yrs_list[[j]], dist_pars_list = dist_pars_list_j)

  }

dist_sims[[i]] <- dist_sims_j

}

toc()


# process the results
#dist_dt_fun2 <- function(dist_sims, dist_sev.r, dist_sev.o, max_yr = 51, n_pars = n_sample1, prop_set = prop_main, par_set2 = NA, par2_name = "none"){
  
  for(i in 1:n_dist_sims){ # for each severity
    
    for(j in 1:n_dist_sims){ # for each frequncy
      
      dt_j <- full_dt_fun(sim_list = dist_sims[[i]][[j]], max_yr = 51, n_pars = 1, prop_set = prop_main[2], par_set2 = NA, par2_name = "none")
  
   dt_j$dist_sev_r <- dist_sev.r[i]
   dt_j$dist_freq <- dist_yrs_list[[j]][2] - dist_yrs_list[[j]][1]
      
    if(j == 1){
     dt_i <- dt_j
   } else{
     dt_i <- rbind(dt_i, dt_j)
   }
   
    }
  
  

   if(i == 1){
     dt_out <- dt_i
   } else{
     dt_out <- rbind(dt_out, dt_i)
   }
  
  }
 
#  return(dt_out)
#}

dt_dist <- dt_out

#View(dt_out)

```



Make the plots: 
2A: Probability of the population falling below the quasi-extinction threshold within 50 years, summarized across 100 parameter sets. Three bars: no intervention, 50% to reef, 100% to reef. 

2B: Reef cover (m²) over 50 years, three restoration levels matching 2A, no disturbance. Mean line + 90% quantile envelope from 100 parameter replicates. Maybe make this panel A instead?

2C: Heatmap of reef cover at year 50, severity (x) × frequency (y). Second contour layer shows CV across replicates so reviewers can see the ranking is stable under parameter uncertainty. Finding: severity dominates frequency over the 3–7 year window.


```{r}
s_cols <- c("#888888", "#FB9A29", "#993404")


pdf("TempFigs/Figure2.pdf", height = 3, width = 9)

par(mfrow = c(2, 1))
layout(matrix(c(1, 2, 3, 4), nrow = 1, ncol = 4, byrow = T), widths = c(1, 1, 1, 0.2), heights = rep(1, 4))
#layout.show(4)
par(mar=c(0, 3.5, 2.5, 1), oma = c(4.5, 1.5, 0.1, 0))


# panel A: time series

lab_cx <- 0.85

xlab1 <- "Year in simulation"
ylab1 <- expression("Reef cover ("*m^2*")")
#plotlab <- "c) Reef ROI"
plot(x = c(1:years), y = rep(0, years), type = "l", col = NA, ylim = c(0, 4000), las = 1, xlab = NA, ylab = NA, xaxt = "n")
axis(side = 1, at = seq(0, 50, 10), labels = seq(0, 50, 10) + 2025)
mtext(side = 1, xlab1, line = 2.5, cex = lab_cx)
mtext(side = 2, ylab1, line = 2.9, cex = lab_cx)
mtext(side = 3, "a) Reef cover over 50 years", cex = lab_cx, adj = 0)

  plot_dt <- ts_fun(sim_list = orch_D0_0, n_pars = n_sample1, max_yr = years, prop_choice = 1, par2_choice = 1)
  lines(x = c(1:years), y = plot_dt$reef_cover_mn, col = s_cols[1], lwd = 2)
  polygon(x = c(c(1:years), rev(c(1:years))), y = c(plot_dt$reef_cover_low, rev(plot_dt$reef_cover_up)), col = adjustcolor(s_cols[1], alpha.f = 0.2), border = NA)
  
  plot_dt <- ts_fun(sim_list = orch_D0_50, n_pars = n_sample1, max_yr = years, prop_choice = 1, par2_choice = 1)
  lines(x = c(1:years), y = plot_dt$reef_cover_mn, col = s_cols[2], lwd = 2)
  polygon(x = c(c(1:years), rev(c(1:years))), y = c(plot_dt$reef_cover_low, rev(plot_dt$reef_cover_up)), col = adjustcolor(s_cols[2], alpha.f = 0.2), border = NA)
  
  plot_dt <- ts_fun(sim_list = orch_D0_100, n_pars = n_sample1, max_yr = years, prop_choice = 1, par2_choice = 1)
  lines(x = c(1:years), y = plot_dt$reef_cover_mn, col = s_cols[3], lwd = 2)
  polygon(x = c(c(1:years), rev(c(1:years))), y = c(plot_dt$reef_cover_low, rev(plot_dt$reef_cover_up)), col = adjustcolor(s_cols[3], alpha.f = 0.2), border = NA)
  
legend("topleft", legend = c("none", "50% to reef", "100% to reef"), col = s_cols, lwd = c(2, 2, 2), lty = c(1, 1, 1), title = "Restoration strategy", bty = "n", cex = 1.1, fill = c(adjustcolor(s_cols[1], alpha.f = 0.2), adjustcolor(s_cols[2], alpha.f = 0.2), adjustcolor(s_cols[3], alpha.f = 0.2)), border = NA, merge = T, seg.len = rep(0.5, 3))


# panel B: extinction probabilities
ext_probs <- c(ext_N0_0, ext_N0_50, ext_N0_100)

plot(x = c(1, 2, 3), y = rep(0, 3), col = NA, ylim = c(0, 1), xlim = c(0.5, 3.5), las = 1, xaxt = "n", xlab = NA, ylab = NA, xaxs = "i")
mtext(side = 1, "Restoration strategy", line = 2.5, cex = lab_cx)
#mtext(side = 2, "Prop. simulations below \nextinction threshold", line = 2.5, cex = lab_cx)
mtext(side = 2, "Extinction probability", line = 2.5, cex = lab_cx)
mtext(side = 3, "b) Probabilities of extinction", cex = lab_cx, adj = 0)
axis(side = 1, at = c(1,2,3), labels = c("No \nintervention", "50% reef \ninvestment", "100% reef \ninvestment"), tick = F, cex.axis = 0.9)
for(i in 1:3){
  polygon(x = c(i-0.25, i + 0.25, i + 0.25, i-0.25), y = c(0, 0, ext_probs[i], ext_probs[i]), col = s_cols[i])

}

# panel C: disturbance sensitivity
  
xset <- unique(dt_dist$dist_sev_r)
yset <- unique(dt_dist$dist_freq)

dt_plot <- dt_dist
# color palette
nHalf <- 100
rc1 <- colorRampPalette(colors = c("#E8E6E6", "#993404"), space = "Lab")(nHalf)
rampcols1 <- c(rc1) # all the colors

col_vals <- seq(from = 0, to = max(dt_plot$reef_cover_mean), length.out = nHalf)

  L_HM <- n_dist_sims
  
  #par(pty = "s")
plot(x = xset*L_HM, y = yset*L_HM, col = NA, xlim = c(0, L_HM), ylim = c(0, L_HM), las = 1, xaxs = "i", yaxs = "i", xlab = NA, ylab = NA, xaxt = "n", yaxt = "n") # asp = 1,xaxs = "i" removes white space around the plot
axis(side = 1, at = seq(1, L_HM, length.out = 6), tick = F, line = -0.5, labels = seq(0,1,0.2), cex.axis = 1, padj = -1)
axis(side = 2, at = seq(1, L_HM, 4), tick = F, line = -0.5, cex.axis = 1, las = 1, labels = rev(seq(1, L_HM, 4)))
mtext(side = 1, "Disturbance severity \n(prop. reduction in survival)", line = 2.75, cex = lab_cx)
mtext(side = 2, "Disturbance frequency (years \nbetween disturbance events)", line = 1.75, cex = lab_cx)
mtext(side = 3, "c) Senstivity to disturbance", cex = lab_cx, adj = 0)
for(i in 1:L_HM){ # x value
  for(j in 1:L_HM){ # y value
    
    # get point to plot
    plot_pt <- dt_plot$reef_cover_mean[which(dt_plot$dist_sev_r==xset[i] & dt_plot$dist_freq == yset[j])]
    
    # reverse y-axis
    j2 <- rev(c(1:L_HM))[j]
    
    # get closest color
    if(plot_pt < 1){
      index_ij <- which(abs(plot_pt-col_vals) == min(abs(plot_pt-col_vals)))
      
      polygon(x = c(i-1, i, i, i-1), y = c(j2-1, j2-1, j2, j2), col = rampcols1[index_ij], border = NA) 
    
    } else{
      index_ij <- which(abs(plot_pt-col_vals) == min(abs(plot_pt-col_vals)))
      
      polygon(x = c(i-1, i, i, i-1), y = c(j2-1, j2-1, j2, j2), col = rampcols1[index_ij], border = NA) 
    }
    
  }
}


# make the legend
xl <- 1
yb <- 1
xr <- 1.25
yt <- 2

par(mar=c(0, 0,2.5,0))
plot(NA, type = "n", ann = F, xlim = c(1, 2), ylim = c(1, 2), xaxt = "n", yaxt = "n", bty = "n")
mtext(side = 3, expression(m^2~"coral"), cex = 0.8, adj = -0.025)
text(x = 1.6, y = 1, "0")
text(x = 1.6, y = 1.5, "250")
text(x = 1.6, y = 2-0.025, "500") # max(dt_plot$reef_cover_mean)
rect(
  xl,
  head(seq(yb, yt, (yt-yb)/nHalf), -1), # the denominator needs to match the length of the color vector
  xr,
  tail(seq(yb, yt, (yt - yb)/nHalf), -1),
  col = rampcols1,
  border = NA
)


dev.off()

```


SM figure: λ distribution + parameter distributions
-- for parameter distribution: show histogram of survival values, with different color for different size classes


lambda distributions

```{r}
# calculate lambda for each of the parameter sets

for(i in 1:n_sample1){
  lambda_set[i] <- pop_lambda_fun(all_pars_R$surv_pars_L.r[[i]][[1]][[1]], all_pars_R$growth_pars_L.r[[i]][[1]][[1]], all_pars_R$shrink_pars_L.r[[i]][[1]][[1]], all_pars_R$frag_pars_L.r[[i]][[1]][[1]])
}


# get survival parameters for each size class and recruits

surv_mat <- matrix(NA, nrow = n_sample1, ncol = 6)

#s1_rand <- rep(NA, n_sample1)

for(i in 1:n_sample1){
  #s1_rand[i] <- all_pars_R$surv_pars_L.rc[[i]][[1]][[1]] 
  
  surv_mat[i,1] <- all_pars_R$surv_pars_L.rc[[i]][[1]][[1]] 
  
  for(k in 1:5){
    surv_mat[i,k+1] <- all_pars_R$surv_pars_L.r[[i]][[1]][[1]][k]
  }
}

#unique(s1_rand)

size_cols <- c("#882E72", "#5289C7", "#4EB265", "#F7F056", "#F1932D", "#DC050C")

a1 <- 0.3

pdf("TempFigs/SM_par_hist.pdf", height = 5, width = 4)

par(mfrow = c(2, 1))
layout(matrix(c(1, 2), nrow = 2, ncol = 1, byrow = T), widths = c(1), heights = rep(1, 2))
#layout.show(2)
par(mar=c(2.5, 3.5, 2.5, 1), oma = c(0.75, 0.25, 0, 0))


hist(lambda_set, main = NA, xlab = NA, ylab = NA, las = 1, breaks = 15)
mtext(side = 1, expression("Population"~lambda), line = 2.25)
mtext(side = 2, "Frequency", line = 2)
mtext(side = 3, "a) Population growth rates", line = 0, adj = 0)
abline(v = 1, lty = 2)

hist(surv_mat[,1], las = 1, xlab = NA, ylab = NA, main = NA, xlim = c(0,1), breaks = 5, col = adjustcolor(size_cols[1], alpha.f = a1), border = size_cols[1]) # xaxt = "n",, breaks = 20
#axis(side = 1, at = c(0 + 0.01, 0.05, 0.1-0.01, 0.15), labels = sort(unique(s1_rand)))
mtext(side = 1, "Proportion surviving one year", line = 2)
mtext(side = 2, "Frequency", line = 2)
mtext(side = 3, "b) Survival rates by size class", adj = 0)
hist(surv_mat[,2], las = 1, xlab = NA, main = NA, xlim = c(0,1), add = T, breaks = 20, col = adjustcolor(size_cols[2], alpha.f = a1), border = size_cols[2])
hist(surv_mat[,3], las = 1, xlab = NA, main = NA, xlim = c(0,1), add = T, breaks = 10, col = adjustcolor(size_cols[3], alpha.f = a1), border = size_cols[3])
hist(surv_mat[,4], las = 1, xlab = NA, main = NA, xlim = c(0,1), add = T, breaks = 5, col = adjustcolor(size_cols[4], alpha.f = a1), border = size_cols[4])
hist(surv_mat[,5], las = 1, xlab = NA, main = NA, xlim = c(0,1), add = T, breaks = 5, col = adjustcolor(size_cols[5], alpha.f = a1), border = size_cols[5])
hist(surv_mat[,6], las = 1, xlab = NA, main = NA, xlim = c(0,1), add = T, breaks = 2, col = adjustcolor(size_cols[6], alpha.f = a1), border = size_cols[6])
legend("topleft", legend = c("recruits", "1", "2", "3", "4", "5"), fill = c(adjustcolor(size_cols[1], alpha.f = a1), adjustcolor(size_cols[2], alpha.f = a1), adjustcolor(size_cols[3], alpha.f = a1), adjustcolor(size_cols[4], alpha.f = a1), adjustcolor(size_cols[5], alpha.f = a1), adjustcolor(size_cols[6], alpha.f = a1)), ncol = 2, title = "Size class", inset = c(0.1,0.05), bty = "n", cex = 0.75)
abline(v = apply(surv_mat, MARGIN = 2, FUN = mean), col = size_cols, lwd = 1.5)

dev.off()

#min(surv_mat[,1])
#max(surv_mat[,1])

```


# Figure 3 & SM

petal simulations

```{r}

orch_D0_petals <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = prop_main, par_set2 = lambda_Rs[2], par2_name = "lambda_R", dist = F, dist_yrs = NULL, dist_pars_list = NULL)

# process output
orch_D0_petals_dt5 <- full_dt_fun(sim_list = orch_D0_petals, max_yr = 6, n_pars = n_sample1, prop_set = prop_main, par_set2 = lambda_Rs[2], par2_name = "lambda_R")

orch_D0_petals_dt50 <- full_dt_fun(sim_list = orch_D0_petals, max_yr = 51, n_pars = n_sample1, prop_set = prop_main, par_set2 = lambda_Rs[2], par2_name = "lambda_R")

```

petal function

```{r}

# make a function for making the OHI-type plots

#' OHI-style radar plot
#' Creates a radar (spider) plot showing normalized performance metrics for
#' each orchard-to-reef allocation proportion. Highlights the optimal proportion.
petal_fun <- function(dt_plot, prop_choice){
  
#metric_cols <- c("#4477AA", "#AA3377", "#66CCEE", "#CCBB44","#EE6677", "#228833") # coral cover on reef, orchard function, reef ROI, total ROI, total costs, recruits outplanted to reef

metric_cols <- c("#99DDFF", "#FFAABB", "#AAAA00", "#BBCC33","#EE8866", "#44BB99") # coral cover on reef, orchard function, reef ROI, total ROI, total costs, recruits outplanted to reef

Lcol <- "gray"
lab_cex <- 0.9
a1 <- 0.6

lwd_ci <- 2
#View(dt_plot)
  
#dt_plot2 <- dt_plot %>% filter(prop_out == prop_choice) %>% left_join(ref_vals, by = c("par_rep", "par2"))

#dt_plot2 <- dt_plot %>% filter(prop_out == prop_choice)

dt_plot1 <- dt_plot %>% select(par_rep, prop_out, reef_cover_mean,orch_function_mean,reef_ROI_mean,tot_ROI_mean,tot_costs,reef_recruits_out_tot) %>% group_by(par_rep) %>% mutate(reef_cover_max = max(reef_cover_mean), orch_function_max = max(orch_function_mean), reef_ROI_max = max(reef_ROI_mean), tot_ROI_max = max(tot_ROI_mean), tot_costs_max = max(tot_costs), reef_recruits_out_max = max(reef_recruits_out_tot)) %>% ungroup()

dt_plot2 <- dt_plot1 %>% filter(prop_out == prop_choice)

par(pty = "s")
r_in <- 0.5
r_out0 <- r_in + 1
test_x <- seq(from = -r_in, to = r_in, length.out = 100)
plot(x = 1, y = 1, col = NA, xlim = c(-5, 5), ylim = c(-5, 5), asp = 1, las = 1, xlab = NA, ylab = NA, xaxt = "n", yaxt = "n", bty = "n") # , xaxs = "i", yaxs = "i"
polygon(x = c(test_x, rev(test_x)), y = c(sqrt((r_in)^2 - test_x^2), rev(-1*sqrt(r_in^2 - test_x^2))), border = Lcol) # , col = "gray"
test_x <- seq(from = -r_out0, to = r_out0, length.out = 100)
polygon(x = c(test_x, rev(test_x)), y = c(sqrt((r_out0)^2 - test_x^2), rev(-1*sqrt(r_out0^2 - test_x^2))), border = Lcol)
#text(x = 0, y = 0, paste(as.character(round(opt_prop1*100,0)), "%", sep = ""))
#text(x = 0, y = 0, paste(as.character(prop_choice), as.character(0.9), sep = "/")) 
text(x = 0, y = 0, as.character(prop_choice)) 

# add background lines
th1_set <- seq(from = 0, to = 300, by = 60)
th2_set <- seq(from = 60, to = 360, by = 60)
for(i in 1:6){
  r_in_x <- seq(from = r_in*cos(th1_set[i]*pi/180), to = r_in*cos(th2_set[i]*pi/180), length.out = 100)
r_out_x <- seq(from = r_out0*cos(th1_set[i]*pi/180), to = r_out0*cos(th2_set[i]*pi/180), length.out = 100)

if(i < 4){
  polygon(x = c(r_in_x, rev(r_out_x)), y = c(sqrt(r_in^2 - r_in_x^2), rev(sqrt(r_out0^2 - r_out_x^2))), border = Lcol)
} else{
  polygon(x = c(r_in_x, rev(r_out_x)), y = c(-sqrt(r_in^2 - r_in_x^2), rev(-sqrt(r_out0^2 - r_out_x^2))), border = Lcol)
}

}

# value to plot: on average, how far was each metric from it's max possible value for that parameter set? (outer ring is one); actual value to show in text is the average of the raw values at the proportion outplanted?

# reef_cover_mean
r_out <- mean(dt_plot2$reef_cover_mean/dt_plot2$reef_cover_max) + r_in
th1 <- 0 # starting angle (in degrees)
th2 <- 60 # ending angle (in degrees)
r_in_x <- seq(from = r_in*cos(th1*pi/180), to = r_in*cos(th2*pi/180), length.out = 100)
r_out_x <- seq(from = r_out*cos(th1*pi/180), to = r_out*cos(th2*pi/180), length.out = 100)
polygon(x = c(r_in_x, rev(r_out_x)), y = c(sqrt(r_in^2 - r_in_x^2), rev(sqrt(r_out^2 - r_out_x^2))), col = adjustcolor(metric_cols[1], alpha.f = a1), border = NA)
# add 90% confidence intervals
r_low <- quantile(dt_plot2$reef_cover_mean/dt_plot2$reef_cover_max, 0.05) + r_in
r_up <- quantile(dt_plot2$reef_cover_mean/dt_plot2$reef_cover_max, 0.95) + r_in
r_low_x <- seq(from = r_low*cos(th1*pi/180), to = r_low*cos(th2*pi/180), length.out = 100)
r_up_x <- seq(from = r_up*cos(th1*pi/180), to = r_up*cos(th2*pi/180), length.out = 100)
lines(x = r_low_x, y = sqrt(r_low^2 - r_low_x^2), col = metric_cols[1], lwd = lwd_ci)
lines(x = r_up_x, y = sqrt(r_up^2 - r_up_x^2), col = metric_cols[1], lwd = lwd_ci)

# orch_function_mean
r_out <- mean(dt_plot2$orch_function_mean/dt_plot2$orch_function_max) + r_in
th1 <- 60 # starting angle (in degrees)
th2 <- 120 # ending angle (in degrees)
r_in_x <- seq(from = r_in*cos(th1*pi/180), to = r_in*cos(th2*pi/180), length.out = 100)
r_out_x <- seq(from = r_out*cos(th1*pi/180), to = r_out*cos(th2*pi/180), length.out = 100)
polygon(x = c(r_in_x, rev(r_out_x)), y = c(sqrt(r_in^2 - r_in_x^2), rev(sqrt(r_out^2 - r_out_x^2))), col = adjustcolor(metric_cols[2], alpha.f = a1), border = NA)
# add 90% confidence intervals
r_low <- quantile(dt_plot2$orch_function_mean/dt_plot2$orch_function_max, 0.05) + r_in
r_up <- quantile(dt_plot2$orch_function_mean/dt_plot2$orch_function_max, 0.95) + r_in
r_low_x <- seq(from = r_low*cos(th1*pi/180), to = r_low*cos(th2*pi/180), length.out = 100)
r_up_x <- seq(from = r_up*cos(th1*pi/180), to = r_up*cos(th2*pi/180), length.out = 100)
lines(x = r_low_x, y = sqrt(r_low^2 - r_low_x^2), col = metric_cols[2], lwd = lwd_ci)
lines(x = r_up_x, y = sqrt(r_up^2 - r_up_x^2), col = metric_cols[2], lwd = lwd_ci)


# reef_ROI_mean
r_out <- mean(dt_plot2$reef_ROI_mean/dt_plot2$reef_ROI_max) + r_in
th1 <- 120 # starting angle (in degrees)
th2 <- 180 # ending angle (in degrees)
r_in_x <- seq(from = r_in*cos(th1*pi/180), to = r_in*cos(th2*pi/180), length.out = 100)
r_out_x <- seq(from = r_out*cos(th1*pi/180), to = r_out*cos(th2*pi/180), length.out = 100)
polygon(x = c(r_in_x, rev(r_out_x)), y = c(sqrt(r_in^2 - r_in_x^2), rev(sqrt(r_out^2 - r_out_x^2))), col = adjustcolor(metric_cols[3], alpha.f = a1), border = NA)
# add 90% confidence intervals
r_low <- quantile(dt_plot2$reef_ROI_mean/dt_plot2$reef_ROI_max, 0.05) + r_in
r_up <- quantile(dt_plot2$reef_ROI_mean/dt_plot2$reef_ROI_max, 0.95) + r_in
r_low_x <- seq(from = r_low*cos(th1*pi/180), to = r_low*cos(th2*pi/180), length.out = 100)
r_up_x <- seq(from = r_up*cos(th1*pi/180), to = r_up*cos(th2*pi/180), length.out = 100)
lines(x = r_low_x, y = sqrt(r_low^2 - r_low_x^2), col = metric_cols[3], lwd = lwd_ci)
lines(x = r_up_x, y = sqrt(r_up^2 - r_up_x^2), col = metric_cols[3], lwd = lwd_ci)


# past 180: need to switch to negative square root
# tot_ROI_mean
r_out <- mean(dt_plot2$tot_ROI_mean/dt_plot2$tot_ROI_max) + r_in
th1 <- 180 # starting angle (in degrees)
th2 <- 240 # ending angle (in degrees)
r_in_x <- seq(from = r_in*cos(th1*pi/180), to = r_in*cos(th2*pi/180), length.out = 100)
r_out_x <- seq(from = r_out*cos(th1*pi/180), to = r_out*cos(th2*pi/180), length.out = 100)
polygon(x = c(r_in_x, rev(r_out_x)), y = c(-sqrt(r_in^2 - r_in_x^2), rev(-sqrt(r_out^2 - r_out_x^2))), col = adjustcolor(metric_cols[4], alpha.f = a1), border = NA)
# add 90% confidence intervals
r_low <- quantile(dt_plot2$tot_ROI_mean/dt_plot2$tot_ROI_max, 0.05) + r_in
r_up <- quantile(dt_plot2$tot_ROI_mean/dt_plot2$tot_ROI_max, 0.95) + r_in
r_low_x <- seq(from = r_low*cos(th1*pi/180), to = r_low*cos(th2*pi/180), length.out = 100)
r_up_x <- seq(from = r_up*cos(th1*pi/180), to = r_up*cos(th2*pi/180), length.out = 100)
lines(x = r_low_x, y = -sqrt(r_low^2 - r_low_x^2), col = metric_cols[4], lwd = lwd_ci)
lines(x = r_up_x, y = -sqrt(r_up^2 - r_up_x^2), col = metric_cols[4], lwd = lwd_ci)



# tot_costs
# r_out <- 1- (mean(dt_plot2$tot_costs/dt_plot2$tot_costs_min))^(-1) + r_in
r_out <- mean(dt_plot2$tot_costs/dt_plot2$tot_costs_max) + r_in
th1 <- 240 # starting angle (in degrees)
th2 <- 300 # ending angle (in degrees)
r_in_x <- seq(from = r_in*cos(th1*pi/180), to = r_in*cos(th2*pi/180), length.out = 100)
r_out_x <- seq(from = r_out*cos(th1*pi/180), to = r_out*cos(th2*pi/180), length.out = 100)
polygon(x = c(r_in_x, rev(r_out_x)), y = c(-sqrt(r_in^2 - r_in_x^2), rev(-sqrt(r_out^2 - r_out_x^2))), col = adjustcolor(metric_cols[5], alpha.f = a1), border = NA)
# add 90% confidence intervals
r_low <- quantile(dt_plot2$tot_costs/dt_plot2$tot_costs_max, 0.05) + r_in
r_up <- quantile(dt_plot2$tot_costs/dt_plot2$tot_costs_max, 0.95) + r_in
r_low_x <- seq(from = r_low*cos(th1*pi/180), to = r_low*cos(th2*pi/180), length.out = 100)
r_up_x <- seq(from = r_up*cos(th1*pi/180), to = r_up*cos(th2*pi/180), length.out = 100)
lines(x = r_low_x, y = -sqrt(r_low^2 - r_low_x^2), col = metric_cols[5], lwd = lwd_ci)
lines(x = r_up_x, y = -sqrt(r_up^2 - r_up_x^2), col = metric_cols[5], lwd = lwd_ci)



# reef_recruits_out_tot
r_out <- mean(dt_plot2$reef_recruits_out_tot/dt_plot2$reef_recruits_out_max) + r_in
th1 <- 300 # starting angle (in degrees)
th2 <- 360 # ending angle (in degrees)
r_in_x <- seq(from = r_in*cos(th1*pi/180), to = r_in*cos(th2*pi/180), length.out = 100)
r_out_x <- seq(from = r_out*cos(th1*pi/180), to = r_out*cos(th2*pi/180), length.out = 100)
polygon(x = c(r_in_x, rev(r_out_x)), y = c(-sqrt(r_in^2 - r_in_x^2), rev(-sqrt(r_out^2 - r_out_x^2))), col = adjustcolor(metric_cols[6], alpha.f = a1), border = NA)
# add 90% confidence intervals
r_low <- quantile(dt_plot2$reef_recruits_out_tot/dt_plot2$reef_recruits_out_max, 0.05) + r_in
r_up <- quantile(dt_plot2$reef_recruits_out_tot/dt_plot2$reef_recruits_out_max, 0.95) + r_in
r_low_x <- seq(from = r_low*cos(th1*pi/180), to = r_low*cos(th2*pi/180), length.out = 100)
r_up_x <- seq(from = r_up*cos(th1*pi/180), to = r_up*cos(th2*pi/180), length.out = 100)
lines(x = r_low_x, y = -sqrt(r_low^2 - r_low_x^2), col = metric_cols[6], lwd = lwd_ci)
lines(x = r_up_x, y = -sqrt(r_up^2 - r_up_x^2), col = metric_cols[6], lwd = lwd_ci)


}

```


outplanting proportion simulations

```{r}
# no disturbance


# severe disturbance
```



3A Petal plots: 50% vs. 100% to reef. Add error bars or annotate the to show uncertainty across parameter sets


```{r}

pdf("TempFigs/petal_5_D0_05.pdf", height = 8, width = 8)
petal_fun(dt_plot = orch_D0_petals_dt5, prop_choice = 0.5)
dev.off()

pdf("TempFigs/petal_5_D0_1.pdf", height = 8, width = 8)
petal_fun(dt_plot = orch_D0_petals_dt5, prop_choice = 1)
dev.off()

pdf("TempFigs/petal_50_D0_05.pdf", height = 8, width = 8)
petal_fun(dt_plot = orch_D0_petals_dt50, prop_choice = 0.5)
dev.off()

pdf("TempFigs/petal_50_D0_1.pdf", height = 8, width = 8)
petal_fun(dt_plot = orch_D0_petals_dt50, prop_choice = 1)
dev.off()

```
```{r}
#prop_all <- seq(from = 0, to = 1, length.out = 20) # main strategies to focus on

prop_all <- seq(from = 0, to = 1, length.out = 20)

lambda_Rs <- c(0, 1255111)

#lab_max <- 12000

#orch_D0 <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = prop_main, par_set2 = NULL, par2_name = "none", dist = F, dist_yrs = NULL, dist_pars_list = NULL)

tic()
orch_D0_all <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = prop_all, par_set2 = lambda_Rs, par2_name = "lambda_R", dist = F, dist_yrs = NULL, dist_pars_list = NULL)

orch_D3_all <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = prop_all, par_set2 = lambda_Rs, par2_name = "lambda_R", dist = T, dist_yrs = dist_yrs3, dist_pars_list = dist_pars_list3)
toc()

# process the output

orch_D0_all_dt5 <- full_dt_fun(sim_list = orch_D0_all, max_yr = 6, n_pars = n_sample1, prop_set = prop_all, par_set2 = lambda_Rs, par2_name = "lambda_R")
orch_D0_all_dt50 <- full_dt_fun(sim_list = orch_D0_all, max_yr = 51, n_pars = n_sample1, prop_set = prop_all, par_set2 = lambda_Rs, par2_name = "lambda_R")

orch_D3_all_dt5 <- full_dt_fun(sim_list = orch_D3_all, max_yr = 6, n_pars = n_sample1, prop_set = prop_all, par_set2 = lambda_Rs, par2_name = "lambda_R")
orch_D3_all_dt50 <- full_dt_fun(sim_list = orch_D3_all, max_yr = 51, n_pars = n_sample1, prop_set = prop_all, par_set2 = lambda_Rs, par2_name = "lambda_R")


#lab_max <- 4000

```


3B: Continuous allocation curve. X = proportion to reef (at least 0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0). Main panel full-range; zoomed inset on the 0.5–1 region where the nonlinearities live. Y = reef cover (top panel) and total ROI (bottom panel). Stack the two panels sharing the x-axis.

3C — Allocation × disturbance. Same continuous curve with vs. without severe disturbance. Shows the nursery advantage grows under stress.

```{r}

#colnames(orch_D0_all_dt5)

metric_cols <- c("#99DDFF", "#FFAABB", "#AAAA00", "#BBCC33","#EE8866", "#44BB99") # coral cover on reef, orchard function, reef ROI, total ROI, total costs, recruits outplanted to reef

plot_cols <- c("#99DDFF","#BBCC33") 

plot_colsD <- c("#7EA0A8","#8E9470")

dt_plot <- orch_D0_all_dt5

dt_plot1 <- dt_plot %>% filter(par2 == lambda_Rs[2]) %>% rename(metric = reef_cover_mean)

dt_plot2 <- dt_plot1 %>% select(metric, prop_out) %>% group_by(prop_out) %>% summarize(mean = mean(metric), Q05 = quantile(metric, probs = 0.05), Q95 = quantile(metric, probs = 0.95)) %>% ungroup()

plot(x = dt_plot2$prop_out, y = dt_plot2$mean, type = "l", ylim = c(0, max(dt_plot2$Q95)), las = 1, lwd = 2, col = plot_cols[1], xlab = NA, ylab = NA)
polygon(x = c(dt_plot2$prop_out, rev(dt_plot2$prop_out)), y = c(dt_plot2$Q05, rev(dt_plot2$Q95)), border = NA, col = adjustcolor(plot_cols[1], alpha.f = 0.25))
points(x = dt_plot2$prop_out[which(dt_plot2$mean == max(dt_plot2$mean))], y = dt_plot2$mean[which(dt_plot2$mean == max(dt_plot2$mean))], pch = 16, col = plot_cols[1], cex = 1.1)


# repeat with disturbance
dt_plot <- orch_D3_all_dt5

dt_plot1 <- dt_plot %>% filter(par2 == lambda_Rs[2]) %>% rename(metric = reef_cover_mean)

dt_plot2 <- dt_plot1 %>% select(metric, prop_out) %>% group_by(prop_out) %>% summarize(mean = mean(metric), Q05 = quantile(metric, probs = 0.05), Q95 = quantile(metric, probs = 0.95)) %>% ungroup()

lines(x = dt_plot2$prop_out, y = dt_plot2$mean, type = "l", lwd = 2, col = plot_colsD[1], lty = 2)
polygon(x = c(dt_plot2$prop_out, rev(dt_plot2$prop_out)), y = c(dt_plot2$Q05, rev(dt_plot2$Q95)), border = NA, col = adjustcolor(plot_colsD[1], alpha.f = 0.25))
points(x = dt_plot2$prop_out[which(dt_plot2$mean == max(dt_plot2$mean))], y = dt_plot2$mean[which(dt_plot2$mean == max(dt_plot2$mean))], pch = 17, col = plot_colsD[1])


# total ROI

dt_plot <- orch_D0_all_dt5

dt_plot1 <- dt_plot %>% filter(par2 == lambda_Rs[2]) %>% rename(metric = tot_ROI_mean)

dt_plot2 <- dt_plot1 %>% select(metric, prop_out) %>% group_by(prop_out) %>% summarize(mean = mean(metric), Q05 = quantile(metric, probs = 0.05), Q95 = quantile(metric, probs = 0.95)) %>% ungroup()

plot(x = dt_plot2$prop_out, y = dt_plot2$mean, type = "l", ylim = c(0, max(dt_plot2$Q95)), las = 1, lwd = 2, col = plot_cols[2], xlab = NA, ylab = NA)
polygon(x = c(dt_plot2$prop_out, rev(dt_plot2$prop_out)), y = c(dt_plot2$Q05, rev(dt_plot2$Q95)), border = NA, col = adjustcolor(plot_cols[2], alpha.f = 0.25))
points(x = dt_plot2$prop_out[which(dt_plot2$mean == max(dt_plot2$mean))], y = dt_plot2$mean[which(dt_plot2$mean == max(dt_plot2$mean))], pch = 16, col = plot_cols[2], cex = 1.1)


# repeat with disturbance
dt_plot <- orch_D3_all_dt5

dt_plot1 <- dt_plot%>% filter(par2 == lambda_Rs[2]) %>% rename(metric = tot_ROI_mean)

dt_plot2 <- dt_plot1 %>% select(metric, prop_out) %>% group_by(prop_out) %>% summarize(mean = mean(metric), Q05 = quantile(metric, probs = 0.05), Q95 = quantile(metric, probs = 0.95)) %>% ungroup()

lines(x = dt_plot2$prop_out, y = dt_plot2$mean, type = "l", lwd = 2, col = plot_colsD[2], lty = 2)
polygon(x = c(dt_plot2$prop_out, rev(dt_plot2$prop_out)), y = c(dt_plot2$Q05, rev(dt_plot2$Q95)), border = NA, col = adjustcolor(plot_colsD[2], alpha.f = 0.25))
points(x = dt_plot2$prop_out[which(dt_plot2$mean == max(dt_plot2$mean))], y = dt_plot2$mean[which(dt_plot2$mean == max(dt_plot2$mean))], pch = 17, col = plot_colsD[2])


# repeat with no reference reef supply


dt_plot <- orch_D0_all_dt5

dt_plot1 <- dt_plot %>% filter(par2 == lambda_Rs[1]) %>% rename(metric = reef_cover_mean)

dt_plot2 <- dt_plot1 %>% select(metric, prop_out) %>% group_by(prop_out) %>% summarize(mean = mean(metric), Q05 = quantile(metric, probs = 0.05), Q95 = quantile(metric, probs = 0.95)) %>% ungroup()

plot(x = dt_plot2$prop_out, y = dt_plot2$mean, type = "l", ylim = c(0, max(dt_plot2$Q95)), las = 1, lwd = 2, col = plot_cols[1], xlab = NA, ylab = NA)
polygon(x = c(dt_plot2$prop_out, rev(dt_plot2$prop_out)), y = c(dt_plot2$Q05, rev(dt_plot2$Q95)), border = NA, col = adjustcolor(plot_cols[1], alpha.f = 0.25))
points(x = dt_plot2$prop_out[which(dt_plot2$mean == max(dt_plot2$mean))], y = dt_plot2$mean[which(dt_plot2$mean == max(dt_plot2$mean))], pch = 16, col = plot_cols[1], cex = 1.1)

# repeat with disturbance
dt_plot <- orch_D3_all_dt5

dt_plot1 <- dt_plot %>% filter(par2 == lambda_Rs[1]) %>% rename(metric = reef_cover_mean)

dt_plot2 <- dt_plot1 %>% select(metric, prop_out) %>% group_by(prop_out) %>% summarize(mean = mean(metric), Q05 = quantile(metric, probs = 0.05), Q95 = quantile(metric, probs = 0.95)) %>% ungroup()

lines(x = dt_plot2$prop_out, y = dt_plot2$mean, type = "l", lwd = 2, col = plot_colsD[1], lty = 2)
polygon(x = c(dt_plot2$prop_out, rev(dt_plot2$prop_out)), y = c(dt_plot2$Q05, rev(dt_plot2$Q95)), border = NA, col = adjustcolor(plot_colsD[1], alpha.f = 0.25))
points(x = dt_plot2$prop_out[which(dt_plot2$mean == max(dt_plot2$mean))], y = dt_plot2$mean[which(dt_plot2$mean == max(dt_plot2$mean))], pch = 17, col = plot_colsD[1])


# total ROI

dt_plot <- orch_D0_all_dt5

dt_plot1 <- dt_plot %>% filter(par2 == lambda_Rs[1]) %>% rename(metric = tot_ROI_mean)

dt_plot2 <- dt_plot1 %>% select(metric, prop_out) %>% group_by(prop_out) %>% summarize(mean = mean(metric), Q05 = quantile(metric, probs = 0.05), Q95 = quantile(metric, probs = 0.95)) %>% ungroup()

plot(x = dt_plot2$prop_out, y = dt_plot2$mean, type = "l", ylim = c(0, max(dt_plot2$Q95)), las = 1, lwd = 2, col = plot_cols[2], xlab = NA, ylab = NA)
polygon(x = c(dt_plot2$prop_out, rev(dt_plot2$prop_out)), y = c(dt_plot2$Q05, rev(dt_plot2$Q95)), border = NA, col = adjustcolor(plot_cols[2], alpha.f = 0.25))
points(x = dt_plot2$prop_out[which(dt_plot2$mean == max(dt_plot2$mean))], y = dt_plot2$mean[which(dt_plot2$mean == max(dt_plot2$mean))], pch = 16, col = plot_cols[2])

# repeat with disturbance
dt_plot <- orch_D3_all_dt5

dt_plot1 <- dt_plot %>% filter(par2 == lambda_Rs[1]) %>% rename(metric = tot_ROI_mean)

dt_plot2 <- dt_plot1 %>% select(metric, prop_out) %>% group_by(prop_out) %>% summarize(mean = mean(metric), Q05 = quantile(metric, probs = 0.05), Q95 = quantile(metric, probs = 0.95)) %>% ungroup()

lines(x = dt_plot2$prop_out, y = dt_plot2$mean, type = "l", lwd = 2, col = plot_colsD[2], lty = 2)
polygon(x = c(dt_plot2$prop_out, rev(dt_plot2$prop_out)), y = c(dt_plot2$Q05, rev(dt_plot2$Q95)), border = NA, col = adjustcolor(plot_colsD[2], alpha.f = 0.25))
points(x = dt_plot2$prop_out[which(dt_plot2$mean == max(dt_plot2$mean))], y = dt_plot2$mean[which(dt_plot2$mean == max(dt_plot2$mean))], pch = 17, col = plot_colsD[2])


```

update: portfolio effect isn't very strong until you get to very high proportions outplanted to reef. So maybe instead of repeating the middle panel with disturbance, instead make some sort of heatmap plot/countour plot showing outplanting proportion that maximizes reef cover for different combinations of 1) orchard resistance to disturbance and 2) reference reef supply? OR orchard resistance to disturbance and orchard fecundity, repeated with and without reference reef supply?


make the plots


SM: 50-year petals, full metric plots at 5 and 50 years, time to full orchard



# Figure 4 & SM
Fill with one size class — which wins? SC2. Paired sub-panel: stacked final-size-distribution at year 50 (mechanism: SC2s grow up to SC5).

```{r}

# area outplanted: try to approximate the area of the orchard, so number of reef start in orchard x area per reef star. Estimate area of reef star as 0.337 m2 (from https://onlinelibrary.wiley.com/doi/full/10.1111/rec.12866)
area_out <- 170 # 500*0.337 = 168.5; round up to 170
lambda_R_12345 <- round(170/(A_mids*0.0001)) #c(10000, 1000, 100, 10, 1)
out_sizes <- c(1, 2, 3, 4, 5)

# make sure there's no orchard investment:
rest_pars$reef_prop <- c(1,1)

# holding list for model output

mod_out_L <- list()

mod_ind <- matrix(NA, nrow = years, ncol = 5) # total number of individuals on reef
mod_area <- matrix(NA, nrow = years, ncol = 5) # total area covered on the reef

# total number in each size class after 5 and 50 years
mod_mat5 <- matrix(NA, nrow = 5, ncol = 5) # row = outplanting scenario
mod_mat50 <- matrix(NA, nrow = 5, ncol = 5)

dens_pars.r <- list()
dens_pars.r[[1]] <- list() # first reef treatment/subpop
dens_pars.r[[1]][[1]] <- 0 # first source to first reef (external recruits)
dens_pars.r[[1]][[2]] <- dens_pars.r[[1]][[1]] # second source to first reef (tiles outplanted immediately)
dens_pars.r[[1]][[3]] <- dens_pars.r[[1]][[1]] # third source to first reef (tiles retained in lab)

N0.l <- list()
N0.l[[1]] <- 0 # make sure there's nothing already in the lab
N0.l[[2]] <- 0 

for(j in 1:5){
  
 lambda_R_j <- lambda_R_12345[j]
  
  size_props <- matrix(NA, nrow = length(lab_treatments), ncol = n) # matrix for fractions of retained recruits in each size class at the end of their year in the lab
size_props[1, ] <- rep(0, 5) # first lab treatment (0_T1)
size_props[1, out_sizes[j]] <- 1
lab_pars <- list(s0 = matrix(1, nrow = years, ncol = 2), s1 = matrix(1, nrow = years, ncol = 2), m0 = 0, m1 = 0, sett_props = list(T1 = 1), size_props = size_props, size_props1 = size_props1)

sim_j <- rse_mod1(years, n, A_mids, surv_pars.rc, surv_pars.r, dens_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, surv_pars.o, dens_pars.o, growth_pars.o, shrink_pars.o, frag_pars.o, fec_pars.o, lambda, lambda_R = lambda_R_j, sigma_s, sigma_f, ext_rand, seeds, dist_yrs, dist_pars.r, dist_effects.r, dist_pars.o, dist_effects.o, orchard_treatments, reef_treatments, lab_treatments, lab_pars, rest_pars, N0.r, N0.o, N0.l)


  
mod_out_L[[j]] <- sim_j


mod_ind[,j] <- model_summ(model_sim = sim_j, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

mod_area[,j] <- model_summ(model_sim = sim_j, location = "reef", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

mod_ind1 <- model_summ(model_sim = sim_j, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(1))

mod_ind2 <- model_summ(model_sim = sim_j, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(2))

mod_ind3 <- model_summ(model_sim = sim_j, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(3))

mod_ind4 <- model_summ(model_sim = sim_j, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(4))

mod_ind5 <- model_summ(model_sim = sim_j, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(5))

mod_mat5[,j] <- c(mod_ind1[6], mod_ind2[6], mod_ind3[6], mod_ind4[6], mod_ind5[6])
mod_mat50[,j] <- c(mod_ind1[51], mod_ind2[51], mod_ind3[51], mod_ind4[51], mod_ind5[51])

}


# reset size props to default
size_props <- matrix(NA, nrow = length(lab_treatments), ncol = n)
size_props[1, ] <- c(1, 0, 0, 0, 0) # first lab treatment 

lab_pars <- list(s0 = s0, s1 = s1, m0 = m0, m1 = m1, sett_props = sett_props, size_props = size_props, size_props1 = size_props1)

# reset orchard investment
rest_pars$reef_prop <- c(0.5, 1)

dens_pars.r <- list()
dens_pars.r[[1]] <- list() # first reef treatment/subpop
dens_pars.r[[1]][[1]] <- 0.02 # first source to first reef (external recruits)
dens_pars.r[[1]][[2]] <- dens_pars.r[[1]][[1]] # second source to first reef (tiles outplanted immediately)
dens_pars.r[[1]][[3]] <- dens_pars.r[[1]][[1]] # third source to first reef (tiles retained in lab)

N0.l <- list()
N0.l[[1]] <- 1000 # start with some babies in the lab
N0.l[[2]] <- 0 


# lambda_R_12345[1]
# mod_out_L[[1]]$reef_pops[[1]][[2]][1,2]
# lambda_R_12345[2]
# mod_out_L[[2]]$reef_pops[[1]][[2]][2,2]
# lambda_R_12345[3]
# mod_out_L[[3]]$reef_pops[[1]][[2]][3,2]
# lambda_R_12345[4]
# mod_out_L[[4]]$reef_pops[[1]][[2]][4,2]
# lambda_R_12345[5]
# mod_out_L[[5]]$reef_pops[[1]][[2]][5,2]

```



```{r}
cols1 <- c("#5289C7", "#4EB265", "#F7F056", "#F1932D", "#DC050C")


# plot(x = c(1:years), y = rep(0, years), ylim = c(0, max(mod_ind)), col = NA, xlab = NA, ylab = NA, las = 1)
# for(j in 1:5){
#   lines(x = c(1:years), y = mod_ind[,j], type = "l", lty = j, col = cols1[j], lwd = 2)
# }

plot(x = c(1:years), y = rep(0, years), ylim = c(0, max(mod_area)), col = NA, xlab = NA, ylab = NA, las = 1)
for(j in 1:5){
  lines(x = c(1:years), y = mod_area[,j], type = "l", lty = j, col = cols1[j], lwd = 2)
}
legend("topleft", legend = c("1 (340,000)", "2 (30,909)", "3 (3,400)", "4 (694)", "5 (182)"), col = cols1, lty = c(1:5), lwd = 2, bty = "n", ncol = 1, cex = 0.9)


# look at final size distributions

end_mat <- mod_mat5
plot(x = c(1, 2, 3, 4, 5), y = rep(0, 5), col = NA, ylim = c(0, max(mod_ind[6,])), xlim = c(0.5, 5.5), las = 1)
for(i in 1:5){
  polygon(x = c(i-0.25, i + 0.25, i + 0.25, i-0.25), y = c(0, 0, sum(end_mat[,i]), sum(end_mat[,i])), col = adjustcolor(col = "gray", alpha.f = 0))
  for(k in 1:5){
    
    if(k == 1){
      polygon(x = c(i-0.25, i + 0.25, i + 0.25, i-0.25), y = c(0, 0, end_mat[k,i], end_mat[k,i]), col = cols1[k])
    } else{
      polygon(x = c(i-0.25, i + 0.25, i + 0.25, i-0.25), y = c(sum(end_mat[1:(k-1),i]), sum(end_mat[1:(k-1),i]), sum(end_mat[1:(k-1),i]) + end_mat[k,i], sum(end_mat[1:(k-1),i]) + end_mat[k,i]), col = cols1[k])
    }
    
  }
  
}


end_mat <- mod_mat50
plot(x = c(1, 2, 3, 4, 5), y = rep(0, 5), col = NA, ylim = c(0, max(mod_ind[51,])), xlim = c(0.5, 5.5), las = 1)
for(i in 1:5){
  polygon(x = c(i-0.25, i + 0.25, i + 0.25, i-0.25), y = c(0, 0, sum(end_mat[,i]), sum(end_mat[,i])), col = adjustcolor(col = "gray", alpha.f = 0))
  for(k in 1:5){
    
    if(k == 1){
      polygon(x = c(i-0.25, i + 0.25, i + 0.25, i-0.25), y = c(0, 0, end_mat[k,i], end_mat[k,i]), col = cols1[k])
    } else{
      polygon(x = c(i-0.25, i + 0.25, i + 0.25, i-0.25), y = c(sum(end_mat[1:(k-1),i]), sum(end_mat[1:(k-1),i]), sum(end_mat[1:(k-1),i]) + end_mat[k,i], sum(end_mat[1:(k-1),i]) + end_mat[k,i]), col = cols1[k])
    }
    
  }
  
}


```


for supplement: 1) outplanting equal numbers and 2) equal survival

```{r}

# area outplanted: try to approximate the area of the orchard, so number of reef start in orchard x area per reef star. Estimate area of reef star as 0.337 m2 (from https://onlinelibrary.wiley.com/doi/full/10.1111/rec.12866)
area_out <- 170 # 500*0.337 = 168.5; round up to 170
lambda_R_12345 <- rep(200, 5) #c(10000, 1000, 100, 10, 1)
out_sizes <- c(1, 2, 3, 4, 5)

# make sure there's no orchard investment:
rest_pars$reef_prop <- c(1,1)

# holding list for model output

mod_out_L2 <- list()

mod_ind2 <- matrix(NA, nrow = years, ncol = 5) # total number of individuals on reef
mod_area2 <- matrix(NA, nrow = years, ncol = 5) # total area covered on the reef

# total number in each size class after 5 and 50 years
mod_mat5_2 <- matrix(NA, nrow = 5, ncol = 5) # row = outplanting scenario
mod_mat50_2 <- matrix(NA, nrow = 5, ncol = 5)

dens_pars.r <- list()
dens_pars.r[[1]] <- list() # first reef treatment/subpop
dens_pars.r[[1]][[1]] <- 0 # first source to first reef (external recruits)
dens_pars.r[[1]][[2]] <- dens_pars.r[[1]][[1]] # second source to first reef (tiles outplanted immediately)
dens_pars.r[[1]][[3]] <- dens_pars.r[[1]][[1]] # third source to first reef (tiles retained in lab)

N0.l <- list()
N0.l[[1]] <- 0 # make sure there's nothing already in the lab
N0.l[[2]] <- 0 

for(j in 1:5){
  
 lambda_R_j <- lambda_R_12345[j]
  
  size_props <- matrix(NA, nrow = length(lab_treatments), ncol = n) # matrix for fractions of retained recruits in each size class at the end of their year in the lab
size_props[1, ] <- rep(0, 5) # first lab treatment (0_T1)
size_props[1, out_sizes[j]] <- 1
lab_pars <- list(s0 = matrix(1, nrow = years, ncol = 2), s1 = matrix(1, nrow = years, ncol = 2), m0 = 0, m1 = 0, sett_props = list(T1 = 1), size_props = size_props, size_props1 = size_props1)

sim_j <- rse_mod1(years, n, A_mids, surv_pars.rc, surv_pars.r, dens_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, surv_pars.o, dens_pars.o, growth_pars.o, shrink_pars.o, frag_pars.o, fec_pars.o, lambda, lambda_R = lambda_R_j, sigma_s, sigma_f, ext_rand, seeds, dist_yrs, dist_pars.r, dist_effects.r, dist_pars.o, dist_effects.o, orchard_treatments, reef_treatments, lab_treatments, lab_pars, rest_pars, N0.r, N0.o, N0.l)


  
mod_out_L2[[j]] <- sim_j


mod_ind2[,j] <- model_summ(model_sim = sim_j, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

mod_area2[,j] <- model_summ(model_sim = sim_j, location = "reef", metric = "area_m2", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments))

mod_ind1 <- model_summ(model_sim = sim_j, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(1))

mod_ind2 <- model_summ(model_sim = sim_j, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(2))

mod_ind3 <- model_summ(model_sim = sim_j, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(3))

mod_ind4 <- model_summ(model_sim = sim_j, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(4))

mod_ind5 <- model_summ(model_sim = sim_j, location = "reef", metric = "ind", n_reef = length(reef_treatments), n_orchard = length(orchard_treatments), n_lab = length(lab_treatments), size_classes = c(5))

mod_mat5_2[,j] <- c(mod_ind1[6], mod_ind2[6], mod_ind3[6], mod_ind4[6], mod_ind5[6])
mod_mat50_2[,j] <- c(mod_ind1[51], mod_ind2[51], mod_ind3[51], mod_ind4[51], mod_ind5[51])

}


# reset size props to default
size_props <- matrix(NA, nrow = length(lab_treatments), ncol = n)
size_props[1, ] <- c(1, 0, 0, 0, 0) # first lab treatment 

lab_pars <- list(s0 = s0, s1 = s1, m0 = m0, m1 = m1, sett_props = sett_props, size_props = size_props, size_props1 = size_props1)

# reset orchard investment
rest_pars$reef_prop <- c(0.9, 1)

dens_pars.r <- list()
dens_pars.r[[1]] <- list() # first reef treatment/subpop
dens_pars.r[[1]][[1]] <- 0.02 # first source to first reef (external recruits)
dens_pars.r[[1]][[2]] <- dens_pars.r[[1]][[1]] # second source to first reef (tiles outplanted immediately)
dens_pars.r[[1]][[3]] <- dens_pars.r[[1]][[1]] # third source to first reef (tiles retained in lab)

N0.l <- list()
N0.l[[1]] <- 1000 # start with some babies in the lab
N0.l[[2]] <- 0 



```


Equal survival: REMEMBER to update recruit survival as well



Budgeting figure: need to decide on lab grow out or nursery. Nursery might make more sense given 1) there is lower parameter uncertainty (already have growth estimates) and 2) it matches the paired panel with fixed area, and 3) the budgets are more comparable (diver/boat costs either spent on outplanting more babies or moving size class 2 corals). 

50% orchard investment

maybe plot everything relative to the default case? (for each parameter combination, calculate prop change relative to default for that parameter combination)

base number of corals to transplant on number of outplanting day in budget

start with some size class 2 and size class 1 corals in the orchard?

```{r}

prop_set1 <- c(0.5)

# update orchard initial conditions 
N0.o <- list()
N0.o[[1]] <- list() # first orchard treatment
N0.o[[1]][[1]] <- c(0.25*orchard_size, 0.5*orchard_size, rep(0, 3)) # first source
N0.o[[1]][[2]] <- rep(0, n) # second source

# 50% investment in reef
orch_D0_def_50 <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = prop_set1, par_set2 = lambda_Rs[2], par2_name = "lambda_R", dist = T, dist_yrs = dist_yrs_i, dist_pars_list = dist_pars_list_i)

# make metric data frames
orch_D0_def_dt5 <- full_dt_fun(sim_list = orch_D0_def_50, max_yr = 6, n_pars = n_sample1, prop_set = prop_set1, par_set2 = lambda_Rs[2], par2_name = "lambda_R", annual_costs = T)
orch_D0_def_dt50 <- full_dt_fun(sim_list = orch_D0_def_50, max_yr = 51, n_pars = n_sample1, prop_set = prop_set1, par_set2 = lambda_Rs[2], par2_name = "lambda_R", annual_costs = T)

#orch_D0_def_dt5$costs_list

# update transplanting parameters
transplant <- rep(1, years)
rest_pars$transplant <- transplant

orch_D0_SC2_50 <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = prop_set1, par_set2 = lambda_Rs[2], par2_name = "lambda_R", dist = T, dist_yrs = dist_yrs_i, dist_pars_list = dist_pars_list_i)

orch_D0_SC2_dt5 <- full_dt_fun(sim_list = orch_D0_SC2_50, max_yr = 6, n_pars = n_sample1, prop_set = prop_set1, par_set2 = lambda_Rs[2], par2_name = "lambda_R", annual_costs = T)
orch_D0_SC2_dt50 <- full_dt_fun(sim_list = orch_D0_SC2_50, max_yr = 51, n_pars = n_sample1, prop_set = prop_set1, par_set2 = lambda_Rs[2], par2_name = "lambda_R", annual_costs = T)

# reset initial conditions in orchard

N0.o <- list()
N0.o[[1]] <- list() # first orchard treatment
N0.o[[1]][[1]] <- rep(0, n) # first source
N0.o[[1]][[2]] <- rep(0, n) # second source

# reset transplanting
transplant <- rep(0, years)
rest_pars$transplant <- transplant

```


```{r}
# process costs
# differences in costs are due to differences in number of corals in the orchard (due to differences in orchard parameters)
cost_mat_fun <- function(dt_list, max_year){
  
  cost_mat <- matrix(NA, nrow = n_sample1, ncol = max_year)

for(i in 1:n_sample1){
  cost_mat[i,] <- dt_list[[i]][[1]][[1]]
}
  return(cost_mat)
}

costs_D0_def_5 <- cost_mat_fun(dt_list = orch_D0_def_dt5$costs_list, max_year = 5)

costs_D0_SC2_5 <- cost_mat_fun(dt_list = orch_D0_SC2_dt5$costs_list, max_year = 5)

# calculate the average increase in costs each year across all parameter combinations

# calculate the total difference in cost across all 5 years for each parameter combination, then calculate the average total difference?

# or calculate the max annual cost for each parameter combination 

#mean_diff_5 <- apply(costs_D0_SC2_5 - costs_D0_def_5, MARGIN = 2, mean)
#apply(costs_D0_SC2_5 - costs_D0_def_5, MARGIN = 2, sd)

mean(apply(costs_D0_SC2_5, MARGIN = 1, max)- apply(costs_D0_def_5, MARGIN = 1, max))


mean_diff_5

#max(costs_D0_def_5 - costs_D0_SC2_5)
#min(costs_D0_def_5 - costs_D0_SC2_5)


```

complicated because it's not just orchard outplanting vs. reef outplanting; also have orchard maintenance which depends on how many corals are in the orchard, and there's inflation so things get more expensive each year. But could still try to estimate budgets each year, and just assume you need to have enough to perform maintenance on the orchard corals? 


Budgeting

```{r}
# plan out 5 year budget: need to make sure you have enough to collect spawn from reef only (more expensive), perform maintenance on full orchard, and outplant 50% of substrates to reef and 50% to orchard each year

# cost*(1 + r_inf)^(2:max_yr)
# r_inf <- 0.056

# can use the same data function with fake simulation results

# lab_max = max number of substrates
# n_trans = max number of orchard colonies transplanted each year
      
# sum(rest_pars$trans_mats[[1]][[1]][1,])
budget_fun <- function(max_yr, lab_max, n_trans){
  
  
budget_sim <- list()


budget_sim$reef_pops_pre <- list()
budget_sim$reef_pops_pre[[1]] <- list()
budget_sim$reef_pops_pre[[1]][[1]] <- matrix(0, 5, max_yr)
budget_sim$reef_pops_pre[[1]][[2]] <- matrix(0, 5, max_yr)
budget_sim$reef_pops_pre[[1]][[3]] <- matrix(0, 5, max_yr)

budget_sim$orchard_pops_pre <- list()
budget_sim$orchard_pops_pre[[1]] <- list()
budget_sim$orchard_pops_pre[[1]][[1]] <- matrix(0, 5, max_yr)
budget_sim$orchard_pops_pre[[1]][[2]] <- matrix(0, 5, max_yr)

budget_sim$reef_tiles_out <- rep(0.5*lab_max, max_yr)
budget_sim$orchard_tiles_out <- rep(0.5*lab_max, max_yr)
budget_sim$orchard_tiles <- list()

budget_sim$trans_colonies_tot <- rep(n_trans, max_yr)
budget_sim$reef_babies_used <- rep(lambda_R, max_yr)
budget_sim$orchard_babies_used <- rep(0, max_yr)

for(kk in 1:length(orchard_treatments)){ # in case there's more than one orchard
  
  if(kk == 1){
    budget_sim$orchard_tiles[[kk]] <- rep(rest_pars$orchard_size, max_yr)
  } else{
    budget_sim$orchard_tiles[[kk]] <- rep(0, max_yr)
  }
         
}

# calculate the costs
dt_out <- metrics_dt_fun(sim = budget_sim, max_yr, annual_costs = T)
  
return(dt_out$tot_costs_yr)

}


def_budget <- budget_fun(max_yr = 6, lab_max, n_trans = 0)

SC2_budget <- budget_fun(max_yr = 6, lab_max, n_trans = sum(rest_pars$trans_mats[[1]][[1]][1,]))

# def_budget
# SC2_budget

c_extra <- SC2_budget - def_budget # put that difference into lab size
c_extra <- min(c_extra)


# 11700 per year = how many extra tiles?
# x = number of tiles, costs = purchasing each year plus outplanting (50% to reef and 50% to orchard)

x_set <- seq(from = 0, to = 4000, by = 1)

c_subs_set <- x_set*(0.5*1 + 0.5*7)

c_reef_set <- 300*ceiling(0.5*x_set/1250) + 200*0.5*x_set/1250 + 380/2000*0.5*x_set
      
c_orch_set <- 300*ceiling(0.5*x_set/800) + 200*0.5*x_set/800 + 250/2000 * 0.5*x_set

c_tot_set <- c_subs_set + c_reef_set + c_orch_set

c_diff_set <- c_extra - c_tot_set

x_set[which(abs(c_diff_set)==min(abs(c_diff_set)))]
c_diff_set[2476]

c_extra - c_tot_set[2476]

c_extra 
c_tot_set[2476]

# so the answer is 2476 extra tiles

SC1L_budget <- budget_fun(max_yr = 6, lab_max + 1600, n_trans =0)

#SC1L_budget <- budget_fun(max_yr = 6, lab_max + 2476, n_trans =0)

(SC2_budget-SC1L_budget)/SC1L_budget*100 # within 10%
```


```{r}
# now repeat SC1 outplanting with bigger lab


rest_pars$lab_max <- 4000 + 1600

prop_set1 <- c(0.5)

# update orchard initial conditions 
N0.o <- list()
N0.o[[1]] <- list() # first orchard treatment
N0.o[[1]][[1]] <- c(0.25*orchard_size, 0.5*orchard_size, rep(0, 3)) # first source
N0.o[[1]][[2]] <- rep(0, n) # second source

# 50% investment in reef
orch_D0_SC1L_50 <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = c(0.5), par_set2 = lambda_Rs[2], par2_name = "lambda_R", dist = T, dist_yrs = dist_yrs_i, dist_pars_list = dist_pars_list_i)

# make metric data frames
orch_D0_SC1L_dt5 <- full_dt_fun(sim_list = orch_D0_SC1L_50, max_yr = 6, n_pars = n_sample1, prop_set = prop_set1, par_set2 = lambda_Rs[2], par2_name = "lambda_R", annual_costs = T)
orch_D0_SC1L_dt50 <- full_dt_fun(sim_list = orch_D0_SC1L_50, max_yr = 51, n_pars = n_sample1, prop_set = prop_set1, par_set2 = lambda_Rs[2], par2_name = "lambda_R", annual_costs = T)

#orch_D0_def_dt5$costs_list



# reset initial conditions in orchard

N0.o <- list()
N0.o[[1]] <- list() # first orchard treatment
N0.o[[1]][[1]] <- rep(0, n) # first source
N0.o[[1]][[2]] <- rep(0, n) # second source

# rest lab capacity

rest_pars$lab_max <- 4000



```


```{r}
#orch_D0_SC1L_50[[1]][[1]][[1]]$trans_colonies_tot

costs_D0_SC1L_5 <- cost_mat_fun(dt_list = orch_D0_SC1L_dt5$costs_list, max_year = 5)


mean(apply(costs_D0_SC2_5, MARGIN = 1, max))
     
mean(apply(costs_D0_def_5, MARGIN = 1, max))

mean(apply(costs_D0_SC1L_5, MARGIN = 1, max))


mean(orch_D0_SC1L_dt5$dt_i$reef_cover_mean)
mean(orch_D0_def_dt5$dt_i$reef_cover_mean)
mean(orch_D0_SC2_dt5$dt_i$reef_cover_mean)

mean(orch_D0_SC1L_dt5$dt_i$tot_ROI_mean)
mean(orch_D0_def_dt5$dt_i$tot_ROI_mean)
mean(orch_D0_SC2_dt5$dt_i$tot_ROI_mean)

mean(orch_D0_SC1L_dt5$dt_i$reef_ROI_mean)
mean(orch_D0_def_dt5$dt_i$reef_ROI_mean)
mean(orch_D0_SC2_dt5$dt_i$reef_ROI_mean)

# check differences in costs
max((orch_D0_SC2_dt5$dt_i$tot_costs - orch_D0_SC1L_dt5$dt_i$tot_costs)/orch_D0_SC1L_dt5$dt_i$tot_costs)*100

max((orch_D0_SC2_dt5$dt_i$tot_costs - orch_D0_def_dt5$dt_i$tot_costs)/orch_D0_def_dt5$dt_i$tot_costs)*100


# check over 50 years
max((orch_D0_SC2_dt50$dt_i$tot_costs - orch_D0_SC1L_dt50$dt_i$tot_costs)/orch_D0_SC1L_dt50$dt_i$tot_costs)*100

max((orch_D0_SC2_dt50$dt_i$tot_costs - orch_D0_def_dt50$dt_i$tot_costs)/orch_D0_def_dt50$dt_i$tot_costs)*100

```


```{r}
#View(orch_D0_SC2_dt5$dt_i)

# for each parameter set, get coral cover on reef, total costs, and total ROI relative to the default simulation

relative_metrics_fun <- function(dt2, dt_def, n_sample1){
  
dt2$par_rep <- c(1:n_sample1)  
dt_def$par_rep <- c(1:n_sample1)  

reef_coral <- rep(NA, n_sample1)
tot_ROI <- rep(NA, n_sample1)
totI_ROI <- rep(NA, n_sample1)
tot_costs <- rep(NA, n_sample1)

for(i in 1:n_sample1){
  dt_def_i <- dt_def[which(dt_def$par_rep == i),]
  dt2_i <- dt2[which(dt2$par_rep == i),]
  
  reef_coral[i] <- dt2_i$reef_cover_mean/dt_def_i$reef_cover_mean
  tot_ROI[i] <- dt2_i$tot_ROI_mean/dt_def_i$tot_ROI_mean
  totI_ROI[i] <- dt2_i$totI_ROI_mean/dt_def_i$totI_ROI_mean
  tot_costs[i] <- dt2_i$tot_costs/dt_def_i$tot_costs
}

return(list(reef_coral = reef_coral, tot_ROI = tot_ROI, totI_ROI = totI_ROI, tot_costs = tot_costs))
  
}



```


plot results

```{r}

pch_set <- c(16, 17)

col_set <- c("black", "blue") # CHANGE THESE to match the size 1 and size 2 size classes from the outplanting size figure (once that color palette is decided)

SC1L_def <- relative_metrics_fun(dt2 = orch_D0_SC1L_dt5$dt_i, dt_def = orch_D0_def_dt5$dt_i, n_sample1)

SC2_def <- relative_metrics_fun(dt2 = orch_D0_SC2_dt5$dt_i, dt_def = orch_D0_def_dt5$dt_i, n_sample1)


xmax <- max(c(SC1L_def$reef_coral, SC2_def$reef_coral))
ymax <- max(c(SC1L_def$tot_ROI, SC2_def$tot_ROI))

plot(x = c(0, xmax), y = c(0, ymax), type = "l", col = NA, las = 1, xlab = NA, ylab = NA)
points(x = mean(SC1L_def$reef_coral), y = mean(SC1L_def$tot_ROI), pch = pch_set[1], col = col_set[1])
arrows(x0 = mean(SC1L_def$reef_coral), y0 = quantile(SC1L_def$tot_ROI, 0.05), y1 = quantile(SC1L_def$tot_ROI, 0.95), code = 3, angle = 90, length = 0.025, col = col_set[1])
arrows(x0 = quantile(SC1L_def$reef_coral, 0.05), y0 = mean(SC1L_def$tot_ROI), x1 = quantile(SC1L_def$reef_coral, 0.95), code = 3, angle = 90, length = 0.025, col = col_set[1])
points(x = mean(SC2_def$reef_coral), y = mean(SC2_def$tot_ROI), pch = pch_set[2], col = col_set[2])
arrows(x0 = mean(SC2_def$reef_coral), y0 = quantile(SC2_def$tot_ROI, 0.05), y1 = quantile(SC2_def$tot_ROI, 0.95), code = 3, angle = 90, length = 0.025, col = col_set[2])
arrows(x0 = quantile(SC2_def$reef_coral, 0.05), y0 = mean(SC2_def$tot_ROI), x1 = quantile(SC2_def$reef_coral, 0.95), code = 3, angle = 90, length = 0.025, col = col_set[2])
abline(h = 1, lty = 2, col = "gray30")
abline(v = 1, lty = 2, col = "gray30")


```

repeat for 50 years

```{r}
SC1L_def <- relative_metrics_fun(dt2 = orch_D0_SC1L_dt50$dt_i, dt_def = orch_D0_def_dt50$dt_i, n_sample1)

SC2_def <- relative_metrics_fun(dt2 = orch_D0_SC2_dt50$dt_i, dt_def = orch_D0_def_dt50$dt_i, n_sample1)


xmax <- max(c(SC1L_def$reef_coral, SC2_def$reef_coral))
ymax <- max(c(SC1L_def$tot_ROI, SC2_def$tot_ROI))

plot(x = c(0, xmax), y = c(0, ymax), type = "l", col = NA, las = 1, xlab = NA, ylab = NA)
points(x = mean(SC1L_def$reef_coral), y = mean(SC1L_def$tot_ROI), pch = pch_set[1], col = col_set[1])
arrows(x0 = mean(SC1L_def$reef_coral), y0 = quantile(SC1L_def$tot_ROI, 0.05), y1 = quantile(SC1L_def$tot_ROI, 0.95), code = 3, angle = 90, length = 0.025, col = col_set[1])
arrows(x0 = quantile(SC1L_def$reef_coral, 0.05), y0 = mean(SC1L_def$tot_ROI), x1 = quantile(SC1L_def$reef_coral, 0.95), code = 3, angle = 90, length = 0.025, col = col_set[1])
points(x = mean(SC2_def$reef_coral), y = mean(SC2_def$tot_ROI), pch = pch_set[2], col = col_set[2])
arrows(x0 = mean(SC2_def$reef_coral), y0 = quantile(SC2_def$tot_ROI, 0.05), y1 = quantile(SC2_def$tot_ROI, 0.95), code = 3, angle = 90, length = 0.025, col = col_set[2])
arrows(x0 = quantile(SC2_def$reef_coral, 0.05), y0 = mean(SC2_def$tot_ROI), x1 = quantile(SC2_def$reef_coral, 0.95), code = 3, angle = 90, length = 0.025, col = col_set[2])
abline(h = 1, lty = 2, col = "gray30")
abline(v = 1, lty = 2, col = "gray30")
```


repeat with disturbance?

# Misc


Hypothesis for Figure 3: benefits of orchard investment depends in part on lab capacity, otherwise orchard benefits max out when the lab is full. So if orchard is maxing out, it's better to put more on the reef just to get as many corals out there as possible.

Update: wasn't much variation -- portfolio effect doesn't seem to be very strong

```{r}

# years <- 6 # just do 6 years for now to reduce computation time
# 
# prop_all <- seq(from = 0, to = 1, length.out = 20)
# #prop_all <- seq(from = 0, to = 1, length.out = 5)
# lab_max_set <- seq(from = 3000, to = 12000, length.out = 20)
# 
# lambda_R <- 0
# 
# #orch_D0 <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = prop_main, par_set2 = NULL, par2_name = "none", dist = F, dist_yrs = NULL, dist_pars_list = NULL)
# 
# tic()
# orch_D0_all_lab_max <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = prop_all, par_set2 = lab_max_set, par2_name = "lab_max", dist = F, dist_yrs = NULL, dist_pars_list = NULL)
# 
# orch_D3_all_lab_max <- orch_exp_fun2(dem_pars_all = all_pars_R, n_pars = n_sample1, prop_set = prop_all, par_set2 = lab_max_set, par2_name = "lab_max", dist = T, dist_yrs = dist_yrs3, dist_pars_list = dist_pars_list3)
# toc()
# 
# # reset number of years to default
# years <- 51
# 
# lambda_R <- 1255111
# 
# # process the output
# 
# orch_D0_all_lab_max_dt5 <- full_dt_fun(sim_list = orch_D0_all_lab_max, max_yr = 6, n_pars = n_sample1, prop_set = prop_all, par_set2 = lab_max_set, par2_name = "lab_max")
# 
# orch_D3_all_lab_max_dt5 <- full_dt_fun(sim_list = orch_D3_all_lab_max, max_yr = 6, n_pars = n_sample1, prop_set = prop_all, par_set2 = lab_max_set, par2_name = "lab_max")
# 

```

takes 2-3 minutes


```{r}


# for each parameter combination
# for each proportion outplanted, find the range of lab sizes for which that proportion results in the max coral cover on the reef (NA if coral cover was never maximized at that proportion outplanted)

# for each parameter combination and lab size, get the max coral cover and the proportion that it occurs, then arrange by proportions
# 
# plot_cols <- c("#99DDFF","#7EA0A8") # without and with disturbance
# a1 <- 0.2
# lwd1 <- 3
# 
# dt_plot <- orch_D0_all_lab_max_dt5
# 
# dt_plot1 <- dt_plot %>% select(reef_cover_mean, par2, par_rep, prop_out) %>% group_by(par2, par_rep) %>% mutate(max_coral = max(reef_cover_mean)) %>% ungroup() %>% filter(reef_cover_mean == max_coral)
# 
# plot(x = prop_all, lab_max_set, col = NA)
# #points(x = dt_plot_i$prop_out, y = dt_plot_i$par2)
# for(i in 1:n_sample1){
#   for(j in 1:length(unique(dt_plot_i$prop_out))){
#     
#     dt_plot_i <- dt_plot1 %>% filter(par_rep == i)
#   dt_plot_j <- dt_plot_i %>% filter(prop_out == unique(dt_plot_i$prop_out)[j])
#   lines(x = dt_plot_j$prop_out, y = dt_plot_j$par2, col = adjustcolor(plot_cols[1], alpha.f = a1), lwd = lwd1)
# }
# 
# }
# 
# 
# dt_plot <- orch_D3_all_lab_max_dt5
# 
# dt_plot1 <- dt_plot %>% select(reef_cover_mean, par2, par_rep, prop_out) %>% group_by(par2, par_rep) %>% mutate(max_coral = max(reef_cover_mean)) %>% ungroup() %>% filter(reef_cover_mean == max_coral)
# 
# plot(x = prop_all, lab_max_set, col = NA)
# for(i in 1:n_sample1){
#   for(j in 1:length(unique(dt_plot_i$prop_out))){
#     
#     dt_plot_i <- dt_plot1 %>% filter(par_rep == i)
#   dt_plot_j <- dt_plot_i %>% filter(prop_out == unique(dt_plot_i$prop_out)[j])
#   lines(x = dt_plot_j$prop_out, y = dt_plot_j$par2, col = adjustcolor(plot_cols[2], alpha.f = a1), lwd = lwd1, lty = 1)
# }
# 
# }



```





