---
title: "rse_population_viability"
output: html_document
date: "2025-07-22"
---


README: population viability analyses (how much do you have to outplant (and of what sizes) to get a growing/persistent population, how long does it take to reach this point, and once there can you stop outplanting?)


```{r}

source("coral_demographic_funs.R")
source("rse_funs.R")

```


```{r}

years <- 100
n <- 5

A_mids <- c(0.1, 43, 369, 2158, 11171) # mean areas in each size class (cm^2)


# demographic parameters for each orchard and reef treatment
# reef parameters: survival
surv_pars.r <- list()
surv_pars.r[[1]] <- list() 
surv_pars.r[[1]][[1]] <- c(0.010, 0.760, 0.870, 0.950, 0.999) # survival probabilities for first source of recruits to the reef (external recruits); from 1st matrix in Vardi et al. 2012
#surv_pars.r[[1]][[1]] <- c(0.010, 0.760, 0.870, 0.950, 0.999) # dying (2nd matrix)
surv_pars.r[[1]][[2]] <- surv_pars.r[[1]][[1]] # survival probabilities for second source of individuals to the reef (outplanted individuals)


# reef parameters: growth
growth_pars.r <- list()
growth_pars.r[[1]] <- list() # first reef subpop
growth_pars.r[[1]][[1]] <- list(c(0.5, 0, 0, 0), c(0.197, 0.0132, 0), c(0.126, 0), 0.1158, NULL) # first source to first reef subpop (external recruits)
growth_pars.r[[1]][[2]] <- growth_pars.r[[1]][[1]]

# reef parameters: shrinkage
shrink_pars.r <- list()
shrink_pars.r[[1]] <- list() # first subpop
shrink_pars.r[[1]][[1]] <- list(NULL, c(0.01), c(0, 0.08), c(0, 0, 0.105), c(0, 0, 0.06, 0.11)) # first source
shrink_pars.r[[1]][[2]] <- shrink_pars.r[[1]][[1]]

# reef parameters: fragmentation
frag_pars.r <- list()
frag_pars.r[[1]] <- list() # first subpop
#frag_pars.r[[1]][[1]] <- list(NULL, c(0), c(0, 0), c(0, 0.06, 0.06), c(0, 0.28, 0.3, 0.23)) # first source
frag_pars.r[[1]][[1]] <- list(NULL, c(0), c(0, 0), c(0, 0.015, 0.015), c(0, 0.07, 0.08, 0.06)) # dying
#frag_pars.r[[1]][[1]] <- list(NULL, c(0), c(0, 0), c(0, 0, 0), c(0, 0.4, 0.5, 0.9))
frag_pars.r[[1]][[2]] <- frag_pars.r[[1]][[1]]

# reef parameters: fecundity
fec_pars.r <- list()
fec_pars.r[[1]] <- list() # first subpop
fec_pars.r[[1]][[1]] <- c(0, 0, 5000, 50000, 100000) # first source
fec_pars.r[[1]][[2]] <- fec_pars.r[[1]][[1]] 

lambda <- 0 # external recruitment to reef

# stochasticity parameters
sigma_s <- 0
sigma_f <- 0
ext_rand <- FALSE
seeds <- c(1000, 5000, 10000)

# initial conditions in each reef subpopulation
N0.r <- list()
N0.r[[1]] <- list() # first reef subpop
N0.r[[1]][[1]] <- rep(0, n) # first source in the reef (external recruits)
N0.r[[1]][[2]] <- rep(0, n) # second source in the reef (outplants)

# disturbance parameters
dist_yrs <- c(years + 10) # years when disturbance occurs, NA if none

# effects of disturbance on each reef subpop
dist_effects.r <- list()
dist_effects.r[[1]] <- list() # list where each element is the effects of disturbnce on corals from each source in the 1st reef subpop
dist_effects.r[[1]][[1]] <- list() # effects of disturbances on corals from first source in first reef subpop
dist_effects.r[[1]][[1]][[1]] <- c("survival") # effects of 1st disturbance on corals from first source in first reef subpop ([[1]][[1]][[2]] would be second disturbance, etc.)

dist_effects.r[[1]][[2]] <- list() # effects of disturbances on corals from first source in first reef subpop
dist_effects.r[[1]][[2]][[1]] <- c("survival") # effects of 1st disturbance on corals from second source in first reef subpop 

# disturbance parameters for each reef subpop
dist_pars.r <- list()

# first reef subpop
dist_pars.r[[1]] <- list() # first source in first reef subpop
dist_pars.r[[1]][[1]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[1]], dist_surv0 = list(surv_pars.r[[1]][[1]]*0.1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL) # list of disturbance parameters for the third source in the third reef population (defaults for each parameter type are NULL unless the disturbance affects them, as specified in dist_effects)
# dist_surv0 = list where ith element is the survival probabilities for the ith disturbance

dist_pars.r[[1]][[2]] <- dist_pars_fun(dist_yrs = dist_yrs, dist_effects.r[[1]][[2]], dist_surv0 = list(surv_pars.r[[1]][[2]]*0.1), dist_Tmat0 = NULL, dist_Fmat0 = NULL, dist_fec0 = NULL) 


# restoration parameters
out_pars <- matrix(0, nrow = years, ncol = n)
out_pars[2,1] <- 0.1
#out_pars[ ,1] <- c(1, 1, rep(0, years-2))#c(rep(1000, years/2), rep(0, years/2))
#out_pars[ ,5] <- c(rep(0, years/2), rep(0, years/2))

sim1 <- popvi_mod(years, n, A_mids, surv_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r,
                      fec_pars.r, lambda, sigma_s, sigma_f, ext_rand, seeds, dist_yrs,
                      dist_pars.r, dist_effects.r, out_pars, N0.r)


```


```{r}

reef_tot <- apply(sim1$reef_pops_pre[[1]][[1]], MARGIN = 2, sum) + apply(sim1$reef_pops_pre[[1]][[2]], MARGIN = 2, sum) 

reef_Atot <- rep(NA, years)

for(i in 1:years){
  
  reef_Atot[i] <- sum(sim1$reef_pops_pre[[1]][[1]][,i]*A_mids) + sum(sim1$reef_pops_pre[[1]][[2]][,i]*A_mids)
  
}


plot(x = c(1:years), y = reef_tot, type = "l", las = 1)

plot(x = c(1:years), y = reef_Atot, type = "l", las = 1)


```

If it's a dying population, it's going to stay that way (i.e., it will start to die as soon as outplanting stops). If it's a growing population, then it's going to grow and outplanting will only affect how quickly it grows. Also it seems like if it's dying, then external recruitment can't allow it to grow (so it will just maintain a population size approximately equal to the number of external recruits)



start with a dying population, know that outplanting needs to be continuous to maintain the population? but the goal is to get a self sustaining population...but ignore that for now and just explore the number you'd need to outplant to reach some population threshold by some amount of time

THOUGHT: could also vary the frequency of outplanting (e.g., once every year, once evey two years, etc.)

and make the x-axis 1) external recruitment rate or 2) fragmentation of largest corals?

First: x-axis = external recruitment rate

```{r}
 
ext_rec_set <- seq(from = 0, to = 50, length.out = 10) # external recruitment rate

out_set <- seq(from = 0, to = 100, by = 1) # number outplanted each year

t_goal <- 10 # amount of time in which you want to reach the restoration target

# holding matrices for population sizes after 10 years
pop1 <- matrix(NA, nrow = length(out_set), ncol = length(ext_rec_set)) # outplanting size 1
pop2 <- matrix(NA, nrow = length(out_set), ncol = length(ext_rec_set)) # outplanting size 2
pop3 <- matrix(NA, nrow = length(out_set), ncol = length(ext_rec_set)) # outplanting size 3
pop4 <- matrix(NA, nrow = length(out_set), ncol = length(ext_rec_set)) # outplanting size 4
pop5 <- matrix(NA, nrow = length(out_set), ncol = length(ext_rec_set)) # outplanting size 5

# holding matrices for area covered after 10 years
area1 <- matrix(NA, nrow = length(out_set), ncol = length(ext_rec_set)) # outplanting size 1
area2 <- matrix(NA, nrow = length(out_set), ncol = length(ext_rec_set)) # outplanting size 2
area3 <- matrix(NA, nrow = length(out_set), ncol = length(ext_rec_set)) # outplanting size 3
area4 <- matrix(NA, nrow = length(out_set), ncol = length(ext_rec_set)) # outplanting size 4
area5 <- matrix(NA, nrow = length(out_set), ncol = length(ext_rec_set)) # outplanting size 5

# outer for loop = external recruitment rate
# inner for loop = number outplanted

for(k in 1:length(ext_rec_set)){
  
  for(j in 1:length(out_set)){
    
    # outplanting size 1
    out_pars <- matrix(0, nrow = years, ncol = n)
    out_pars[ ,1] <- rep(out_set[j])
    
    simjk <- popvi_mod(years = t_goal, n, A_mids, surv_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, lambda = ext_rec_set[k], sigma_s, sigma_f, ext_rand, seeds, dist_yrs,
                      dist_pars.r, dist_effects.r, out_pars, N0.r)
    
    pop1[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal])
    area1[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]*A_mids) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal]*A_mids)
    
    
    # outplanting size 2
    out_pars <- matrix(0, nrow = years, ncol = n)
    out_pars[ ,2] <- rep(out_set[j])
    
    simjk <- popvi_mod(years = t_goal, n, A_mids, surv_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, lambda = ext_rec_set[k], sigma_s, sigma_f, ext_rand, seeds, dist_yrs,
                      dist_pars.r, dist_effects.r, out_pars, N0.r)
    
    pop2[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal])
    area2[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]*A_mids) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal]*A_mids)
    
     # outplanting size 3
    out_pars <- matrix(0, nrow = years, ncol = n)
    out_pars[ ,3] <- rep(out_set[j])
    
    simjk <- popvi_mod(years = t_goal, n, A_mids, surv_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, lambda = ext_rec_set[k], sigma_s, sigma_f, ext_rand, seeds, dist_yrs,
                      dist_pars.r, dist_effects.r, out_pars, N0.r)
    
    pop3[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal])
    area3[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]*A_mids) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal]*A_mids)
    
    # outplanting size 4
    out_pars <- matrix(0, nrow = years, ncol = n)
    out_pars[ ,4] <- rep(out_set[j])
    
    simjk <- popvi_mod(years = t_goal, n, A_mids, surv_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, lambda = ext_rec_set[k], sigma_s, sigma_f, ext_rand, seeds, dist_yrs,
                      dist_pars.r, dist_effects.r, out_pars, N0.r)
    
    pop4[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal])
    area4[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]*A_mids) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal]*A_mids)
    
    # outplanting size 5
    out_pars <- matrix(0, nrow = years, ncol = n)
    out_pars[ ,5] <- rep(out_set[j])
    
    simjk <- popvi_mod(years = t_goal, n, A_mids, surv_pars.r, growth_pars.r, shrink_pars.r, frag_pars.r, fec_pars.r, lambda = ext_rec_set[k], sigma_s, sigma_f, ext_rand, seeds, dist_yrs,
                      dist_pars.r, dist_effects.r, out_pars, N0.r)
    
    pop5[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal])
    area5[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]*A_mids) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal]*A_mids)
    
  }
  
}

```



```{r}

# plot results

N_goal <- 100 # goal for population size on reef

n_cols <- c("red", "orange", "green", "royalblue", "violet")

# now for each external recruitment rate, want to plot the minimum number of outplants needed to reach the restoration goal

plot(NULL, xlim=c(0,max(ext_rec_set)), ylim=c(0,50), ylab="Min. number outplanted", xlab="External recruitment rate", las = 1)
popn <- rep(NA, length(ext_rec_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(pop1[,k] >= N_goal))]
}
lines(x = ext_rec_set, y = popn, lty = 1, col = n_cols[1], lwd = 2)
popn <- rep(NA, length(ext_rec_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(pop2[,k] >= N_goal))]
}
lines(x = ext_rec_set, y = popn, lty = 2, col = n_cols[2], lwd = 2)
popn <- rep(NA, length(ext_rec_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(pop3[,k] >= N_goal))]
}
lines(x = ext_rec_set, y = popn, lty = 3, col = n_cols[3], lwd = 2)
popn <- rep(NA, length(ext_rec_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(pop4[,k] >= N_goal))]
}
lines(x = ext_rec_set, y = popn, lty = 4, col = n_cols[4], lwd = 2)
popn <- rep(NA, length(ext_rec_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(pop5[,k] >= N_goal))]
}
lines(x = ext_rec_set, y = popn, lty = 5, col = n_cols[5], lwd = 2)
legend("topleft", legend = as.character(c(1:5)), lty = c(1:5), bty = "n", ncol = 5, col = n_cols, lwd = 2)



```


```{r}
# what if goal is based on area?
# pop5[1,10]
# area5[1,10]
# 
# area1[100, 10]

A_goal <- 100000

plot(NULL, xlim=c(0,max(ext_rec_set)), ylim=c(0,50), ylab="Min. number outplanted", xlab="External recruitment rate", las = 1)
popn <- rep(NA, length(ext_rec_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(area1[,k] >= A_goal))]
}
lines(x = ext_rec_set, y = popn, lty = 1, col = n_cols[1], lwd = 2)
popn <- rep(NA, length(ext_rec_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(area2[,k] >= A_goal))]
}
lines(x = ext_rec_set, y = popn, lty = 2, col = n_cols[2], lwd = 2)
popn <- rep(NA, length(ext_rec_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(area3[,k] >= A_goal))]
}
lines(x = ext_rec_set, y = popn, lty = 3, col = n_cols[3], lwd = 2)
popn <- rep(NA, length(ext_rec_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(area4[,k] >= A_goal))]
}
lines(x = ext_rec_set, y = popn, lty = 4, col = n_cols[4], lwd = 2)
popn <- rep(NA, length(ext_rec_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(area5[,k] >= A_goal))]
}
lines(x = ext_rec_set, y = popn, lty = 5, col = n_cols[5], lwd = 2)
legend("topleft", legend = as.character(c(1:5)), lty = c(1:5), bty = "n", ncol = 5, col = n_cols, lwd = 2)

```


now repeat but vary fragmentation instead


```{r}
#frag_set <- seq(from = 0, to = 0.25, length.out = 10) # external recruitment rate
frag_set <- seq(from = 1, to = 10, length.out = 10) # external recruitment rate

out_set <- seq(from = 0, to = 100, by = 1) # number outplanted each year

t_goal <- 10 # amount of time in which you want to reach the restoration target

# holding matrices for population sizes after 10 years
pop1 <- matrix(NA, nrow = length(out_set), ncol = length(frag_set)) # outplanting size 1
pop2 <- matrix(NA, nrow = length(out_set), ncol = length(frag_set)) # outplanting size 2
pop3 <- matrix(NA, nrow = length(out_set), ncol = length(frag_set)) # outplanting size 3
pop4 <- matrix(NA, nrow = length(out_set), ncol = length(frag_set)) # outplanting size 4
pop5 <- matrix(NA, nrow = length(out_set), ncol = length(frag_set)) # outplanting size 5

# holding matrices for area covered after 10 years
area1 <- matrix(NA, nrow = length(out_set), ncol = length(frag_set)) # outplanting size 1
area2 <- matrix(NA, nrow = length(out_set), ncol = length(frag_set)) # outplanting size 2
area3 <- matrix(NA, nrow = length(out_set), ncol = length(frag_set)) # outplanting size 3
area4 <- matrix(NA, nrow = length(out_set), ncol = length(frag_set)) # outplanting size 4
area5 <- matrix(NA, nrow = length(out_set), ncol = length(frag_set)) # outplanting size 5

# outer for loop = external recruitment rate
# inner for loop = number outplanted

for(k in 1:length(frag_set)){
  
frag_pars.k <- list()
frag_pars.k[[1]] <- list() # first subpop
frag_pars.k[[1]][[1]] <- list(NULL, c(0), c(0, 0), c(0, 0.015, 0.015), c(0, 0.07, 0.08, 0.06)*frag_set[k]) 
frag_pars.k[[1]][[2]] <- frag_pars.k[[1]][[1]]

  
  for(j in 1:length(out_set)){
    
    # outplanting size 1
    out_pars <- matrix(0, nrow = years, ncol = n)
    out_pars[ ,1] <- rep(out_set[j])
    
    simjk <- popvi_mod(years = t_goal, n, A_mids, surv_pars.r, growth_pars.r, shrink_pars.r, frag_pars.k, fec_pars.r, lambda, sigma_s, sigma_f, ext_rand, seeds, dist_yrs,
                      dist_pars.r, dist_effects.r, out_pars, N0.r)
    
    pop1[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal])
    area1[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]*A_mids) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal]*A_mids)
    
    
    # outplanting size 2
    out_pars <- matrix(0, nrow = years, ncol = n)
    out_pars[ ,2] <- rep(out_set[j])
    
    simjk <- popvi_mod(years = t_goal, n, A_mids, surv_pars.r, growth_pars.r, shrink_pars.r, frag_pars.k, fec_pars.r, lambda, sigma_s, sigma_f, ext_rand, seeds, dist_yrs,
                      dist_pars.r, dist_effects.r, out_pars, N0.r)
    
    pop2[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal])
    area2[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]*A_mids) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal]*A_mids)
    
     # outplanting size 3
    out_pars <- matrix(0, nrow = years, ncol = n)
    out_pars[ ,3] <- rep(out_set[j])
    
    simjk <- popvi_mod(years = t_goal, n, A_mids, surv_pars.r, growth_pars.r, shrink_pars.r, frag_pars.k, fec_pars.r, lambda, sigma_s, sigma_f, ext_rand, seeds, dist_yrs,
                      dist_pars.r, dist_effects.r, out_pars, N0.r)
    
    pop3[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal])
    area3[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]*A_mids) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal]*A_mids)
    
    # outplanting size 4
    out_pars <- matrix(0, nrow = years, ncol = n)
    out_pars[ ,4] <- rep(out_set[j])
    
    simjk <- popvi_mod(years = t_goal, n, A_mids, surv_pars.r, growth_pars.r, shrink_pars.r, frag_pars.k, fec_pars.r, lambda, sigma_s, sigma_f, ext_rand, seeds, dist_yrs,
                      dist_pars.r, dist_effects.r, out_pars, N0.r)
    
    pop4[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal])
    area4[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]*A_mids) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal]*A_mids)
    
    # outplanting size 5
    out_pars <- matrix(0, nrow = years, ncol = n)
    out_pars[ ,5] <- rep(out_set[j])
    
    simjk <- popvi_mod(years = t_goal, n, A_mids, surv_pars.r, growth_pars.r, shrink_pars.r, frag_pars.k, fec_pars.r, lambda, sigma_s, sigma_f, ext_rand, seeds, dist_yrs,
                      dist_pars.r, dist_effects.r, out_pars, N0.r)
    
    pop5[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal])
    area5[j,k] <- sum(simjk$reef_pops_pre[[1]][[1]][,t_goal]*A_mids) + sum(simjk$reef_pops_pre[[1]][[2]][,t_goal]*A_mids)
    
  }
  
}
```




```{r}
plot(NULL, xlim=c(min(frag_set),max(frag_set)), ylim=c(0,50), ylab="Min. number outplanted", xlab="fragmentation", las = 1)
popn <- rep(NA, length(frag_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(pop1[,k] >= N_goal))]
}
lines(x = frag_set, y = popn, lty = 1, col = n_cols[1], lwd = 2)
popn <- rep(NA, length(frag_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(pop2[,k] >= N_goal))]
}
lines(x = frag_set, y = popn, lty = 2, col = n_cols[2], lwd = 2)
popn <- rep(NA, length(frag_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(pop3[,k] >= N_goal))]
}
lines(x = frag_set, y = popn, lty = 3, col = n_cols[3], lwd = 2)
popn <- rep(NA, length(frag_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(pop4[,k] >= N_goal))]
}
lines(x = frag_set, y = popn, lty = 4, col = n_cols[4], lwd = 2)
popn <- rep(NA, length(frag_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(pop5[,k] >= N_goal))]
}
lines(x = frag_set, y = popn, lty = 5, col = n_cols[5], lwd = 2)
legend("topleft", legend = as.character(c(1:5)), lty = c(1:5), bty = "n", ncol = 5, col = n_cols, lwd = 2)
```


```{r}
plot(NULL, xlim=c(min(frag_set),max(frag_set)), ylim=c(0,50), ylab="Min. number outplanted", xlab="External recruitment rate", las = 1)
popn <- rep(NA, length(frag_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(area1[,k] >= A_goal))]
}
lines(x = frag_set, y = popn, lty = 1, col = n_cols[1], lwd = 2)
popn <- rep(NA, length(frag_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(area2[,k] >= A_goal))]
}
lines(x = frag_set, y = popn, lty = 2, col = n_cols[2], lwd = 2)
popn <- rep(NA, length(frag_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(area3[,k] >= A_goal))]
}
lines(x = frag_set, y = popn, lty = 3, col = n_cols[3], lwd = 2)
popn <- rep(NA, length(frag_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(area4[,k] >= A_goal))]
}
lines(x = frag_set, y = popn, lty = 4, col = n_cols[4], lwd = 2)
popn <- rep(NA, length(frag_set))
for(k in 1:length(popn)){
  popn[k] <- out_set[min(which(area5[,k] >= A_goal))]
}
lines(x = frag_set, y = popn, lty = 5, col = n_cols[5], lwd = 2)
legend("topleft", legend = as.character(c(1:5)), lty = c(1:5), bty = "n", ncol = 5, col = n_cols, lwd = 2)
```


